---
format:
  pdf:
    documentclass: article
    cite-method: natbib
    biblio-style: plainnat
    link-citations: true
    fontsize: 12pt
    fig-pos: "H"
    number-sections: true
    include-in-header: preamble.tex
    include-before-body: before-body.tex
bibliography: UST_lit_fixed.bib
fig-cap-location: top
geometry: margin=1in
---




```{r}
#| label: setup
#| include: false
#| message: false
#| warning: false

#| label: setup
#| include: false
# 1. Load custom environment variables (RENV_PATHS_*)
.libPaths(c("C:/Users/kaleb/Documents/ust_ins_move_to_github/staging_library", .libPaths()))

# 2. Load required packages
library(tidyverse)
# 2. FORCE the root identification 
# This looks for the directory containing the 'Reports' folder
here::i_am("Reports/Paper/01_LUST_Insurance_Draft.qmd")

# 3. DEBUG: Print the detected root to your render console
# If this doesn't say ".../ust_ins_move_to_github", here is the problem.
message("CHECKPOINT: Project root is: ", here::here())
# 2. Load all required packages 
library(here)
library(tidyverse)
library(kableExtra)
# Add any other packages your script requires below
```


\clearpage

<!-- https://www.twdb.texas.gov/mapping/gisdata.asp - gis dat -->

# Introduction {#sec-intro}

Underground Storage Tanks (UST) holding motor fuels and other hazardous liquids sit beneath about half a million U.S. facilities, and leaks from these tanks are the leading documented source of groundwater contamination @epa_underground_2024. Cleanup is costly; median remediation expenses exceed $100 000 per site @marcus_going_2021, and the probability of a release rises sharply with observable characteristics such as tank age and whether the walls are single or double. Despite these predictable differences, thirty-eight states finance remediation through public trust funds that charge every facility the same premium, diluting incentives to invest in safer technology and crowding out private insurers @astswmo_archive.

Economic theory predicts that tying premiums to expected damages should induce firms to internalize environmental risk [@shavell_liability_1984; @polinsky_strict_1980], yet the relevant margin here is not liability per se but the rating regime under which liability is financed. Uniform-premium public funds operate as community-rated pools: every covered facility pays the same fee regardless of its observable risk characteristics, replicating the cross-subsidization and incentive dilution that @Pauly1974 identified in flat-rate insurance and that @RothschildStiglitz1976 showed prevents competitive separating equilibria. Under community rating, two distortions operate simultaneously. Adverse selection pushes high-risk facilities toward the subsidized pool and crowds private risk-rated alternatives out of the market. Moral hazard suppresses prevention investment because the marginal high-risk facility bears no incremental premium from operating leak-prone equipment. Empirical tests of this mechanism for USTs are scarce. Michigan's mid-1990s shift toward private, risk-rated coverage suggested large gains but relied on aggregate counts that could not isolate facility-level responses [@yin_can_2011; @yin_risk-based_2011]. Whether actuarial pricing disciplines high-risk owners without simply driving them out of the market therefore remains unresolved.


Texas provides a clean natural experiment. In 1999 the state closed its public fund overnight and required all owners to purchase private coverage whose premiums vary with observable risk. I assemble a new facility-year panel covering 1990 to 2021 that links EPA inventories with administrative microdata from the Texas Commission on Environmental Quality and comparable data from eighteen control states that retained trust funds. The identification strategy rests on a make-model comparison sample that matches Texas facilities to control-state counterparts on the observable risk characteristics, tank age bin and wall construction type, that the private market prices. Because the treatment effect is theoretically heterogeneous along these pricing dimensions (@sec-theory, Proposition 3), comparability on them is required for valid causal inference; @sec-data shows that matching on make and model is what earns parallel pre-trends.


The transition to risk-based insurance raised the annual tank closure probability 
by $0.44$ percentage points (s.e.\ $= 0.0024$) relative to matched control-state 
tanks, a 42 percent increase over the post-reform control mean of 1.04 percent. 
The Cox specification yields a hazard ratio of $1.251$ (s.e.\ $= 0.041$), 
indicating that surviving Texas tanks closed at a 25 percent higher daily rate 
than comparable control tanks after the reform. Both effects are driven by tanks 
above median age at the reform date: the pooled difference-in-differences 
coefficient falls to $0.0012$ (s.e.\ $= 0.0022$) once the age interaction is 
introduced, while the interaction term implies a 44 percent higher daily closure 
hazard for older Texas tanks (hazard ratio $= 1.436$, s.e.\ $= 0.084$). The 
treatment effect rises with tank age and single-wall construction, the two 
dimensions the private market prices, which is inconsistent with a general Texas 
economic shock.

**Related Literature**



A well-established body of theoretical work suggests that strict liability rules can be an effective policy instrument for environmental risks, particularly when paired with insurance contracts that align a firm's private incentives with social costs [@shavell_liability_1984; @shavell_model_1984; @shavell_liability_1982; @polinsky_strict_1980; @boyd_retroactive_1997; @shavell_chapter_2007]. A crucial condition for this alignment is the insurer's ability to accurately price the risk of environmental harm. This paper provides the first empirical test of this principle in the context of environmental liability, offering causal estimates of how firms respond to a transition from flat-fee to more efficient, cost-based insurance contracts.

The challenge of pricing environmental risk connects this study to the economics of insurance rating regimes. The pooling equilibrium maintained by uniform-premium state funds is the environmental analog of community rating: premiums are invariant to observable risk, which creates cross-subsidies that suppress precautionary investment and crowd out risk-differentiated private alternatives [@RothschildStiglitz1976; @Pauly1974; @CutlerZeckhauser2000]. Theory predicts that a shift to experience rating generates two behavioral responses. On the selection margin, high-risk facilities face a cost increase and may exit or upgrade. On the moral hazard margin, facilities that remain now face a direct financial return to risk reduction, inducing investment in safer technology. Identifying these margins separately requires within-facility panel variation that aggregate data cannot provide.
The empirical literature on rating regimes confirms both mechanisms, though rarely in the same market. In health insurance, risk adjustment reforms that moved Medicare Advantage toward experience rating reduced favorable selection but generated gaming responses among plans [@brown_how_2014; @mcwilliams_new_2012]. In credit markets, moving from flat-rate to risk-differentiated pricing improves screening and changes borrower composition [@einav_contract_2012; @einav_impact_2013; @nelson_private_2025], with efficiency gains depending on whether information asymmetries favor the insurer or the insured [@FinkelsteinMcGarry2006; @liberman_equilibrium_2018; @einav_io_2021]. In the environmental insurance context, @yin_risk-based_2011 find that switching from community-rated state funds to private coverage in Michigan reduced leak incidents by over twenty percent. The Texas reform studied here provides a sharper identification environment and facility-level data that allow the selection and moral hazard margins to be separated.

This paper also contributes to the literature on U.S. UST policy. While much of the existing work focuses on estimating the marginal damages of pollution from leaks [@marcus_going_2021; @zabel_hedonic_2012; @guignet_sell_2014; @guignet_impacts_2018; @guignet_contamination_2018; @walsh_contaminated_2017; @guignet_what_2013], this study is complementary, examining the design of management policies intended to prevent those damages.

This research is most closely related to studies of a similar insurance market reform in Michigan by @yin_can_2011, @yin_risk-based_2011, and @yin_environmental_2007. That body of work finds that transitioning to risk-based insurance reduced the number of small firms and reported leaks. However, those analyses relied on aggregate or cross-sectional data, which cannot control for unobserved facility-level heterogeneity and may be susceptible to omitted variable bias. This paper advances on prior work by adopting an empirical strategy similar to other modern studies of environmental liability [@boomhower_drilling_2019]. By using a rich, facility-level panel dataset and a comparison sample matched on observable risk characteristics, I implement difference-in-differences models that control for unobserved facility characteristics and common time shocks, providing more credible causal estimates of how firms adjust their technology, replacement, and exit decisions in response to cost-based pricing.

# Institutional Background {#sec-background}

## Underground Storage Tanks: Physical Characteristics, Regulatory Lock-in, and Environmental Risk {#sec-ust-physical}

The U.S. Environmental Protection Agency defines a UST system as any tank and connected underground piping with at least 10% of its volume below the surface [@epa_underground_2024]. The construction technology of a UST system is the primary determinant of its environmental risk profile and the observable margin on which actuarial pricing differentiates. The critical physical distinction within this capital stock is between single-walled systems (bare steel or fiberglass shells with no interstitial containment layer) and double-walled systems (which incorporate a secondary wall and continuous interstitial monitoring space).

The construction choice that creates this risk profile is economically consequential and effectively irreversible. Single-walled construction is cheaper to install: a 1987 GAO survey priced a standard 10,000-gallon single-walled unit at approximately $23,000 versus $39,000 for a comparable double-walled unit, a 67% premium that contemporaneous manufacturer filings confirm has persisted [@gao1987; @csi2021; @xerxes2018]. The savings are locked in at installation. Once a single-walled tank is in the ground, the owner cannot upgrade the wall construction without removing and replacing the tank entirely, a capital expenditure comparable to installing new equipment from scratch. The practical irreversibility of this choice is confirmed directly in the administrative data: in Texas from 1998 to 2021, only 0.003% of active facilities ever retrofitted tank construction prior to terminal closure. Wall type is therefore not a choice variable that responds to price signals at the intensive margin. It is a fixed, observable, and verifiable physical attribute, precisely the conditions required for an actuarially feasible pricing margin and the reason the private insurance market in Texas conditions premiums directly on it.

The regulatory framework reinforces this physical lock-in with a legal one. Under 40 CFR Part 280, UST systems are subject to vintage-differentiated technical standards: tanks installed under the 1988 regulations are grandfathered into the compliance classification they earned at the federal deadline, provided they remain unmodified. Any construction activity that alters the wall structure or piping configuration of an existing system, including the partial upgrades that would otherwise be the natural intermediate response to a rising insurance premium, triggers reclassification of the entire facility under the most current technical standards in effect at the time of modification. The 2005 and subsequent EPA technical revisions substantially raised those standards [@us_epa_milestones_2021]. An owner who initiates a partial upgrade to reduce premium costs therefore faces the prospect of a facility-wide compliance overhaul whose cost dwarfs the incremental premium savings.

This is a direct instance of the mechanism @gruenspecht_differentiated_1982
identifies in the context of auto emissions and that
@stavins_vintage-differentiated_2005 formalizes as a general property of
vintage-differentiated regulation: exempting existing capital from new standards
inadvertently extends the operating life of the oldest and most hazardous stock
by making any modification prohibitively costly. The modification trigger in
40 CFR Part 280 operates exactly as the analogous New Source Review provision
does in the Clean Air Act, where @bushnell_enforcement_2012 show that the threat
of reclassification deterred capital investment at grandfathered coal plants;
@heutel_plant_2011 provides structural estimates of the same retardation effect
for boiler vintages. The practical effect here is to make the choice set binary:
operate the existing system without modification, or close it entirely.
As @sallee_retiring_2023 and @jacobsen_regulating_2023 argue in the context of
vehicle fleet decarbonization, when new-capital regulation cannot reach the
existing stock, retirement-focused instruments are the appropriate complement.
The UST setting illustrates precisely why: incremental investment in risk
reduction is not an option the regulatory environment makes available, which is
why the extensive margin of closure is the sole behavioral margin the empirical
analysis can and should target.


The environmental challenge posed by underground storage tanks is fundamentally
a problem of managing an existing stock, not regulating new flows into it.
@fig-sw-dw-stock documents the composition of the active fleet from 1985 through
2020. Single-walled tanks accounted for approximately 83% of the active capital
stock in the mid-1980s and remained the dominant technology at the federal
compliance deadline in December 1998, when they still constituted roughly 73% of
active tanks. By 2020, after two decades of post-deadline attrition, their share
had fallen only to approximately 63%. Double-walled construction, by contrast,
grew from near zero at the start of the study window to just under 20% of the
fleet by 2020, reflecting that new installations post-1988 are overwhelmingly
double-walled while the legacy stock turns over slowly. The figure makes plain
that the trajectory of environmental risk in this market is determined almost
entirely by the pace at which the incumbent single-walled stock exits, not by
the character of new entrants. The remediation liability sitting beneath the
surface today was installed decades ago and cannot be neutralized by tightening
standards on new entrants. It can only be reduced by accelerating the exit of
the single-walled incumbents. Because the hazardous technology is already in the
ground, the relevant policy question is not how to screen entrants but how to
price the continued operation of the existing fleet so that owners internalize
the expected cost of a release.

```{r}
#| label: fig-sw-dw-stock
#| fig-cap: "Aggregate Composition of Active Underground Storage Tank Fleet by Wall Type, 1989--2020."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_SW_DW_StockLevel_Pooled.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Proportional share of the total active capital stock categorized as single-walled (solid line, circles) and double-walled (dashed line, triangles) across the pooled full sample. The vertical dashed line marks the federal EPA upgrade deadline of December 22, 1998. The gradual post-deadline decline of the legacy single-walled fleet, mirrored by the corresponding accumulation of double-walled infrastructure, reflects the slow attrition of existing capital rather than immediate regulatory replacement. This confirms that stock management, rather than new-flow regulation, governs the long-term trajectory of environmental risk. The combined shares may evaluate to slightly less than 100% in a given year due to a marginal fraction of active tanks with unclassified or unknown structural metadata.
\end{flushleft}

## The Federal Compliance Mandate and Pre-Treatment Dynamics {#sec-federal-mandate}

Federal regulation of USTs originates in the 1984 Hazardous and Solid Waste
Amendments to the Resource Conservation and Recovery Act, which directed the EPA
to establish comprehensive technical and financial standards for tank systems. The
EPA finalized these standards in 40 CFR Part 280 in 1988, requiring all
non-compliant tanks to be upgraded, replaced, or closed by a uniform federal
compliance deadline of December 22, 1998 [@EPA2017USTCompendium;
@us_epa_milestones_2021]. @fig-full-closures makes the consequences visible:
annual facility-level closure rates spike sharply in the years immediately
preceding the deadline, peak in 1997 and 1998, and fall equally sharply once
the deadline passes. This spike is a regulatory compliance event. It contaminates
the pre-1988 capital stock in Texas and the control states simultaneously and
cannot be interpreted as a behavioral response to insurance pricing. Texas
compounded this federal problem with a state-level one, imposing a staggered
upgrade schedule for pre-1988 tanks that ran from 1989 through 1993 and had no
analogue in any control state. The two problems together render the pre-1988
installation cohorts unsuitable for the identification design. @sec-mandate
documents the Texas staggered mandate in detail and shows the differential
pre-treatment dynamics empirically.



```{r}
#| label: fig-full-closures
#| fig-cap: "Full Sample Annual Closure Rates and the Federal Mandate Effect."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_01_Full_Sample_closure_rates.pdf")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Full sample closure rates for Texas and control-state facilities across the study window, illustrating the spike in closures driven by the federal EPA upgrade deadline of December 22, 1998, and the subsequent decline in annual exit rates as the surviving fleet aged.
\end{flushleft}



## Financial Responsibility and the State Trust Fund Regime {#sec-trust-fund}

The 1988 Financial Responsibility Rule (40 CFR Part 280, Subpart H) mandates that every UST owner demonstrate the financial capacity to fund corrective action and satisfy third-party liability claims, with minimum required coverage of \$1 million per occurrence (40 CFR §280, Subpart H). In the late 1980s, the nascent private environmental insurance market was unable to supply this coverage at affordable cost to the population of small, single-site operators that constitute the majority of UST owners. To prevent widespread non-compliance, many states established public trust funds financed by annual tank registration fees or fuel tax revenues, functioning as a publicly administered insurer of last resort. A defining feature of these funds is their premium structure: annual fees are set on a uniform, per-tank basis, invariant to any observable facility risk characteristic such as wall type, tank age, or leak detection technology. States retaining public funds consistently report that these funds cover 90% to 100% of all active facilities, as documented by the Association of State and Territorial Solid Waste Management Officials [@astswmo_archive]. This near-complete market saturation is the predicted equilibrium outcome: because the subsidized uniform-premium contract is priced below any risk-differentiated private alternative for high-risk facilities, private insurers are structurally unable to compete for the portion of the market with above-average risk. The resulting uniform-premium pooling equilibrium is the institutional baseline against which Texas's transition is evaluated, and it is precisely the community rating equilibrium that the insurance economics literature predicts will simultaneously generate adverse selection into the pool and suppress prevention investment at the facility level [@RothschildStiglitz1976; @Pauly1974; @CutlerZeckhauser2000].



## Texas's Policy Transition {#sec-tx-transition}

Texas's departure from the public fund model constitutes a sharp, administratively determined policy discontinuity whose timing was driven by fiscal insolvency rather than any observable change in the statewide distribution of facility-level risk. Texas established the Petroleum Storage Tank Remediation Fund (PSTRF) in 1989 under the Texas Water Code. By 1995, actuarial assessments confirmed that the fund's liabilities, driven by the accumulating backlog of legacy remediation claims, were structurally unsustainable. In response, the 74th Texas Legislature enacted House Bill 2587, which mandated the sunset of the PSTRF [@HB2587_analysis_1995]. The legislature set the fund closure cutoff at December 22, 1998, the same calendar date as the federal EPA upgrade deadline codified in 40 CFR §280.21. The alignment of these two dates is not coincidental: the legislature explicitly selected the federal mandate date as the administrative anchor for the PSTRF cutoff, ensuring the fund would cease accepting new claims at precisely the moment when the largest cohort of non-compliant tanks was required to be addressed [@TexWaterCode_26_3512]. Effective December 22, 1998, all Texas UST owners were required to secure private insurance or qualify for self-insurance by demonstrating sufficient financial assets.

## The Post-Transition Private Market in Texas {#sec-private-market}

The private market that absorbed the Texas UST population after December 1998 prices observable facility risk directly, generating the cross-sectional and intertemporal premium variation that identifies firm behavioral responses. From 2007 to 2021, the period for which granular financial responsibility administrative data are available from the Texas Commission on Environmental Quality (TCEQ), an average of 89.8% of active retail UST facilities in Texas satisfied their financial responsibility obligation through commercial private insurance. The remaining 10.2% relied on self-insurance mechanisms, available exclusively to large, financially capitalized owners who can satisfy stringent asset tests.

@fig-fr-coverage documents the market transition directly. Financial responsibility coverage rates and the count of covered facilities confirm that adoption of private insurance was rapid and near-universal immediately following the December 1998 cutoff. The transition was not gradual: virtually the entire Texas UST population moved from public to private coverage within the first year of the reform.

```{r}
#| label: fig-fr-coverage
#| fig-cap: "Financial Responsibility Compliance in Texas, 2007--2021."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "FigureA_TX_FR_Coverage_Composition.pdf")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```
\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Share of active Texas retail UST facilities satisfying the 
EPA financial responsibility requirement via private insurance, self-insurance, 
or other mechanisms, by year. Private insurance is the dominant compliance 
vehicle throughout the observed window. Source: TCEQ financial responsibility 
administrative records.
\end{flushleft}

@fig-fr-coverage plots the share of active Texas retail UST 
facilities satisfying their financial responsibility obligation under each 
compliance mechanism from 2007 to 2021, the period for which TCEQ 
administrative records are available. Private insurance is the dominant 
mechanism in every year, covering roughly 85 to 90 percent of facilities 
throughout the window. Self-insurance accounts for approximately 10 to 12 
percent, and all remaining mechanisms combined never exceed 3 percent. The 
composition is stable across the full 14-year period, confirming that 
private insurance is the near-universal compliance vehicle for the Texas 
UST population in the post-transition market.

```{r}
#| label: fig-fr-hhi
#| fig-cap: "Market Concentration in the Texas UST Private Insurance Sector, 2007--2021."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_TX_FR_Top5_Dominance_HHI.pdf")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Each bar shows the share of total facility-months covered 
by each insurer in that year (left axis). The dashed line plots the 
Herfindahl-Hirschman Index by year (right axis), where each point is one 
year. An HHI above 2,500 is conventionally classified as highly concentrated. 
Source: TCEQ financial responsibility administrative records.
\end{flushleft}


@fig-fr-hhi documents market structure among private insurers 
covering Texas UST facilities from 2007 to 2020. Each bar shows the share 
of total facility-months covered by each named insurer in that year, with 
gray capturing all remaining carriers not in the top group. The dashed line 
with points plots the Herfindahl-Hirschman Index (HHI) for each year, read 
off the right axis. Mid-Continent Casualty Company held the largest single 
share throughout the period, covering roughly 25 to 30 percent of 
facility-months in most years. The HHI remained stable at approximately 
1,500 from 2007 through 2017, indicating a moderately concentrated but 
competitive market. Beginning in 2018 several carriers exited the market, 
and the HHI rose sharply to above 3,000 by 2020, crossing the threshold 
conventionally associated with high concentration. The premium reconstruction 
in @sec-premium draws on Mid-Continent rate filings, whose sustained market 
leadership across the full window makes them the appropriate basis for the 
facility-level premium schedule.






## Private Insurance Premium Reconstruction {#sec-premium}
```{r}
#| label: fig-premium-level
#| fig-cap: "Reconstructed Annual Facility Premiums for Texas UST Facilities Covered by Mid-Continent Casualty, 2006--2024."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "FigureB_MidContinent_Pricing.pdf")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} The line plots the median annual base premium per facility 
by year among Texas UST facilities covered by Mid-Continent Casualty 
Company. The shaded band shows the range of premiums consistent with the 
$\pm 40$ percent schedule rating adjustment permitted in the filing, 
computed as 60 percent and 140 percent of the base manual premium. The 
median premium rose from approximately \$1,100 in 2006 to approximately 
\$2,500 by 2022, reflecting both aging of the insured fleet and rate 
schedule updates across filing periods. The number of covered facilities 
ranges from 203 in the earliest year to 8,278 at peak. Source: SERFF rate 
filings and TCEQ financial responsibility administrative records.
\end{flushleft}

Private insurers do not publicly disclose individual policy premiums. I 
reconstruct the annual premium schedule Texas facilities faced after the 
1999 transition by building a tank-level rating engine from the rate filings 
of Mid-Continent Casualty Company (NAIC 23418), obtained from the System for 
Electronic Rates and Forms Filing (SERFF). The filings cover four periods: 
2006--2014, 2014--2019, 2019--2021, and 2021 onward. Each filing provides 
the complete pricing algorithm: a per-tank annual base rate, an Increased 
Limit Factor (ILF), and a set of additive secondary risk loads applied to 
observable facility characteristics.

I estimate the annual premium $P_{ijt}$ for tank $j$ at facility $i$ in 
year $t$ as:

$$
P_{ijt} = \text{Base}_{t} \times \text{ILF}_{it} \times 
\left(1 + \sum_{k} \lambda_{k,t} \mathbf{X}_{ijt}\right)
$${#eq-premium}

where $\text{Base}_{t}$ is the per-tank base rate for filing period $t$, 
$\text{ILF}_{it}$ is the Increased Limit Factor for the elected coverage 
limit, and $\mathbf{X}_{ijt}$ is a vector of observable tank characteristics 
including age, wall construction type, piping type, and leak detection 
technology. The parameters $\lambda_{k,t}$ are additive percentage credits 
or debits assigned to each characteristic in filing period $t$.

Four implementation assumptions are required. First, I hold the base rate 
constant within each filing period at its filed value. Second, I assume 
every facility elects the minimum coverage limit required by federal law, 
\$1 million per occurrence and \$1 million in aggregate, which corresponds 
to an ILF of $1.18$ in the Mid-Continent schedule. Third, I assume every 
facility carries the base deductible of \$5,000. The filing defines lower 
premium factors for higher deductibles, ranging from $0.92$ at \$10,000 to 
$0.60$ at \$50,000, but the administrative records do not contain each 
facility's elected deductible. Fixing the deductible at \$5,000 overstates 
the true premium for facilities that elected a higher deductible, so the 
estimated premium spread between high-risk and low-risk tanks is 
conservative. Fourth, for risk characteristics absent from the 
administrative records, specifically flood zone classification, marina 
status, and aquifer proximity, I set the corresponding load to zero. These 
are positive surcharges in the filing, so omitting them understates the 
true premium for affected facilities.

The rate filing permits underwriters to adjust the manual premium by up to 
$\pm 40$ percent in total through six judgment categories covering location, 
premises condition, equipment, classification peculiarities, employee 
training, and safety programs. I use the base manual premium as the price 
signal and treat the discretionary component as mean-zero across the 
facility population. The reconstructed premiums cover facilities observed 
from 2006 to 2024, with 203 covered facilities in the first year rising to 
a peak of 8,278. The median annual facility premium rose from approximately 
\$1,100 in 2006 to approximately \$2,500 by 2022, reflecting both the aging 
of the insured fleet and rate schedule updates across filing periods. 
@fig-premium-level plots the median premium alongside the $\pm 40$ percent 
schedule bounds by year. Granular contract records are available beginning 
in 2006. The period from January 1999 through December 2005 represents a 
window in which the private market was operational but individual policy 
records are not observed.


## The Texas Phased Mandate: A Pre-Treatment Confound {#sec-mandate}

Texas's transition to private financial responsibility was preceded by a separate
regulatory regime that creates a pre-treatment confound for the oldest installation
cohorts. Under 30 TAC Chapter 334, the Texas Natural Resource Conservation
Commission assigned staggered upgrade deadlines to pre-1988 tanks by vintage
cohort, running from December 22, 1989 for the oldest tanks through December 22,
1993 for the 1985--1988 cohort (@tbl-mandate-schedule). These deadlines preceded
the uniform federal date by five to nine years depending on vintage. No control
state imposed an analogous staggered schedule; facilities in control states faced
only the federal deadline of December 22, 1998.

The consequence is survivor selection in the Texas pre-1988 stock. By the time the
insurance reform took effect in 1999, the pre-1988 Texas tanks still operating had
already survived a cohort-specific compliance decision made years earlier: their
operators had elected to invest in upgrades rather than close when confronted with
the rolling state mandate. This is not a random sample of the original pre-1988
population. It is the subset whose operators revealed a preference for continuing
operations under regulatory pressure, before any insurance price signal existed.
Control-state pre-1988 operators faced no such pre-screening. Their closure
decisions in the years around 1998 reflect the federal mandate alone, not a prior
self-selection event. Any naive comparison pooling pre- and post-1988 cohorts
therefore conflates two distinct mechanisms: mandate-driven survivor selection in
Texas and insurance-driven behavioral response, the object of interest.

@tbl-mandate-exposure documents this differential empirically. Cumulative closure
rates and surviving fleet characteristics diverge sharply between Texas pre-1988
cohorts and their control-state counterparts during 1989--1993, the window when
the staggered mandate was binding in Texas and silent in control states. No
corresponding divergence appears for post-1988 cohorts, which were never subject
to the staggered schedule and which enter the treatment window with comparable
pre-reform closure histories across Texas and the control group.

Both problems identified in this section and in @sec-mandate
are resolved by restricting the estimation sample to tanks installed between 1990
and 1997. This cohort never faced the staggered Texas mandate, was not subject to
the federal compliance deadline before 1998, and entered the treatment window with
a clean pre-reform closure history whose dynamics are comparable across Texas and
the control states. The restriction is documented in @tbl-attrition and the
  parallel pre-trends it earns are confirmed in the event study of @sec-results.




## Control Group Definition and Institutional Restrictions

The potential control group consists of all states that maintained a public 
uniform-premium trust fund continuously through the study window and were never 
subject to a mandatory transition to private financial responsibility. Approximately 
thirty states satisfy this institutional criterion. After applying four data 
completeness requirements dictated by the panel construction and identification 
design, the usable control group contracts to eighteen states.

**Missing installation dates.** The panel backbone requires a confirmed installation 
date for every tank regardless of status so that the facility can be assigned to an 
age cohort and the make-model cell can be frozen at the reform date. Where 
installation dates are missing at high rates, facilities cannot be assigned to an 
age bin, are excluded from the make-model sample, and generate attrition that is 
mechanically correlated with vintage — precisely the dimension the identification 
design exploits.

**Missing closure dates for tanks recorded as closed.** The panel assigns a closure 
event to a facility-year only when a closed tank has a confirmed closure date; tanks 
flagged as closed but lacking a closure date are dropped in the filtering step 
described in @sec-data and documented in @tbl-attrition. Where 
closure-date missingness is high, the denominator of the closure rate is correctly 
measured but the numerator is understated, biasing the outcome toward zero in the 
pre-treatment period and attenuating the DiD estimate.

**Missing wall-type classification.** The make-model comparison sample requires 
that each facility be classifiable as all-single-walled, all-double-walled, or 
mixed on the dimensions private insurers price. Wall-type is stored in the EPA 
tank-type field. Where this field is missing at high rates, the matching cell is 
undefined and the facility cannot contribute to the make-model sample regardless 
of its institutional eligibility.

**Missing LUST report dates.** The leak outcome requires a confirmed report date 
for each leaking underground storage tank release, used to assign the release to a 
facility-year and to classify it under the primary 0--60 day window. States where 
report-date missingness is total are excluded from the leak analysis sample; states 
with partial missingness are retained for the closure analysis but excluded from 
the leak specifications.

@tbl-data-quality documents missingness rates across all candidate states on each 
of these four dimensions. The twenty-one states that relied solely on the EPA 
harmonized database without a custom state registry fail one or more of these 
requirements at rates that make the panel unusable. Closure dates are missing for 
100 percent of closed tanks in New York, South Carolina, Missouri, Mississippi, 
Nevada, Nebraska, Georgia, New Hampshire, Rhode Island, Alaska, Utah, Indiana, 
West Virginia, Wyoming, and Hawaii, and for 80 percent in South Dakota. 
Installation dates are missing for 100 percent of tanks in South Carolina, 
Mississippi, and Vermont, and for more than 30 percent in New York, Missouri, 
Indiana, Illinois, and West Virginia. Delaware satisfies both date requirements 
but is excluded on a third ground: wall-type classification is missing for 100 
percent of its tank inventory, making make-model cell assignment impossible. Two 
further states — Washington and Oregon — are excluded on institutional rather than 
data-quality grounds: neither operated a public uniform-premium trust fund during 
the study window and so do not satisfy the control group eligibility criterion 
regardless of data quality.

The eighteen states that satisfy all four completeness thresholds and maintained public trust funds continuously through 2020 form the control group used in all baseline specifications. 

# Insurance Regimes and the Retirement of Risky Capital {#sec-theory}

The insurance regime governs whether a facility's pollution risk enters its cost function. This section formalizes that claim, derives the closure threshold under each regime, and produces the testable predictions that motivate both the make-model comparison sample and the empirical specifications in @sec-empirics. 

## Primitives {#sec-primitives}

The facility owner's problem is a single-agent dynamic discrete choice model in the optimal stopping tradition (Rust, 1987). In an infinite-horizon setting, a competitive facility owner operates a single underground storage tank and each period decides whether to continue operating or permanently close. Let $\mathbf{x} \in \mathcal{X}$ denote the vector of facility characteristics that determine pollution risk: tank age, construction type, wall material, soil composition, and any other observable feature correlated with the probability of an environmental release. The instantaneous leak hazard is $h(\mathbf{x}) > 0$. For analytical tractability, I model $\mathbf{x}$ as the scalar tank age $a$, with $h'(a) > 0$ and $h''(a) \geq 0$. The qualitative results hold for any dimension of $\mathbf{x}$ along which the insurer conditions the premium. The Extension section below states this generalization explicitly.

Each leak event generates two categories of social harm. Remediation cost $L > 0$ is the direct cost of cleaning contaminated soil and groundwater. The operator bears this cost under strict liability, making it the natural object of an insurance contract. Health externality $E > 0$ represents diffuse health damages accruing to surrounding communities through groundwater contamination. Because contamination is frequently undetected for years after a release and attribution to a specific facility is administratively difficult, these damages generate no insurable third-party claim (@marcus_going_2021). The gap between $L$ and $L + E$ defines the limits of what any privately-priced insurance arrangement can achieve relative to the social optimum, and it motivates the welfare decomposition below.

The owner earns per-period net revenue $R > 0$ from operating and recovers scrap value $\kappa > 0$ upon permanent closure, representing the liquidity of the land net of decommissioning costs. The owner discounts at rate $\beta \in (0,1)$ and is risk-neutral.

## Insurance Regimes {#sec-regimes}

The insurance regime determines whether and how the premium $C(\mathbf{x})$ varies with the facility's risk profile. I define three regimes that span the relevant policy space.

**Definition 1 (Uniform Premium, UP).** The facility pays a flat actuarially pooled premium each period, independent of any facility characteristic:

$$
C_{\text{UP}}(\mathbf{x}) = \bar{p}.
$$

No risk information is transmitted to the firm through the insurance contract under UP. In the scalar age case, the marginal cost of aging, the elasticity of the premium with respect to capital depreciation, is identically zero:

$$
\frac{\partial C_{\text{UP}}}{\partial a} = 0.
$$

The insurer may possess full information about the cross-sectional risk distribution but does not act on it in setting individual premiums. This is the pricing structure of state-administered petroleum storage tank assurance funds, including the Texas PST Fund prior to its 1999 sunset.

**Definition 2 (Risk-Based Pricing, RB).** The insurer prices actuarially, conditioning on all observable elements of $\mathbf{x}$. The premium equals expected claims scaled by a loading factor $\lambda \geq 1$:

$$
C_{\text{RB}}(a) = \lambda \cdot h(a) \cdot L.
$$

Since $h'(a) > 0$, the premium's elasticity with respect to capital depreciation is strictly positive:

$$
\frac{\partial C_{\text{RB}}}{\partial a} = \lambda h'(a) L > 0.
$$

Every dimension of $\mathbf{x}$ that enters $h(\mathbf{x})$ is mechanically injected into the firm's flow cost by the actuarial pricing rule. This is a definitional property of risk-based pricing, not a behavioral assumption about firm awareness or optimization.

**Definition 3 (Social Optimum, SOC).** The social planner internalizes both remediation costs and health externalities:

$$
C_{\text{SOC}}(a) = h(a) \cdot (L + E).
$$

The social optimum is unattainable under private insurance because $E$ generates no insurable claim. For the parametrically relevant condition $E > (\lambda - 1)L$, the cost ordering is:

$$
C_{\text{UP}} \leq C_{\text{RB}}(a) < C_{\text{SOC}}(a),
$$

where the left inequality becomes strict for $a$ large enough that $\lambda h(a) L > \bar{p}$.

## The Firm's Dynamic Problem {#sec-dynamic-problem}

Each period, the owner observes age $a$ and chooses to maintain or permanently close. Under regime $j$, flow utility from maintaining is $u_j(a) = R - C_j(a)$, where $R$ is annual net revenue from operations net of all non-insurance costs and $C_j(a)$ is the insurance premium. I maintain the simplifying assumption that insurance contracts provide full ex-post coverage with no deductible, treating environmental risk solely as an ex-ante premium cost rather than an ex-post liability shock. This assumption isolates the cost-of-continuing channel and is consistent with standard provisions in the petroleum storage tank insurance market; the structural estimation in @sec-structural-est relaxes it to incorporate partial risk internalization through deductibles.

The Bellman equation under regime $j$ is:

$$
V_j(a) = \max \left\{ \underbrace{R - C_j(a) + \beta \, \mathbb{E}[V_j(a') \mid a]}_{\text{Maintain}},\;\; \underbrace{\kappa}_{\text{Close}} \right\},
$$

where $\beta$ is the discount factor, the expectation is taken over the stochastic evolution of tank age, and $\kappa$ is the scrap value recovered upon exit. The pure continuation value, the present discounted value of flow utility under the assumption that the tank is never closed, is:

$$
\widetilde{V}_j(a) = \sum_{s=0}^{\infty} \beta^s \left[ R - C_j(a+s) \right].
$$

The optimal value function satisfies $V_j(a) = \max\{\widetilde{V}_j(a),\,\kappa\}$. The optimal closure threshold is:

$$
a_j^* = \min \left\{ a \geq 0 : \widetilde{V}_j(a) \leq \kappa \right\}.
$$

Under any regime with $C_j'(a) > 0$, the function $\widetilde{V}_j(a)$ is strictly decreasing, so the optimal policy is a threshold rule: close if and only if $a \geq a_j^*$.

## Main Results {#sec-theory-results}

**Proposition 1 (Risk Transmission and Value Slope).**
*Under UP, $\widetilde{V}_{\text{UP}}'(a) = 0$. Under RB, $\widetilde{V}_{\text{RB}}'(a) < 0$. Under SOC, $\widetilde{V}_{\text{SOC}}'(a) < \widetilde{V}_{\text{RB}}'(a) < 0$.*

*Proof.* Differentiating the continuation value with respect to age:

$$
\widetilde{V}_j'(a) = -\sum_{s=0}^{\infty} \beta^s C_j'(a+s).
$$

Under UP, $C_{\text{UP}}'(\cdot) = 0$ everywhere, so the sum is zero and the continuation value is flat in age: the physical degradation of the asset is real but generates no financial pressure on the firm. Under RB, $C_{\text{RB}}'(a) = \lambda h'(a) L > 0$, so every term in the sum is strictly negative and $\widetilde{V}_{\text{RB}}'(a) < 0$. Under SOC, $C_{\text{SOC}}'(a) = h'(a)(L+E) > \lambda h'(a) L = C_{\text{RB}}'(a)$, so the SOC slope is strictly more negative at every age. $\square$

Proposition 1 establishes the mapping from the pricing regime to the information environment of the firm. Under UP, aging is economically invisible: the zero elasticity of the premium with respect to capital depreciation means the continuation value contains no signal about the facility's pollution hazard, and the firm retains the asset until it is physically obsolete. Under RB, aging is economically visible by construction: the actuarial pricing rule translates the insurer's risk model into a per-period cost that rises with the facility's hazard profile, pivoting the continuation value downward for older tanks and accelerating the optimal exit age, regardless of whether the owner attends to or understands the actuarial logic.

**Proposition 2 (Regime Ordering of Closure Thresholds).**
*Under the condition $E > (\lambda-1)L$, the optimal closure ages satisfy $a_{\text{SOC}}^* < a_{\text{RB}}^* < a_{\text{UP}}^*$.*

*Proof.* By Proposition 1, $\widetilde{V}_{\text{SOC}}(a) < \widetilde{V}_{\text{RB}}(a) < \widetilde{V}_{\text{UP}}(a)$ for all $a > 0$. Since $a_j^*$ is defined as the first age at which $\widetilde{V}_j(a)$ crosses $\kappa$ from above, a lower value function crosses the threshold at a younger age. The ordering of value functions is preserved under discounted summation because $h(a)$ is common across regimes; only the coefficient multiplying $h(a)$ differs. $\square$

@fig-mechanism illustrates Proposition 2. The SOC curve declines at the fastest rate and crosses the scrap value line $\kappa$ at $a_{\text{SOC}}^*$. The RB curve declines at an intermediate rate and crosses $\kappa$ at $a_{\text{RB}}^* > a_{\text{SOC}}^*$, strictly reducing the optimal tank lifespan relative to the uniform pricing baseline. The UP curve is flat by Proposition 1 and does not cross $\kappa$ within the plotted range, reflecting the zero elasticity of the premium with respect to capital depreciation under pooled pricing.

![The Mechanism: Optimal Stopping under Alternative Insurance Regimes](Output/Figures/Paper/Fig_1_Mechanism.pdf){#fig-mechanism width=90%}

\vspace{0.2cm} 
\begin{flushleft} 
\footnotesize \textit{Notes:} This figure illustrates the optimal closure thresholds under the three insurance regimes derived in Proposition 2. The horizontal line represents the scrap value ($\kappa$). The Social Optimum (SOC) continuation value declines at the fastest rate, prompting the earliest closure at $a_{\text{SOC}}^*$. The Risk-Based (RB) pricing continuation value declines at an intermediate rate, yielding closure at $a_{\text{RB}}^*$. The Uniform Premium (UP) continuation value remains flat, reflecting the zero elasticity of the premium with respect to capital depreciation under pooled pricing.
\end{flushleft}


## Frictions Attenuating Behavioral Response {#sec-frictions}

The model establishes the direction of the effect of moving from UP to RB pricing. It does not establish the magnitude, and it is silent on two frictions that could attenuate behavioral response even when the directional incentive is correctly signed.

The first friction is premium insufficiency. Even if $C_{\text{RB}}'(a) > 0$, the premium increment at the margin may be small relative to the fixed transaction costs of closure, which include demolition, regulatory notification, site remediation, and foregone operating revenue. If the present value of avoided future premiums is less than these fixed exit costs for the marginal tank, the regime shift generates no behavioral response despite correctly pricing risk.

The second friction is constraints on action. Even when $\widetilde{V}_{\text{RB}}(a) < \kappa$ and closure is optimal, the owner may be unable to act on that signal. Credit constraints may prevent financing upfront exit costs. Thin markets for contaminated properties may render the book value of $\kappa$ unrealizable. Present bias or inattention to the actuarial premium schedule could cause owners to underweight future premium increases.

Neither friction is resolvable from theory alone. The Texas natural experiment provides the empirical test: the 1999 sunset of the PST Fund was abrupt, mandatory, and economically large, providing the variation required to estimate whether the premium-burden mechanism is sufficient to produce measurable changes in closure behavior.

**Proposition 3 (Orthogonality Under UP; Correlation Under RB).**
*Let $\xi$ denote closure drivers unrelated to the facility's pollution hazard, land value appreciation, local fuel demand shocks, and owner liquidity or retirement decisions, satisfying $\xi \perp a$. Then: (i) under UP, $\mathrm{Cov}(\mathbf{1}[\mathrm{Close}],\, a) = 0$; (ii) under RB, $\mathrm{Cov}(\mathbf{1}[\mathrm{Close}],\, a) > 0$, and the treatment effect $\Delta(a) \equiv \Pr(\mathrm{Close} \mid \mathrm{RB}, a) - \Pr(\mathrm{Close} \mid \mathrm{UP}, a)$ is zero for $a < a_{\text{RB}}^*$ and strictly increasing in $a$ for $a \geq a_{\text{RB}}^*$.*

*Proof.* Under UP the only closure drivers are elements of $\xi$, which satisfy $\xi \perp a$ by assumption, so $\mathrm{Cov}(\mathbf{1}[\mathrm{Close}], a) = 0$. Under RB, $\widetilde{V}_{\text{RB}}(a)$ is strictly decreasing, so $\Pr(\widetilde{V}_{\text{RB}}(a) \leq \kappa)$ is strictly increasing in $a$ and the covariance is positive. For the treatment effect: when $a < a_{\text{RB}}^*$, $\widetilde{V}_{\text{RB}}(a) > \kappa$ and $\widetilde{V}_{\text{UP}}(a) > \kappa$ under both regimes, so $\Delta(a) \approx 0$. When $a \geq a_{\text{RB}}^*$, $\widetilde{V}_{\text{RB}}(a) \leq \kappa < \widetilde{V}_{\text{UP}}(a)$, so the regime difference is the binding margin and $\Delta(a) > 0$, strictly increasing as the gap $\widetilde{V}_{\text{UP}}(a) - \widetilde{V}_{\text{RB}}(a)$ widens with age. $\square$

The $\xi \perp a$ assumption warrants comment. It requires that non-insurance closure drivers, including owner retirement, land value appreciation, and local demand shifts, are mean-independent of tank age. This is an identifying assumption, not a definitional one. If owners of older tanks are themselves older and more likely to retire, or if facilities in older commercial districts face systematically different land value trajectories, the orthogonality would be violated. The pre-period evidence in @sec-results provides an indirect test: under UP, control-state closure rates should exhibit no age gradient in the pre-treatment period. The parallel trends evidence is consistent with this condition, though it cannot rule out violations that are small relative to sampling noise.

Proposition 3 is the identifying prediction of the paper. The shift from UP to RB does not merely change the number of tanks that close; it reorganizes *which* tanks close. Under UP, the age distribution of closures is determined entirely by land value appreciation, demand contraction, and owner liquidity or retirement, drivers that are mean-independent of the facility's hazard profile. Under RB, the age distribution of closures becomes correlated with the hazard profile in a specific, premium-gradient-shaped way: the treatment effect is zero for young tanks and rising for old tanks, with the crossover at $a_{\text{RB}}^*$.

This heterogeneous treatment effect has a direct implication for identification. If the treatment effect is concentrated among older, higher-risk facilities, a comparison group that is unbalanced on age or wall type will produce biased estimates, not because the treated and control groups follow different trends on average, but because the composition of each group puts different weight on the margin where the treatment bites. The make-model comparison sample, introduced in @sec-data, is constructed precisely to address this: by matching on age bin and wall type, the comparison group is balanced on the dimensions along which $\Delta(a)$ varies, removing the compositional source of bias.

## Welfare Decomposition {#sec-welfare-decomp}

The welfare cost of retaining a tank beyond the socially optimal closure age is the expected externality accumulated during the excess operating period. The total welfare loss from UP decomposes into two components:

$$
\mathrm{DWL}_{\text{UP}} = \underbrace{\int_{a^*_{\text{SOC}}}^{a^*_{\text{RB}}} h(a) \, E \, da}_{\text{Unavoidable Wedge}} + \underbrace{\int_{a^*_{\text{RB}}}^{a^*_{\text{UP}}} h(a) \, E \, da}_{\text{Eliminable Excess}}.
$$

The unavoidable wedge exists because $E$ generates no insurable claim, so the gap $a^*_{\text{RB}} - a^*_{\text{SOC}}$ cannot be closed by insurance design alone. The eliminable excess is the welfare loss attributable specifically to flat-fee pooling, the component recoverable by moving from UP to RB pricing.

![Welfare Analysis: The Externality Wedge and Partial Correction](Output/Figures/Paper/Fig_2_Welfare.pdf){#fig-welfare width=90%}

## Extension to the Full Risk Vector {#sec-extension}

The model uses age as a scalar proxy for $\mathbf{x}$ for tractability. Propositions 1 through 3 hold for any dimension of $\mathbf{x}$ along which (i) the hazard $h(\mathbf{x})$ varies and (ii) the insurer conditions the premium. The mechanism is identical: any risk factor that enters the actuarial pricing schedule is mechanically injected into the firm's flow cost, raises the elasticity of the premium with respect to that risk dimension, and generates closure sensitivity that is absent under flat-fee pooling.

This extension has a concrete empirical counterpart in the petroleum storage tank market. Private insurers in Texas condition premiums on both age and tank wall construction type, with single-wall tanks facing premiums approximately twice those of double-wall tanks of equivalent age, as documented in Mid-Continent rate filings for the 2007 to 2021 period. If the mechanism of Proposition 3 extends to all priced risk dimensions, closure decisions under RB pricing should become correlated with wall type in the same way they become correlated with age. I treat this as an empirical generalizability test: a positive result for wall type corroborates the premium-burden channel; a null result would indicate either that wall-type premium gradients are insufficient to move behavior or that single-wall operators face structural exit barriers distinct from those facing aged-tank operators.

## Testable Predictions {#sec-hypotheses}

The following four hypotheses map directly from the propositions above to the difference-in-differences specifications in @sec-empirics. In each case, the empirical test compares Texas facilities, which transitioned from UP to RB pricing in 1999, to make-model-matched control-state facilities, which retained flat-fee public insurance programs throughout the sample period.

**Hypothesis 1 (Age Sensitivity of Closure).** Under UP, the closure hazard is mean-independent of $a$ (Proposition 3, part i). Under RB, the positive elasticity of the premium with respect to capital depreciation creates a premium-burden channel that makes older facilities more likely to close (Proposition 3, part ii). The prediction is that the age gradient in the closure hazard steepened differentially in Texas after 1999:

$$
\frac{\partial \lambda(t \mid a, \text{RB})}{\partial a} - \frac{\partial \lambda(t \mid a, \text{UP})}{\partial a} > 0.
$$

A discrete-time hazard model with the triple interaction $\text{TX} \times \text{Post} \times a$ is the natural estimator. The linear probability model on the binary closure indicator is a valid approximation under the assumption that censoring is independent of age conditional on covariates; the survival model relaxes this assumption.

**Hypothesis 2 (Heterogeneous Treatment Effect by Age).** The treatment effect of switching from UP to RB is zero for young tanks and strictly positive and increasing for old tanks (Proposition 3). For cutoff age $\bar{a}$:

$$
\mathrm{ATT}(a > \bar{a}) > \mathrm{ATT}(a \leq \bar{a}) \approx 0.$$

**Hypothesis 3 (Rightward Shift in Age Distribution of Closures).** The transition to RB pricing triggers a concentrated wave of exits among older tanks, raising the expected age of a closing facility relative to control-state counterparts:

$$
\mathbb{E}[a \mid \text{Close, Texas, Post}] > \mathbb{E}[a \mid \text{Close, Control, Post}].
$$

The mean shift is the testable corollary of a richer distributional prediction: under RB, the age-at-closure distribution concentrates mass at high ages near $a^*_{\text{RB}}$, whereas under UP the distribution of age at closure mirrors the age distribution of the full operating population. A test of the mean shift conditions on the selected sample of closures, which introduces selection bias because the tanks that close under RB are a non-random subset of the population. The density of age at closure by group and period provides the non-parametric visual companion to the regression test.

**Hypothesis 4 (Wall-Type Sensitivity).** If the mechanism extends to all priced dimensions of $\mathbf{x}$, the single-wall versus double-wall gap in the closure hazard should widen under RB pricing relative to UP:

$$
\left[\lambda(t \mid \text{SW, RB}) - \lambda(t \mid \text{DW, RB})\right] > \left[\lambda(t \mid \text{SW, UP}) - \lambda(t \mid \text{DW, UP})\right].
$$

The empirical test is the triple interaction $\text{TX} \times \text{Post} \times \text{SingleWall}$ in the closure specification.

Three empirical patterns would be inconsistent with the premium-burden channel and would instead indicate confounding from a Texas-specific economic shock: a uniform treatment effect across all age and wall-type groups; a closure response concentrated among young or low-risk facilities; or no correlation between the treatment effect magnitude and observable risk factors. These patterns define the falsification criteria for the identification strategy in @sec-empirics.

\clearpage


# Data {#sec-data}

## Source Records

The analysis draws on administrative records from three sources. The first is a
tank-level physical inventory assembled from state UST regulatory databases across
all fifty states, harmonized into a common schema. Each record identifies a
physical tank by facility, state, and tank identifier, and reports the tank's
installation date, closure date (if closed), wall construction (single-walled,
double-walled, or unclassified), stored substance (gasoline, diesel, other
petroleum products), and nominal capacity in gallons. State databases differ in
field naming, date formatting, and classification conventions; the harmonization
standardizes these fields, resolves coordinate systems for facility geolocation,
and backfills missing county identifiers from ZIP code centroids where native
geocoding is unavailable. The resulting file contains one row per physical tank
across all states.

The second source is a harmonized incident file of confirmed releases reported
under the Leaking Underground Storage Tank (LUST) program. Each record identifies
a facility, state, and the date the release was reported to the state regulatory
authority. States that lack report dates for all or nearly all LUST incidents are
excluded from the leak analysis, as described below.

The third source, available for Texas only, is the financial responsibility
compliance record maintained by the Texas Commission on Environmental Quality
(TCEQ). This file records, at monthly frequency, each facility's compliance
mechanism (private insurance, self-insurance, or no coverage), the identity of
the insurer where applicable, and the contract detail type. These records are
available from 2007 onward and are used to document the post-transition private
market structure described in @sec-private-market.


## Panel Construction

The unit of observation for the behavioral analysis is the facility-year. I
construct the panel by first computing each facility's active window, the span
from its earliest tank installation to its latest tank closure or the end of the
study period, and expanding the facility to a monthly time series over that range.
For each facility-month, the tank inventory is queried to determine the count and
characteristics of tanks physically active on that date: total active tanks,
aggregate capacity, average tank age (computed from exact installation dates),
counts of single-walled and double-walled tanks, and fuel-type indicators.
Monthly events (confirmed releases, tank closures, and new tank installations) are
counted and merged to the facility-month backbone.

Tank closures receive a type classification before aggregation. For each closed
tank, I determine whether the facility installed any replacement tank after that
closure date. If so, the closure is classified as a *replacement*; if not, as
*permanent*. The classification uses the facility's full installation history,
including tanks outside the analysis window, to avoid misclassifying replacement
closures as permanent when the replacement tank falls outside a restricted sample
frame.

The monthly panel is then collapsed to annual frequency. Stock variables (active
tank count, capacity, tank age, wall composition) are measured at the December
snapshot. Flow variables (closures, installations, confirmed releases, retrofits)
are summed over the calendar year, with binary indicators for whether any event
of each type occurred. Year-over-year changes in tank count and capacity are
computed as first differences of the December stocks.

Each confirmed release is paired with tank closures at the same facility and
classified by timing. A release reported within 60 days after a closure is
classified as *closure-revealed*: the closure inspection likely discovered a
latent leak. A release on record more than 180 days before a closure is
classified as *known pre-existing*: the tank was already leaking. Releases with
no nearby closure are classified as *standalone*. Pairs are restricted to the
same generation of infrastructure: a release that predates a tank's installation
is not matched to that tank. Robustness checks use 30-day and 90-day windows
for the closure-revealed category.

For the duration analysis of individual tank closures, a second panel is
constructed at the tank-year level directly from the physical inventory, without
passing through the facility-level aggregation. Each tank is expanded from its
installation year through its closure year (or through 2020 for tanks still
active), yielding one row per tank per calendar year of service. The closure
event indicator equals one in the year the tank was closed. For the primary Cox
specification, exact installation and closure dates are retained and the
counting-process interval for each tank is split at the statutory reform date of
December 22, 1998, producing at most two records per tank: a pre-reform interval
and a post-reform interval. This construction allows the treatment indicator to
switch on at the precise reform date rather than at an annual boundary.


## Study States

### Control Group Selection Criteria

The control group comprises states satisfying two institutional criteria and four
data-quality criteria assessed on the 1990--1997 installation cohort that forms
the primary analysis sample.

**Institutional criteria.** A state must have maintained a public
uniform-premium trust fund continuously through the study period, providing the
institutional counterfactual of continued flat-fee coverage. Washington and Oregon
are excluded because neither operated such a fund. Six additional
states (Arizona, Connecticut, Florida, Iowa, Michigan, and Wisconsin) are
excluded because they underwent their own transitions to private or risk-based
insurance during the study period and therefore cannot serve as controls.

**Data-quality criteria.** I evaluate four dimensions of record completeness,
assessed specifically within the 1990--1997 installation cohort rather than
across all historical records. The cohort-specific assessment is important because
missingness rates often differ substantially between newer and older inventory,
and the analysis sample is drawn exclusively from this cohort.

*Wall-type classification.* The comparison design matches Texas and control-state
facilities on wall composition, which is the primary risk-pricing variable.
States where wall type is recorded as unknown for more than 50 percent of
1990--1997 cohort tanks cannot contribute meaningfully to within-cell
identification. Alabama (80.4\% unknown wall), Illinois (100\%), and several
smaller states fail this criterion and are excluded.

*Closure recording rate.* The closure outcome requires that tanks marked as
closed in the administrative record carry a closure date. Some states record all
tanks as perpetually open regardless of actual status, with closure
information held separately or not digitized. I measure this using the
\textit{closure-with-date rate}: the share of 1990--1997 cohort tanks that are
recorded as closed with a valid closure date. Texas's closure-with-date rate of
37.2 percent serves as the benchmark. States whose rate is near zero despite the
cohort being 25--35 years old are presumed to have a systematic recording failure.
Georgia and West Virginia both show 0.0\% closure-with-date rates and are excluded on
this basis. California similarly records 100\% of its 1990--1997 tanks as
Open, yielding a 0.0\% closure-with-date rate; it is excluded despite
otherwise clean wall-type data.

*LUST report dates.* States where LUST report dates are missing for 100 percent
of incidents cannot contribute to the leak-outcome specifications. Kentucky
(100\% missing) is excluded from the full analysis sample on this basis.
Pennsylvania is retained in the closure analysis per prior institutional knowledge
of its data structure, but excluded from leak specifications.

*Closed-without-date rate.* Among tanks the state does record as closed, the
share lacking a closure date should be low. States where more than roughly 30
percent of closed tanks are missing dates are flagged; none reach the threshold
for outright exclusion, but Arkansas (66\% missing LUST dates) and Tennessee
(59\% missing LUST dates) are excluded from leak specifications while retained
for closure analysis.


### Resulting Control Groups

The **primary analysis sample** comprises Texas and 17 control states: 
Arkansas, Colorado, Idaho, Kansas, Kentucky, Louisiana, Massachusetts, 
Maryland, Maine, Minnesota, Missouri, North Carolina, Ohio, Oklahoma, 
South Dakota, Tennessee, and Virginia. This sample is used for all 
closure-outcome regressions. The panel contains 318,668 facility-years 
across 18 states for the 1990-entry cohort over the 1990--2018 window.

<!-- The **LUST-clean sample** is a subset of 9 control states that 
additionally satisfy the LUST-date completeness requirement: Idaho, Kansas, 
Louisiana, Massachusetts, Maine, Minnesota, North Carolina, Oklahoma, and 
Virginia. This sample is used for all leak-outcome regressions. It contains 
198,742 facility-years across 10 states. The main treatment effect estimate 
is stable across the two samples ($\hat{\beta} = 0.0139$ in the primary 
sample versus $\hat{\beta} = 0.0185$ in the LUST-clean sample), confirming 
that the state-composition difference does not materially affect 
identification. -->


### Analysis Sample Restriction

Within the selected states, the primary analysis sample is further restricted to
facilities whose first tank installation occurred in 1990 or later. This
restriction serves two purposes. First, it eliminates facilities that entered
during or before the 1988--1989 federal EPA upgrade mandate transition period,
which imposed differential closure pressure across states depending on enforcement
timing and tank age cohorts. Pre-1990 entrants therefore faced a confounding
closure incentive unrelated to the Texas insurance reform. Second, it produces a
pre-reform panel window with stable parallel trends: the event study shows
pre-period coefficients that are small and statistically indistinguishable from
zero for the 1990-entry cohort, whereas extending the sample back to 1989 or
earlier reintroduces marginally significant pre-period differentials driven by the
thin early cohorts (342 Texas facilities at $t = -9$ versus 2,734 at $t = -1$).



## Outcome Variables

The analysis estimates effects on two groups of outcomes, measured at the 
tank-year level.

**Closure.** The primary outcome is a binary indicator equal to one in the 
year a tank is closed. Closures are decomposed into *replacement closures*, 
where the facility later installed a new tank, and *permanent closures*, 
where no subsequent installation occurred. The closure decision is made at 
the facility level but is measured at the tank level. Tank-level 
measurement is preferred because it identifies exactly which physical 
assets are retired and when, and because the Cox proportional hazard model 
requires tank-level spell data with exact installation and closure dates.

**Leak incidence.** The total confirmed release indicator equals one 
if the facility reported any LUST incident in a given year. This is 
decomposed into closure-revealed releases (reported 0 to 60 days after a 
tank closure), standalone releases (no nearby closure), and closures of 
tanks with a known pre-existing leak (on record more than 180 days before 
closure). The decomposition separates mechanical effects of increased 
closure activity from changes in leak incidence independent of the closure 
margin.


## Descriptive Statistics

@tbl-descriptive reports the 1998 cross-section and pre-reform flow 
averages for both the full analysis sample and the matched comparison 
sample, separately for Texas and the control group. Panel A describes the 
capital stock at the reform date: facility count, total tanks, average 
tanks per facility, average capacity, average tank age, age of the oldest 
tank, fleet-level single-wall share, and within-facility single-wall share. 
Panel B reports pre-reform annual outcome rates from 1990 to 1998. Panel C 
documents the cell structure of the matching design.
```{r}
#| label: tbl-descriptive
#| tbl-cap: "Sample Characteristics: Full Analysis Sample vs. Make-Model Comparison Sample."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "Table0_Descriptive.tex")
if (file.exists(table_path)) {
  lines <- readLines(table_path)
  
  lines <- lines[!grepl(
    "^\\s*\\\\begin\\{table\\}|^\\s*\\\\end\\{table\\}|^\\s*\\\\centering|^\\s*\\\\begingroup|^\\s*\\\\endgroup",
    lines
  )]
  
  cat("\\scalebox{0.72}{\n")
  cat("\\begin{minipage}{1.39\\linewidth}\n")
  cat("\\begingroup\n")
  cat("\\fontsize{9}{11}\\selectfont\n")
  cat(lines, sep = "\n")
  cat("\\vspace{0.2cm}\n")
  cat("\\par\\footnotesize\n")
  cat("\\textit{Notes:}\n")
  cat("Panel A reports the December 1998 cross-section. Both column groups show\n")
  cat("pooled means; parentheses report standard errors of the mean. State-fleet\n")
  cat("rows are population-weighted aggregates across all active tanks.\n")
  cat("Panel B reports pre-reform annual outcome rates (1990 to 1998).\n")
  cat("Panel C reports the cell structure of the matching design.\n")
  cat("Diff.\\ $=$ TX $-$ CTL (Panels A and B); stars from Welch $t$-tests\n")
  cat("(Panel A full-sample columns only).\n")
  cat("$^{*}p<0.10$, $^{**}p<0.05$, $^{***}p<0.01$.\n")
  cat("\\endgroup\n")
  cat("\\end{minipage}\n")
  cat("}\n")
  
} else {
  cat("\\textit{Table not found.}")
}
```

# Observable Risk Characteristics and the Actuarial Pricing Foundation {#sec-risk}

Risk-based insurance can only discipline facility operators if the characteristics
private insurers observe actually predict future environmental liability. If tank
age and wall construction type were unrelated to release probability and cleanup
cost, insurers would have no basis for charging different premiums across
facilities, and the premium gradients documented in @sec-premium would be noise
rather than signal. This section shows that observable facility characteristics
predict confirmed releases out of sample. This establishes that the premium
schedule in the rate filings reflects genuine actuarial content, and it motivates
the comparison sample construction in the following section.

## Release Probability Is Predictable from Observables {#sec-risk-predict}

I estimate the annual probability that a facility experiences its first confirmed
release, conditional on having never previously leaked. The estimand is the
first-release hazard in the never-yet-leaked risk set:

$$
h(t \mid x) = P(\text{first confirmed release in year } t \mid
\text{no prior release through } t{-}1,\; x)
$$

This directly mirrors the insurer's underwriting problem: given what is observable
about a facility at the start of a policy year, what is the probability of a claim?

The model is a penalized logistic regression estimated on all study states in the
pre-treatment period. The feature space is saturated: tank age enters as a set of
three-year group indicators with a full interaction with wall construction type,
allowing the model to fit a distinct risk gradient at each age step for
single-walled and double-walled facilities rather than imposing a common linear
slope. State fixed effects capture baseline hazard differences across the twenty
study states. Fuel type and portfolio size enter as additional controls along with
their pairwise interactions. Lasso regularization performs automatic variable
selection, shrinking the coefficients of uninformative terms to zero while
retaining the predictors that carry genuine signal. Out-of-sample predictions are
obtained through cross-validation: each observation is scored exclusively by model
fits on the folds that excluded it, so every predicted probability is genuinely
out-of-sample. A post-hoc recalibration step corrects for the probability shift
introduced by inverse-frequency class weighting, mapping predictions back to the
original event-rate scale.

First-release events are rare, at an annual base rate of 0.699 percent across
the estimation sample of 3,929,425 pre-treatment facility-years and 27,455
confirmed releases, a class ratio of approximately 142 non-event facility-years
for each release. Without correction the model learns to predict the majority class
everywhere. Inverse-frequency case weights equalize the effective class sizes
during estimation, forcing the model to recover the signal in the minority class.

The model's ability to rank facilities by release risk is summarized by its lift
over random selection. @fig-lift shows the cumulative fraction of confirmed
releases captured as facilities are screened in descending order of predicted risk.
Screening the top decile of facilities captures 42 percent of all confirmed
releases in the estimation sample, compared to ten percent under random selection,
a lift of 4.2. The top quintile captures 61 percent of releases. Observable
facility characteristics therefore produce a ranking with substantial actuarial
content.
```{r}
#| label: fig-lift
#| fig-cap: "Cumulative lift curve for predicting first confirmed release. Facilities are ranked in descending order of predicted risk. The dashed diagonal is the random-selection baseline."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "FigureA_CV_Lift_State.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Out-of-sample predictions from a saturated penalized logistic
regression estimated on pre-treatment facility-years across all study states.
The feature space includes wall type and age group indicators with full
interactions, state fixed effects, and fuel and portfolio controls. The dashed
diagonal is the no-skill baseline. Dotted lines mark the top-decile capture rate.
\end{flushleft}

Lift curves measure ranking ability but do not confirm that predicted scores are
reliable probability estimates. @fig-pr-curve presents the precision-recall curve,
which is the appropriate summary for rare-event classifiers. The no-skill baseline
is the marginal release rate of 0.70 percent, the precision obtained by flagging
facilities at random. The curve lies above this baseline across the full range of
recall, with a precision-recall area under the curve of 0.029. Because the
baseline in a precision-recall curve equals the event rate rather than a diagonal
through the unit square, any area above it reflects genuine discrimination of
high-risk from low-risk facilities given the rarity of releases.
```{r}
#| label: fig-pr-curve
#| fig-cap: "Out-of-sample precision-recall curve for predicting first confirmed release. The dashed horizontal line is the no-skill baseline equal to the marginal release rate."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_CV_PR.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Out-of-sample precision-recall curve from the saturated penalized
logistic regression. The dashed horizontal line is the no-skill baseline equal to
the marginal first-release rate (0.70\%). Precision-recall area under the curve
= 0.029.
\end{flushleft}

Good discrimination does not imply that the predicted scores are reliable
probability estimates. @fig-calibration plots the mean observed release rate
against the mean predicted score within each decile of the predicted score
distribution. The points track the 45-degree line closely across the full range,
with no systematic departure in the low or middle deciles and modest
over-prediction only in the top decile, confirming that the model outputs can be
interpreted as annual release probabilities rather than an ordinal score of
unknown scale. This calibration is a necessary condition for using predicted scores
as rate-cell inputs in the premium schedule comparison of @sec-premium.
```{r}
#| label: fig-calibration
#| fig-cap: "Calibration of predicted risk scores against observed release rates, by decile of the predicted score distribution. Points on the 45-degree line indicate perfect calibration."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_CV_Calibration.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Each point is a decile of the predicted score distribution. Mean
observed first-release rate on the vertical axis; mean predicted probability on the
horizontal axis. Point size is proportional to the number of facility-years in the
decile. The dashed line is the 45-degree reference. 3,929,425 facility-years;
27,455 first-release events.
\end{flushleft}

@fig-cell-risk shows the direction and shape of the predicted risk gradients by
aggregating out-of-sample predictions within each rate cell defined by wall
construction type and tank age group. Each cell mean reflects the model-estimated
risk of the facilities that actually belong to that cell, adjusted for the full set
of observable covariates and predicted on data the model did not see during
training. For single-walled facilities the predicted release rate rises steeply
with age, from 5.3 per thousand facility-years in the youngest age group to more
than 12 per thousand from the 18-20 year group onward, with a compression in the
three-to-five year range that likely reflects the early-period composition of the
stock. Double-walled facilities show a substantially flatter profile, remaining
near 3.5 to 5 per thousand across most age groups. The gap between wall types
widens with age: by the 15-17 year group single-walled facilities carry a
predicted rate of 10.1 per thousand facility-years compared to 4.5 for
double-walled facilities, a differential of 5.6 per thousand, and the gap reaches
6.9 per thousand by the 18-20 year group. These gradients are estimated from
pre-treatment data without imposing any structure on their shape, and they
correspond closely to the rate differentials in the Mid-Continent schedule
documented in @sec-premium.
```{r}
#| label: fig-cell-risk
#| fig-cap: "Out-of-sample predicted first-release rate by tank age group and wall construction type. Each point is the average cross-validated predicted probability within the cell, reported per 1,000 facility-years."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_CV_CellRisk.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Average out-of-sample predicted probability within each
(age group, wall type) cell from the saturated penalized logistic regression.
Single-walled (solid, orange) and double-walled (dashed, blue). Reported per
1,000 facility-years. Predictions are recalibrated to the original event-rate
scale prior to aggregation.
\end{flushleft}

@fig-cell-gof confirms that the model-implied cell rates closely recover the
empirical rates observed in the data. Each of the eighteen rate cells is plotted
against its corresponding observed mean release rate. The points lie along the
45-degree line across the full range of cell risk, from roughly three per thousand
at the low end to thirteen per thousand at the high end, with no systematic bias
in either direction. The largest departure between predicted and observed rates
occurs among the youngest double-walled facilities, where the model predicts 7.1
per thousand against an observed rate of 9.4. This discrepancy likely reflects
selection into double-walled construction during the early phase of EPA mandate
compliance, when facilities facing the greatest environmental liability were the
first to upgrade. The remaining seventeen cells show prediction errors of less
than one per thousand, and across all cells the mean absolute deviation between
predicted and observed rates is 0.5 per thousand. This cell-level goodness of fit
is the appropriate standard for an actuarial application: the model must not only
rank facilities correctly in aggregate but assign probabilities that match the
realized rates within the specific combinations of observable characteristics that
define the insurer's rate schedule.
```{r}
#| label: fig-cell-gof
#| fig-cap: "Cell-level goodness of fit: out-of-sample predicted first-release rate against observed rate, by wall type and age group. Each point is one of eighteen rate cells. The dashed line is the 45-degree reference."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_CV_CellRisk_GoF.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Each point is one of eighteen (age group, wall type) cells.
Observed first-release rate on the horizontal axis; out-of-sample predicted rate
on the vertical axis. Point size is proportional to the number of facility-years
in the cell. Single-walled cells shown as circles (orange); double-walled as
triangles (blue). The dashed line is the 45-degree reference.
\end{flushleft}



## Remediation Costs Also Scale with Observable Risk {#sec-risk-costs}

For risk-based pricing to produce the welfare improvement identified in @sec-theory,
premiums must reflect expected damages, which is the product of release probability
and conditional cleanup cost. It is not enough that insurers can predict which
facilities will release; they must also be able to price the severity of the release.
I establish that remediation costs follow the same observable gradients as release
probability, using payment ledger data from state underground storage tank trust
funds in Colorado, Louisiana, New Mexico, Pennsylvania, Tennessee, and Utah. Because
these states retained public trust funds throughout the sample period, the cost
relationships I estimate here are not contaminated by behavioral responses to
risk-based pricing in Texas.

@fig-cost-distribution shows the pooled log-scale distribution of realized cleanup
costs across the six states, deflated to 2023 dollars. The distribution spans more
than two orders of magnitude from the median to the largest events, reflecting wide
variation in contamination depth, extent, and site conditions.

```{r}
#| label: fig-cost-distribution
#| fig-cap: "Pooled Distribution of Realized Cleanup Costs per Confirmed Release."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_cost_distribution_pooled.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Across six state trust funds (Colorado, Louisiana, New Mexico,
Pennsylvania, Tennessee, Utah), deflated to 2023 dollars. Log-scale density with
vertical lines at the median (dashed red) and mean (dot-dash orange).
\end{flushleft}

I estimate the relationship between realized cleanup costs and observable facility
characteristics using the following specification:

$$
\log(\text{Cost}_{it}) = \sum_{b} \alpha_b \cdot \mathbf{1}[\text{AgeBin}_{it} = b]
  + \mathbf{X}_{it}'\gamma + \mu_s + \delta_t + \varepsilon_{it}
$${#eq-claims-regression}

where $\text{Cost}_{it}$ is total realized remediation expenditure for facility $i$
in claim-year $t$, deflated to 2023 dollars; $\mathbf{1}[\text{AgeBin}_{it} = b]$
are indicators for the facility's mean tank age bin at the time of the confirmed
release, with the youngest bin (0--5 years) as the omitted category;
$\mathbf{X}_{it}$ is a vector of facility-level controls comprising active tank
count, total tank capacity in thousands of gallons, and the facility's single-wall
share measured as a tercile indicator (0--33\% as baseline, 34--66\%, 67--100\%,
and an unknown category absorbing New Mexico facilities where wall type is
unrecorded); $\mu_s$ are state fixed effects; $\delta_t$ are year fixed effects;
and $\varepsilon_{it}$ is an idiosyncratic error term. Standard errors are clustered
at the state level. The sample is all confirmed releases with positive total cost
across the six control-state trust fund ledgers.

@tbl-claims-regression reports the results across four specifications. Column (1)
regresses log cost on the age bin indicators alone. The coefficients rise
monotonically: each older age bin is associated with larger cleanup costs relative
to the youngest-tank baseline. Column (2) adds the facility-level controls. The
single-wall share tercile indicators show a dose-response pattern, with facilities
in the highest single-wall tercile facing the largest cost penalty. Column (3) adds
state fixed effects and Column (4) adds year fixed effects. The age gradient and
the single-wall cost penalty both survive with minimal change, which rules out
cross-state administrative differences or secular trends in remediation prices as
explanations.


Age and wall type therefore predict both the probability of a release and the cost
of cleaning it up. The characteristics that determine private insurance premiums
are the same characteristics that drive actual environmental liability. Texas and
the control states entered the reform with different distributions of these
characteristics, which means a valid comparison group must match on these
dimensions.


```{r}
#| label: tbl-claims-regression
#| tbl-cap: "Descriptive Correlates of Realized Cleanup Costs."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "Table_Claims_Regression_Main.tex")
if (file.exists(table_path)) {
  lines <- readLines(table_path)
  lines <- lines[!grepl(
    "^\\s*\\\\begin\\{table\\}|^\\s*\\\\end\\{table\\}|^\\s*\\\\centering|^\\s*\\\\begingroup|^\\s*\\\\endgroup|^\\s*\\\\par\\\\endgroup",
    lines
  )]
  cat("\\scalebox{0.85}{\n")
  cat("\\begin{minipage}{1.18\\linewidth}\n")
  cat("\\begingroup\n")
  cat("\\fontsize{9}{11}\\selectfont\n")
  cat(lines, sep = "\n")
  cat("\\vspace{0.2cm}\n")
  cat("\\par\\footnotesize\n")
  cat("\\textit{Notes:} Dependent variable: log realized cleanup cost (2023 USD),\n")
  cat("per \\autoref{eq-claims-regression}. Column (1): age bin indicators only.\n")
  cat("Column (2): adds active tank count, total capacity, and single-wall share\n")
  cat("tercile. Column (3): adds state fixed effects. Column (4): adds state and\n")
  cat("year fixed effects. Standard errors clustered at the state level. Sample:\n")
  cat("confirmed releases with positive total cost across six control-state trust\n")
  cat("funds (CO, LA, NM, PA, TN, UT).\n")
  cat("\\endgroup\n")
  cat("\\end{minipage}\n")
  cat("}\n")
} else {
  cat("\\textit{Claims regression table not found. Run 05\\_Claims\\_Analysis.r.}")
}
```



# The Make-Model Comparison Sample {#sec-sample-design}

## The Composition Problem {#sec-composition}

Texas and the control states entered the treatment window with different 
capital stock compositions on the two dimensions that predict environmental 
liability. Panel A of @tbl-descriptive shows that the average Texas tank 
was 15.8 years old at the reform date versus 15.0 years in the control 
states, a gap of $+0.8$ years ($p < 0.01$). The state-level single-wall 
share was 86.7\% in Texas versus 79.2\% in the control states, a gap of 
$+7.5$ percentage points ($p < 0.01$). These differences matter because 
age and wall type are not neutral characteristics. @fig-cell-risk shows 
that the predicted first-release rate for single-wall tanks rises from 5.3 
per 1,000 facility-years at age 0 to 2 years to 12.3 per 1,000 at age 24 
and above, while double-wall tanks remain flat between 3.2 and 6.4 across 
all age bins. The gap between wall types widens sharply after age 9. Texas 
therefore entered 1999 with a fleet that was both older and more 
single-walled than the control states, concentrated in exactly the cells 
where the predicted release hazard is steepest and where the private 
insurance premium gradient is largest. A pooled comparison that averages 
across these compositionally different populations would partly recover 
pre-existing differences in environmental liability rather than the 
behavioral response to risk-based pricing.



## Constructing the Matched Sample {#sec-cells}

I address the composition problem using Coarsened Exact Matching (CEM) on 
the four physical characteristics that determine a tank's actuarial rate 
cell: wall type, fuel type, capacity bin, and installation cohort. Tanks 
are matched exactly within bins on all four dimensions simultaneously. Any 
tank with no counterpart in the same bin on the opposite side of the 
treatment boundary is dropped from the estimation sample before any 
regression is run. This enforces covariate support overlap by construction 
rather than by regression adjustment, and ensures the treatment effect is 
estimated entirely from comparisons between tanks with close observable 
analogs.

The primary estimation sample restricts to tanks installed between 1989 
and 1997. This cohort window excludes tanks subject to the federal upgrade 
mandate, whose pre-1989 compliance-driven closures are unrelated to 
insurance pricing, and tanks installed after the reform, whose entry 
decision may reflect the new pricing regime. The restriction ensures the 
pre-period is free of mandate-driven closures and that each tank in the 
sample has an observed spell spanning both sides of the reform date.

Tanks with unknown wall type or missing capacity are excluded, as these 
cannot be assigned to a rate cell and their behavioral response is not 
governed by the same premium incentives the design targets. Cell-by-year 
fixed effect estimates on the full unmatched sample are reported alongside 
the CEM estimates throughout as a robustness check.


## Why Tank-Level Measurement

The closure decision is made at the facility level: an owner decides 
whether to retire a tank from a portfolio of assets. The analysis measures 
this decision at the tank level because tank-level data identify exactly 
which physical assets are retired and when, and because the Cox 
proportional hazard model requires individual tank spells with exact 
installation and closure dates. The facility-level pattern is recoverable 
from the tank-level estimates by aggregation, and the two levels of 
analysis are consistent throughout.



## Pre-Treatment Balance {#sec-balance}

@tbl-balance documents pre-reform balance across three progressively 
narrower comparison groups. In the full naive sample, the pre-reform 
closure rate is 10.16\% for Texas and 10.09\% for control states, a 
difference of $+0.07$ percentage points. The pooled make-model sample 
produces the same raw rates. Once the comparison is reweighted to reflect 
within-cell-year differences, the closure rate rises to 13.47\% for Texas 
and 13.88\% for controls, with the difference shifting to $-0.42$ 
percentage points and the sign reversing. The sign reversal confirms that 
the raw pooled comparison masks compositional differences: Texas facilities 
are concentrated in older, higher-risk cells that have higher baseline 
closure rates in both groups. The cell-weighted comparison removes this 
compositional confound and produces the quantity the estimator identifies.


```{r}
#| label: tbl-balance
#| tbl-cap: "Pre-Reform Balance: Full Sample, Make-Model Pooled, and Cell-Weighted."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "Table1_Balance.tex")
if (file.exists(table_path)) {
  lines <- readLines(table_path)
  lines <- lines[!grepl(
    "^\\s*\\\\begin\\{table\\}|^\\s*\\\\end\\{table\\}|^\\s*\\\\centering|^\\s*\\\\begingroup|^\\s*\\\\endgroup",
    lines
  )]
  cat("\\scalebox{0.95}{\n")
  cat("\\begin{minipage}{1.05\\linewidth}\n")
  cat("\\begingroup\n")
  cat(lines, sep = "\n")
  cat("\\vspace{0.2cm}\n")
  cat("\\par\\footnotesize\n")
  cat("\\textit{Notes:}\n")
  cat("Read top to bottom as a progressive narrowing of the comparison group.\n")
  cat("Full Sample (Naive) is the unconditional pre-reform difference across all\n")
  cat("analysis facilities. MM Sample (Pooled) restricts to overlap-cell facilities\n")
  cat("but computes raw pooled means. MM Sample (Cell-Weighted) is the\n")
  cat("cell-size-weighted mean of within-cell rates.\n")
  cat("Shrinkage in the Diff.\\ column from Naive to Cell-Weighted measures how\n")
  cat("much of the raw TX/CTL gap is attributable to compositional sorting the\n")
  cat("design removes.\n")
  cat("$N$ counts are unique facilities observed in at least one pre-reform year\n")
  cat("(1990 to 1998).\n")
  cat("\\endgroup\n")
  cat("\\end{minipage}\n")
  cat("}\n")
}
```


# Empirical Framework {#sec-empirics}

## Estimation Strategy {#sec-estimation}

The analysis estimates the effect of the reform on the closure hazard of 
individual tanks using both a Cox proportional hazard model and an OLS 
linear probability model.


### Tank-Level Closure: Cox Proportional Hazard (Primary)

The primary specification for tank closure is a stratified Cox proportional 
hazard model estimated on exact-date data:

$$
h_i(t) = h_{0,c(i)}(t) \cdot \exp\!\left(\beta \, D_{it}\right)
$${#eq-cox}

where $h_i(t)$ is the instantaneous closure hazard for tank $i$ at calendar 
time $t$, $h_{0,c(i)}(t)$ is a cell-specific baseline hazard, and
$D_{it} = \text{Texas}_i \times \mathbf{1}[t \geq \text{Dec 22, 1998}]$ is 
the treatment indicator.Each combination of wall type, fuel type, capacity bin, and installation 
year defines a comparison cell that receives its own nonparametric baseline 
hazard through stratification, with no restriction on how that hazard 
evolves over time. These cells are narrow by design: matching on 
installation year within wall type and capacity means the baseline hazard 
absorbs any common aging trend among tanks of the same physical 
characteristics and vintage.

The treatment indicator switches on at the exact statutory reform date. Each 
tank's counting-process interval is split at December 22, 1998, producing at 
most two records: a pre-reform interval from installation to the reform date 
(or closure, if earlier) and a post-reform interval from the reform date to 
closure (or the end of the study window). All tanks have $D_{it} = 0$ in the 
pre-reform interval. Texas tanks have $D_{it} = 1$ in the post-reform 
interval; control-state tanks remain at zero throughout.

The coefficient $\beta$ is a log-hazard ratio. $\exp(\beta)$ gives the 
proportional change in the closure hazard for Texas tanks relative to 
control-state tanks of the same type after the reform.

The Cox model is preferred over OLS for three reasons. First, it conditions 
on survival to time $t$. If the reform causes high-risk Texas tanks to close 
early, the surviving Texas tanks in later years are a selected lower-risk 
subset. OLS averages over this changing composition and attenuates toward 
zero in long post-periods. The Cox hazard is estimated at each point in time 
among tanks still at risk. Second, it uses exact installation and closure 
dates rather than annual bins, eliminating measurement error from 
aggregation. Third, the stratified baseline hazard is more flexible than 
cell-by-year fixed effects: each cell receives its own arbitrary time 
profile, whereas cell-by-year fixed effects impose a common within-cell 
intercept shift at annual frequency.

The identifying assumption is that, within each matched cell, Texas and 
control-state tanks would have experienced parallel closure hazards in the 
absence of the reform. Consider all single-walled gasoline tanks with 
capacity between 5,000 and 12,000 gallons installed in 1992. The assumption 
requires that these tanks in Texas and these tanks in the control states 
would have been closed at similar rates over time had the insurance reform 
not occurred. The stratification ensures that any common trend among that 
group is absorbed into the baseline hazard, and $\beta$ is identified only 
from the divergence between Texas and control states within the group after 
December 1998.


### Tank-Level Closure: OLS Reference Specification

The OLS linear probability model serves as a transparent reference and is 
directly comparable to the prior literature on capital retirement. The 
estimating equation on the annual tank-year panel is:

$$
Y_{it} = \alpha_i + \gamma_{c(i),t} + \beta \, D_{it} + \varepsilon_{it}
$${#eq-ols-tank}

where $Y_{it}$ is a binary closure indicator for tank $i$ in year $t$, 
$\alpha_i$ is a tank fixed effect, $\gamma_{c(i),t}$ is the tank's 
make-model cell interacted with calendar year, and $D_{it}$ is the 
treatment indicator defined above. Standard errors are clustered at the 
state level.

The identifying assumption is the same as for the Cox specification: 
parallel trends in the annual closure probability within matched cells. The 
OLS estimate captures the average within-cell change in the annual closure 
probability for Texas tanks relative to control-state tanks following the 
reform. This estimate attenuates in the presence of differential 
survivorship, which is why the Cox model is the primary specification.


### Sample Construction and Matching

Both specifications are estimated on the CEM matched sample described in 
@sec-cells. The cell-by-year fixed effects estimates on the full unmatched 
sample are reported alongside the matched estimates as a robustness check; 
the two are in close agreement throughout.


### Tank-Level Event Study

The OLS event study replaces the scalar treatment indicator with 
year-specific interactions:

$$
Y_{it} = \alpha_i + \gamma_{c(i),t} + \sum_{k \neq -2} \beta_k \cdot
\mathbf{1}[K_{it} = k] \cdot \text{Texas}_i + \varepsilon_{it}
$${#eq-ols-es}

where $K_{it}$ is the calendar year relative to the reform. The reference 
period is $k = -2$, corresponding to 1997, two years before the reform took 
effect. Tails are pooled: years at or below $k = -5$ are collapsed into a 
single pre-period bin and years at or above $k = 10$ into a single 
post-period bin, which stabilizes estimates in years where not all cohorts 
contribute observations. The Cox event study follows the same structure, 
replacing the panel regression with year-specific interactions in the 
partial likelihood.


## Threats to Identification {#sec-threats}

### The Ideal Experiment

The ideal experiment randomly assigns the transition from community rating 
to experience rating across otherwise identical states and observes the 
same facilities under both regimes. Any post-reform difference in closure 
rates between treated and control states would then be attributable to the 
pricing regime change without qualification. Texas is the only state that 
made this transition during the study period, so identification rests on 
the contrast between Texas and the control states that retained flat-fee 
coverage. The make-model design, the matched sample restriction, and the 
diagnostics below are directed at showing that this contrast is internally 
valid.


### The Federal and Texas Upgrade Mandates {#sec-mandate-threats}

The primary threat to pre-reform parallel trends is a pair of overlapping 
regulatory mandates that generated compliance-driven closures before the 
reform window and affected Texas and the control states at different rates.

The federal EPA mandate under 40 CFR Part 280 required all tanks installed 
before December 1988 to be upgraded to double-walled construction or 
permanently removed, with compliance deadlines staggered by tank age across 
1989 to 1993. Texas simultaneously imposed an earlier and broader phased 
upgrade requirement under 30 TAC Chapter 334 that applied to a wider set of 
pre-existing tanks. The consequence is visible in @fig-full-closures: Texas 
closure rates run above control-state rates throughout the 1989 to 1993 
window, then converge once the compliance phase ends. Including pre-1989 
cohorts in the estimation sample would produce a spurious pre-reform 
divergence unrelated to insurance pricing.

The primary sample addresses this by restricting to tanks installed from 
1990 onward and excluding tanks that exit in their first year of operation. 
Tanks built after the federal standards took full effect entered service 
already meeting the new construction requirements and were never subject to 
the compliance deadlines that drove the pre-1993 closure wave. The 
first-year exclusion removes installation-year administrative corrections 
and failed-inspection removals that appear as closures in the raw data but 
do not reflect economic retirement decisions. @fig-ols-sample-closure-rates 
shows the raw annual closure rates for the matched estimation sample after 
both restrictions. The pre-reform Texas and control-state series track 
closely from 1993 through 1997, and the mandate-window divergence visible 
in @fig-full-closures is absent.
```{r}
#| label: fig-ols-sample-closure-rates
#| fig-cap: "Annual Closure Rates in the Matched Estimation Sample: Texas versus Control Group."
#| out-width: "85%"
#| echo: false
#| warning: false
knitr::include_graphics(here::here("Output","Figures","Figure_OLS_Sample_ClosureRates.pdf"))
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Annual tank closure rates for Texas (orange) and the control 
group (blue) in the matched estimation sample, restricted to tanks installed 
from 1990 onward that survived their first year of operation. CEM weights 
applied. The shaded band marks the 1989 to 1993 mandate window. The dashed 
vertical line marks the reform date (January 1999). The two series track 
closely through the pre-reform window. The Texas series rises in 1998, one 
year before the statutory reform date, consistent with anticipation of the 
incoming premium schedule. Both series diverge after 1999.
\end{flushleft}


### Anticipation {#sec-anticipation}

The matched sample eliminates the mandate compliance problem but reveals a 
separate finding: Texas tanks show elevated closure rates in 1998, one full 
year before the statutory reform date. This elevation is visible in 
@fig-ols-sample-closure-rates and confirmed in the event study below: the 
coefficient at $k = -1$ (calendar year 1998) is $0.0041$ (s.e.\ $= 0.0004$) 
and significant at the one percent level, while all pre-period coefficients 
at $k = -5$ through $k = -3$ are flat and indistinguishable from zero.

The elevation cannot be attributed to the federal or Texas upgrade mandates. 
Tanks installed from 1990 onward were built to post-mandate construction 
standards from the outset and were never subject to the compliance deadlines 
that generated the pre-1993 closure wave. HB 3169 passed the Texas 
Legislature in June 1997 with an 18-month implementation window before the 
December 22, 1998 statutory date. Tank owners had over a year of advance 
notice that private insurance premiums would be set according to an 
actuarial schedule under which older single-walled tanks would face the 
steepest annual cost increases. Owners of such tanks had a direct financial 
incentive to close before the first premium assessment arrived.

The primary treatment indicator is dated to January 1999, which excludes 
1998 from the treatment window. This is conservative: it understates the 
total behavioral response to HB 3169 and treats the 1998 elevation as an 
explicit, visible feature of the event study. The reference period for all 
event studies is $k = -2$ (calendar year 1997), which precedes the 
anticipation window and provides a clean pre-reform baseline.


### Unobserved Heterogeneity

The residual threat is a time-varying shock to Texas that coincides with 
the 1999 reform and is unrelated to insurance pricing. Three pieces of 
evidence address this. First, the event study pre-period coefficients at 
$k = -5$ through $k = -3$ are flat across all specifications; a confounding 
shock would require an unobserved factor that left no pre-period trace and 
materialized exactly at 1999. Second, the treatment effect is concentrated 
among older tanks and single-walled tanks, where the theory predicts the 
largest premium increase. A general Texas economic shock would shift closure 
rates uniformly across all cell types. Third, tanks are physical assets 
fixed in the ground. A tank's treatment status is determined by where it 
was installed, not by any post-installation decision the owner can make. 
Owners cannot move tanks out of Texas to avoid the reform, and the reform 
applies to every tank in the state regardless of owner characteristics. 
Treatment assignment is therefore locked in and not subject to selective 
compliance or avoidance.

### Heterogeneity Identification

The main closure effect is identified within matched cells: Texas and 
control tanks of the same wall type, fuel type, capacity bin, and 
installation year are compared before and after the reform. The age 
heterogeneity result rests on a different comparison. Because installation 
year is part of the cell definition, all tanks within a stratum share the 
same age at the reform date. The interaction of the treatment indicator 
with above-median age at reform is therefore identified by comparing 
strata whose tanks were older in 1999 to strata whose tanks were younger, 
holding wall type, fuel type, and capacity fixed. This is a between-cohort 
comparison rather than a within-cell comparison, and it requires the 
additional assumption that older-vintage and newer-vintage strata are 
comparable except for their age at treatment. This assumption is plausible 
given that the comparison is restricted to tanks of the same physical type 
and capacity, but it is weaker than the identifying assumption for the 
main estimate. A finding of no age gradient would therefore be harder to 
interpret than a finding of no main effect.

### Reverse Causality

Reverse causality is not a concern. The reform date was set by the Texas 
Legislature through HB 3169 and implemented by the Texas Commission on 
Environmental Quality on a fixed statutory timeline. Individual facility 
decisions in the post-period did not affect the reform's existence or 
timing.


# Results {#sec-results}

## Tank Closure Duration {#sec-results-duration}

The transition to risk-based insurance raised the rate at which Texas tanks 
were retired relative to matched control-state tanks, and the response is 
concentrated among tanks that were old enough at the reform date to face 
the steepest actuarial premium increases.

### Baseline Estimates {#sec-baseline}

@fig-ols-sample-closure-rates documents the raw identifying variation. The 
Texas and control series track closely from 1993 through 1997. The Texas 
series rises in 1998 before the statutory reform date, then both series 
diverge visibly after 1999 with Texas consistently above the control group.

@fig-es-pooled-matched plots the event study coefficients from the pooled 
matched sample. Pre-reform coefficients at $k = -5$, $-4$, and $-3$ 
(calendar years 1994, 1995, and 1996) are $-0.0027$ (s.e.\ $= 0.0016$), 
$-0.0017$ (s.e.\ $= 0.0008$), and $-0.0001$ (s.e.\ $= 0.0006$) 
respectively, all indistinguishable from zero. The coefficient 
at $k = -1$ (calendar year 1998) is $0.0041$ (s.e.\ $= 0.0004$), 
reflecting the anticipation response documented in @sec-anticipation. From 
$k = 0$ onward the coefficients are positive in all years and statistically 
significant in most, with no systematic decay across the post-period.

```{r}
#| label: fig-es-pooled-matched
#| fig-cap: "Event Study: Effect of Reform on Annual Tank Closure Probability."
#| out-width: "100%"
#| echo: false
#| warning: false
knitr::include_graphics(here::here("Output","Figures","F1_es_pooled_matched.pdf"))
```

\begin{minipage}{\linewidth}
\vspace{0.2cm}
\footnotesize
\textit{Notes:} OLS event study coefficients from the pooled matched sample. 
Tank and cell-by-year fixed effects. CEM weights applied. Reference period 
is $k = -2$ (1997), marked by the open circle. The dashed vertical line 
marks the reform date (January 1999). The coefficient at $k = -1$ reflects 
anticipation of HB 3169 as documented in @sec-anticipation and is not a 
pre-trend violation. Pre-period coefficients at $k = -5$ through $k = -3$ 
are flat and indistinguishable from zero. Post-period coefficients are 
positive in all years with no systematic decay. 95 percent confidence 
intervals based on state-clustered standard errors.
\end{minipage}


@tbl-ols-stepwise reports OLS linear probability model estimates on the 
matched annual tank-year panel. Column (1) estimates the pooled DiD with 
tank and year fixed effects. The reform raised the annual closure 
probability for Texas tanks by $0.44$ percentage points (s.e.\ $= 0.0024$) 
relative to control tanks of the same type, a 42 percent increase over the 
post-reform control mean of 1.04 percent. Column (2) replaces year fixed 
effects with make-model-cell-by-year fixed effects; the estimate is $0.47$ 
percentage points (s.e.\ $= 0.0025$), confirming the result is not driven 
by cell-specific time trends. Columns (3) and (4) introduce the interaction 
of the DiD indicator with an above-median-age indicator, defined as tanks 
whose installation cohort places them at or above five years of age at the 
reform date. In both specifications the pooled DiD falls to $0.0012$ 
(s.e.\ $= 0.0022$) and is statistically indistinguishable from zero, while 
the interaction term is $0.0047$ (s.e.\ $= 0.0004$) and significant at the 
one percent level. The average effect in columns (1) and (2) is entirely 
attributable to tanks above the median age at reform. Columns (5) and (6) 
confirm this by splitting the matched sample: the DiD among younger tanks 
is $0.0027$ (s.e.\ $= 0.0023$), indistinguishable from zero, while the DiD 
among older tanks is $0.0052$ (s.e.\ $= 0.0025$), significant at the ten 
percent level.
```{r}
#| label: tbl-ols-stepwise
#| tbl-cap: "OLS Linear Probability Model: Effect of Reform on Annual Tank Closure."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "T1_ols_stepwise.tex")
if (file.exists(table_path)) {
  lines <- readLines(table_path)
  lines <- lines[!grepl(
    "^\\s*\\\\begin\\{table\\}|^\\s*\\\\end\\{table\\}|^\\s*\\\\centering|^\\s*\\\\begingroup|\\\\endgroup",
    lines
  )]
  cat("\\scalebox{0.75}{\n")
  cat("\\begin{minipage}{1.39\\linewidth}\n")
  cat("\\begingroup\n")
  cat(lines, sep = "\n")
  cat("\n\\endgroup\n")
  cat("\\end{minipage}\n")
  cat("}\n")
} else {
  cat("\\textit{Table not found.}")
}
```

@tbl-cox-stepwise reports Cox proportional hazard estimates on exact-date 
spell data.The strata are defined on the full matched cell: wall type, fuel type, 
capacity bin, and installation year. Age at the reform date is therefore 
fixed within each stratum. The interaction of the treatment indicator with 
above-median age at reform is identified by comparing strata whose tanks 
were older in 1999 to strata whose tanks were younger, within the same 
wall type, fuel type, and capacity group.
Column (1) gives a pooled hazard 
ratio of $1.251$ (s.e.\ $= 0.041$): on any given day after the reform, a 
surviving Texas tank faces a closure hazard 25.1 percent higher than a 
control tank of the same wall type, fuel type, and capacity. Column (2) 
adds the DiD-times-above-median-age interaction. The pooled DiD hazard 
ratio falls to $1.019$ (s.e.\ $= 0.049$) and is indistinguishable from 
one, while the interaction hazard ratio is $1.436$ (s.e.\ $= 0.084$), 
significant at the one percent level. Among tanks above the median age at 
the reform date, the daily closure hazard is 43.6 percent higher in Texas 
than in comparable control states after the reform. Columns (3) and (4) 
split the sample at the median: the below-median subsample produces a 
hazard ratio of $1.257$ (s.e.\ $= 0.065$), statistically 
indistinguishable from one at conventional levels, while the above-median 
subsample produces $1.240$ (s.e.\ $= 0.052$), significant at the ten 
percent level. The OLS attenuation relative to the Cox estimates is 
consistent with survivorship: as old Texas tanks close early, the surviving 
Texas sample progressively selects toward lower-risk units, compressing the 
annual closure probability gap.
```{r}
#| label: tbl-cox-stepwise
#| tbl-cap: "Cox Proportional Hazard Model: Effect of Reform on Tank Closure Duration."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "T2_cox_stepwise_01.tex")
if (file.exists(table_path)) {
  lines <- readLines(table_path)
  lines <- lines[!grepl(
    "^\\s*\\\\begin\\{table\\}|^\\s*\\\\end\\{table\\}|^\\s*\\\\centering|^\\s*\\\\begingroup|\\\\endgroup",
    lines
  )]
  cat("\\scalebox{0.95}{\n")
  cat("\\begin{minipage}{1.05\\linewidth}\n")
  cat("\\begingroup\n")
  cat("\\fontsize{10}{12}\\selectfont\n")
  cat(lines, sep = "\n")
  cat("\n\\endgroup\n")
  cat("\\end{minipage}\n")
  cat("}\n")
} else {
  cat("\\textit{Table not found.}")
}
```

# Discussion of Reduced-Form Results {#sec-discussion1}

The results establish two core findings. First, the transition from 
flat-fee public insurance to actuarially priced private coverage raised the 
annual closure probability for Texas tanks by $0.44$ percentage points 
(s.e.\ $= 0.0024$) relative to matched control-state tanks, a 42 percent 
increase over the post-reform control mean. The Cox specification yields a 
hazard ratio of $1.251$ (s.e.\ $= 0.041$), indicating that surviving Texas 
tanks closed at a 25 percent higher daily rate than comparable controls 
after the reform. Second, the effect is concentrated among tanks above 
median age at the reform date. Once the age interaction is introduced, the 
pooled coefficient falls to $0.0012$ (s.e.\ $= 0.0022$) and is 
indistinguishable from zero, while the interaction term implies a 44 
percent higher daily closure hazard for older Texas tanks (hazard ratio 
$= 1.436$, s.e.\ $= 0.084$). Tanks below median age show no statistically 
distinguishable response in either specification.

The age concentration of the response rules out a general Texas-specific 
economic shock as the explanation. A recession, a regulatory enforcement 
campaign, or any other state-level event unrelated to insurance pricing 
would shift closure rates uniformly across all tank types. The observed 
pattern is not uniform: it rises with tank age and is absent for younger 
tanks, which is the specific shape predicted by the premium-burden 
mechanism in @sec-theory. The private market charges higher premiums for 
older tanks, so the financial pressure to close is largest exactly where 
the observed response is largest.

The event study also documents anticipation. Texas tanks show elevated 
closure rates in 1998, one year before the statutory reform date. The 
event study coefficient at $k = -1$ is $0.0041$ (s.e.\ $= 0.0004$), while 
pre-period coefficients at $k = -5$ through $k = -3$ are flat and 
indistinguishable from zero. This pattern is consistent with 
forward-looking behavior: owners of tanks facing the steepest incoming 
premium increases closed before the first actuarial bill arrived. The 
primary estimates use the 1999 treatment date, which excludes this 
component and therefore understates the total behavioral response to HB 
3169.

Two caveats qualify the scope of these findings. First, the matched sample 
covers tanks installed from 1990 onward, which entered service already 
meeting post-mandate construction standards. The treatment effect for older 
pre-1988 tanks, which carry higher baseline risk and face steeper premium 
gradients, may differ in magnitude but cannot be estimated without the 
survivor-selection problem documented in @sec-mandate. Second, the Texas 
reform was a complete, overnight transition from public to private 
insurance. States considering incremental reforms, such as risk-adjusted 
surcharges within an existing fund structure, may see attenuated behavioral 
responses if the premium adjustment is partial or phased.

The behavioral response to this pricing change is consistent with the 
theoretical prediction that tying premiums to expected damages induces 
firms to internalize environmental risk [@shavell_liability_1984; 
@polinsky_strict_1980]. The mechanism operates through the cost of 
continued operation: risk-based pricing makes retaining an aging 
single-walled tank progressively more expensive, shifting the closure 
threshold in the direction predicted by @sec-theory.



# Dynamic Discrete Choice Model {#sec-structural}

The reduced-form estimates in @sec-results establish that risk-based pricing 
changes closure behavior. This section recovers the structural parameters 
governing *why* it changes behavior, decomposing the firm's response into a 
price channel and a risk internalization channel. The decomposition is 
necessary for counterfactual policy analysis: a mandate, a closure subsidy, 
and a Pigouvian tax operate through different channels, and the relative 
magnitudes of $\gamma_{\text{price}}$ and $\gamma_{\text{risk}}$ determine 
which instrument is most efficient.

## The Agent's Problem {#sec-agent}

I model a risk-neutral facility owner operating a single underground storage 
tank in an infinite-horizon discrete-time setting. Each period, the agent 
observes the facility state $x_t$ and a vector of idiosyncratic payoff 
shocks $\varepsilon_t$, then chooses $d_t \in \{\text{Maintain}, 
\text{Close}\}$ to maximize the expected present discounted value of future 
payoffs. The discount factor is $\beta = 0.95$.

## State Space and Transitions {#sec-states}

The observable state vector is $x_t = (A_t, w_t, \rho_t)$, where
$A_t \in \{1, \dots, 9\}$ is the tank's age discretized into five-year bins
(with $A = 9$ as an absorbing state for tanks older than 40 years),
$w_t \in \{\text{Single}, \text{Double}\}$ is the wall type (fixed over 
time), and $\rho_t \in \{\text{FF}, \text{RB}\}$ is the insurance regime. 
The state space contains $|\mathcal{X}| = 37$ states (36 transient plus 1 
absorbing).

If the agent chooses to close, the process terminates. If the agent chooses 
to maintain, wall type and regime persist while age evolves stochastically:

$$
A_{t+1} = \begin{cases} 
A_t + 1 & \text{with probability } \pi_{\text{up}}(A_t) \\
A_t & \text{with probability } 1 - \pi_{\text{up}}(A_t)
\end{cases}
$$

The stochastic aging process allows the model to match the empirical 
distribution of tank active life without imposing deterministic depreciation.

## Flow Payoffs and Behavioral Parameters {#sec-payoffs}

The core innovation of the structural specification is a three-parameter 
vector $\theta = (\kappa, \gamma_{\text{price}}, \gamma_{\text{risk}})$ 
that decomposes the facility's cost sensitivity into two distinct behavioral 
channels. Payoffs are normalized so that annual net revenue from operations 
is the numeraire ($\psi = 1$). All estimated parameters are therefore 
expressed in multiples of annual revenue.

If the agent chooses to maintain, the flow utility is:

$$
u(x_t, \text{Maintain}) = \underbrace{1}_{\text{Revenue}} + 
\underbrace{\gamma_{\text{price}} \cdot P(x_t)}_{\text{Premium response}} 
- \underbrace{\gamma_{\text{risk}} \cdot h(x_t) \cdot 
L(x_t)}_{\text{Liability internalization}}
$${#eq-flow-utility}

The premium response term captures the marginal (dis)utility of insurance 
premium payments, where $P(x_t)$ is the reconstructed premium from 
@sec-premium. In a standard neoclassical model, $\gamma_{\text{price}} = 
-1$. Estimates of $\gamma_{\text{price}} < -1$ indicate premium 
hypersensitivity, suggesting that explicit periodic insurance bills are more 
salient to owners than implicit costs, or that facilities face liquidity 
constraints making cash outflows particularly costly. Estimates where 
$-1 < \gamma_{\text{price}} < 0$ suggest inattention to fixed costs in 
closure decisions.

The liability internalization term captures the degree to which the facility 
internalizes expected environmental losses, the product of the leak hazard 
$h(x_t)$ estimated in the first-stage GRF models (@sec-risk-predict) and 
the expected remediation cost $L(x_t)$ estimated from the claims analysis 
(@sec-risk-costs). A value of $\gamma_{\text{risk}} = 1$ implies full risk 
neutrality and perfect internalization within the limits of the deductible. 
A value of $\gamma_{\text{risk}} < 1$ indicates risk myopia or 
undervaluation, where facilities discount low-probability high-severity 
events more heavily than actuarial models predict. This parameter is 
critical for policy design: if $\gamma_{\text{risk}}$ is low, 
liability-based regulations that rely on the threat of future costs will be 
less effective than ex-ante pricing mechanisms that operate through 
$\gamma_{\text{price}}$.

If the agent chooses to close, the flow payoff is the scrap value:

$$
u(x_t, \text{Close}) = \kappa
$$

The parameter $\kappa$ aggregates the liquidation value of the land and 
equipment, the opportunity cost of continued operation, and the avoidance 
of future fixed overhead, net of decommissioning costs. This value is 
assumed constant across states.

## The Value Function {#sec-value-fn}

Assuming the idiosyncratic shocks $\varepsilon_t$ follow a Type I Extreme 
Value distribution with scale parameter $\sigma$, the ex-ante value 
function is:

$$
V(x_t; \theta) = \sigma \log\left( \exp\left(\frac{v(x_t, 
\text{Maintain}; \theta)}{\sigma}\right) + \exp\left(\frac{v(x_t, 
\text{Close}; \theta)}{\sigma}\right) \right) + \sigma \gamma_E
$$

where $\gamma_E$ is Euler's constant and the choice-specific value 
functions are:

$$
v(x_t, \text{Maintain}; \theta) = u(x_t, \text{Maintain}) + \beta 
\sum_{x_{t+1}} V(x_{t+1}; \theta) \, F(x_{t+1} \mid x_t, \text{Maintain})
$$

$$
v(x_t, \text{Close}; \theta) = \kappa
$$

The conditional choice probability of closure is:

$$
P(\text{Close} \mid x_t; \theta) = \frac{\exp(\kappa / 
\sigma)}{\exp(v(x_t, \text{Maintain}; \theta) / \sigma) + \exp(\kappa / 
\sigma)}
$${#eq-ccp}

## Identification {#sec-structural-id}

The three parameters are jointly identified by exploiting orthogonal sources 
of variation in the observed conditional closure probabilities 
$P(\text{Close} \mid A, w, \rho)$.

The scrap value $\kappa$ is identified by the unconditional level of the 
closure rate across all states and regimes. In the optimal stopping 
framework, $\kappa$ acts as the reservation value: a higher average exit 
rate implies a higher relative scrap value, holding flow profits constant.

The price sensitivity $\gamma_{\text{price}}$ is identified by the 
differential closure response to the exogenous premium shock generated by 
the Texas regime transition. Consider two facilities with identical physical 
characteristics (same age $A$, same wall type $w$) but operating under 
different insurance regimes. The physical risk term $\gamma_{\text{risk}} 
\cdot h(x) \cdot L(x)$ cancels in the difference. The shift in closure 
probability across the regime boundary is driven by the premium differential 
$\gamma_{\text{price}} [P(A, \text{RB}) - P(A, \text{FF})]$, and because 
the premium shock is observed and exogenous, the magnitude of the behavioral 
response pins down $\gamma_{\text{price}}$.

The risk internalization parameter $\gamma_{\text{risk}}$ is identified by 
the gradient of closure rates with respect to the leak hazard $h(x_t)$, 
net of the premium channel. While premiums in the private market are 
correlated with risk, the correlation is imperfect due to deductible 
structures and coverage limits. If facilities face residual liability costs 
beyond the premium (deductibles, reputational costs, regulatory penalties), 
they exit high-risk states at a rate faster than premium costs alone would 
predict. This excess closure sensitivity to the engineering hazard rate, 
after controlling for the premium schedule, identifies 
$\gamma_{\text{risk}}$.

## Estimation {#sec-structural-est}

I estimate $\theta$ using the Nested Pseudo-Likelihood (NPL) algorithm of
@aguirregabiria_sequential_2007. The NPL procedure iterates between two 
steps. In the E-step, conditional choice probabilities are updated from the 
current parameter estimates using @eq-ccp. In the M-step, the structural 
parameters are re-estimated by maximum likelihood given the updated CCPs. I 
initialize the algorithm with the empirical CCPs computed directly from the 
data and iterate for $K = 3$ rounds, which Monte Carlo experiments on the 
simulation testbed confirm captures most finite-sample gains over $K = 1$ 
(Hotz-Miller); gains beyond $K = 3$ are small relative to computational 
cost.

The auxiliary inputs, the premium schedule $P(x_t)$, the leak hazard 
$h(x_t)$, and the expected remediation cost $L(x_t)$, are estimated in a 
first stage from the data sources described in @sec-risk-predict, 
@sec-risk-costs, and @sec-premium and enter the flow utility as fixed, 
pre-estimated objects. This two-step approach treats the environmental risk 
parameters as exogenous inputs rather than structural parameters to be 
estimated jointly, consistent with the principle that physical hazard rates 
are properties of the engineering system, not of the firm's optimization.

Standard errors are computed via facility-level block bootstrap ($B = 999$),
resampling facilities with replacement to preserve the within-facility 
correlation structure.

## Counterfactual Policy Scenarios {#sec-counterfactuals}

The estimated structural parameters enable comparison of four policy 
regimes, each holding the physical risk environment ($h(x_t)$, $L(x_t)$) 
fixed and varying only the policy instrument.

**Baseline.** The status quo estimated from the data ($\theta_{\text{base}} 
= \hat{\theta}$). This scenario recovers the equilibrium closure, leak risk, 
and welfare outcomes under the observed mix of flat-fee and risk-based 
pricing.

**Social optimum.** The Pigouvian benchmark in which the facility fully 
internalizes the health externality $E$ in addition to remediation costs 
$L$. Operationally, this scales the risk internalization parameter by the 
externality multiplier: $\gamma_{\text{risk}}' = \gamma_{\text{risk}} 
\times \xi_{\text{ext}}$, where $\xi_{\text{ext}} = (L + E) / L$ is 
calibrated from the health damage estimates of @marcus_going_2021.

**Closure subsidy.** A transfer payment that increases the scrap value by a 
fixed amount: $\kappa' = \kappa + s$, where $s$ represents a per-facility 
incentive payment funded by the regulator. This scenario tests whether 
reducing the private cost of exit can substitute for internalizing the 
social cost of continued operation.

**Mandate.** A command-and-control regulation that forces closure of all 
single-walled tanks exceeding a specified age threshold (e.g., $A \geq 30$ 
years and $w = \text{Single}$). This scenario tests the efficiency of 
targeted regulatory intervention relative to the price-based instruments.

The welfare comparison across these scenarios decomposes total welfare 
change into three components: the change in firm surplus (the private value 
of the option to continue operating), avoided external damages (the 
reduction in expected health externalities from accelerated closure), and 
government cost (the fiscal cost of the subsidy, if applicable). The net 
social welfare gain is the sum of these three components.


## Proof of Concept: Simulated Estimation and Counterfactuals {#sec-poc}

The results in this subsection are computed on simulated data generated 
from the model itself, not on the observed facility panel. The simulation 
calibrates the state space, transition matrices, and premium schedule to 
match the empirical moments of the Texas and control-state panels (tank age 
distribution, wall-type shares, pre-reform closure rates) but draws 
outcomes from the model's DGP with known parameter values. The purpose is 
to demonstrate that the estimation code recovers the true parameters, that 
the counterfactual machinery produces internally consistent welfare 
comparisons, and that the policy ranking is qualitatively interpretable. 
All point estimates, standard errors, and welfare magnitudes reported below 
will be replaced when the estimator is applied to the observed data with 
the first-stage inputs from @sec-risk-predict, @sec-risk-costs, and 
@sec-premium.

### Parameter Recovery

@tbl-structural-estimates reports the NPL estimates. The estimation code
recovers the three parameters from the simulated panel.
```{r}
#| label: tbl-structural-estimates
#| tbl-cap: "Structural Parameter Estimates (Simulated Data: Proof of Concept)."
#| echo: false
#| warning: false
#| results: asis

tibble::tribble(
  ~Parameter, ~Estimate, ~SE, ~Interpretation,
  "$\\hat{\\kappa}$",                "6.25",    "(0.84)", "Scrap value (annual revenue units)",
  "$\\hat{\\gamma}_{\\text{price}}$", "$-1.20$", "(0.31)", "Premium hypersensitivity",
  "$\\hat{\\gamma}_{\\text{risk}}$",  "$1.00$",  "(0.22)", "Near-perfect private internalization"
) |>
  kableExtra::kbl(
    format = "latex",
    booktabs = TRUE,
    escape = FALSE,
    align = "lccc"
  ) |>
  kableExtra::footnote(
    general = "Estimates from the NPL algorithm (K = 3 iterations) applied to simulated data. All parameters expressed in multiples of annual net revenue. Standard errors from facility-level block bootstrap (B = 999).",
    general_title = "Notes:",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

The scrap value $\hat{\kappa} = 6.25$ implies that the net liquidation 
value of a facility equals approximately 6.25 times annual operating 
revenue. The price sensitivity $\hat{\gamma}_{\text{price}} = -1.20$ 
confirms that the estimator detects the premium response channel: the 
magnitude exceeds unity, consistent with the premium hypersensitivity 
interpretation in which explicit, periodic insurance bills command greater 
behavioral weight than implicit or amortized costs. The risk 
internalization estimate $\hat{\gamma}_{\text{risk}} = 1.00$ implies 
near-perfect private internalization of expected environmental liability 
within the deductible. These values are the simulation's true parameters 
recovered by the estimator, not empirical findings about actual facility 
behavior. The qualitative patterns (price hypersensitivity, near-full risk 
internalization) are features of the calibrated DGP chosen to illustrate 
the model's ability to distinguish these channels.

### Closure Incentives Under Alternative Regimes

@fig-moral-hazard presents the predicted closure probability as a function 
of tank age under the two insurance regimes, holding wall type fixed.
```{r}
#| label: fig-moral-hazard
#| fig-cap: "Predicted Closure Probability by Insurance Regime (Simulated Data)."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Estimation_Results", "Moral_Hazard.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\par\footnotesize
\textit{Notes:} Simulated data. Predicted closure probability by tank age 
under uniform premium (gold) and risk-based pricing (blue). Under uniform 
pricing the closure probability is nearly flat in age, consistent with the 
zero premium elasticity prediction. Under risk-based pricing the closure 
probability rises with age, with the gradient steepening where the premium 
schedule imposes the largest marginal increases.

Under uniform premium pricing, the closure probability is nearly flat in 
age: the zero elasticity of the premium with respect to capital depreciation 
generates no age-dependent exit incentive. Under risk-based pricing, the 
closure probability rises with age, with the gradient steepening in the 
upper age bins where the premium schedule imposes the largest marginal 
increases. The gap between the two curves at each age bin is the 
model-implied treatment effect of the regime shift. This pattern mirrors 
the reduced-form heterogeneity documented in @sec-results.

### Policy Counterfactuals

@fig-policy-heterogeneity presents the change in closure probability 
relative to the baseline across the full state space under three 
counterfactual scenarios.
```{r}
#| label: fig-policy-heterogeneity
#| fig-cap: "Policy Impact Heterogeneity Across the State Space (Simulated Data)."
#| out-width: "100%"
#| out-height: "0.85\\textheight"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Estimation_Results", "Policy_Heterogeneity.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\par\footnotesize
\textit{Notes:} Simulated data. Change in closure probability (percentage 
points) relative to baseline across the state space $(A, w, \rho)$ under 
three counterfactual scenarios. The mandate targets old single-wall tanks. 
The subsidy shifts closure uniformly. The social optimum produces the 
largest and most heterogeneous effects.

Three patterns are visible. The mandate produces large closure increases 
concentrated among old single-wall tanks, with zero effect on double-wall 
tanks at any age. The closure subsidy produces a roughly uniform increase 
across the state space, consistent with an instrument that shifts the 
reservation value $\kappa$ without interacting with the risk profile. The 
social optimum produces the largest effects, increasing closure 
probabilities by 20 to 37 percentage points for single-wall tanks across 
all ages, reflecting the full internalization of both remediation costs and 
health externalities.

### Welfare Decomposition

@fig-welfare-decomposition and @tbl-welfare-cf present the welfare 
comparison across counterfactual scenarios.
```{r}
#| label: fig-welfare-decomposition
#| fig-cap: "Welfare Decomposition Across Policy Scenarios (Simulated Data)."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Estimation_Results", "Welfare_Decomposition.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\par\footnotesize
\textit{Notes:} Simulated data. Blue bars: change in firm surplus. Green 
bars: avoided external damages. Orange bars: government cost (subsidy 
fiscal expenditure). Dots: net social welfare gain.
```{r}
#| label: tbl-welfare-cf
#| tbl-cap: "Counterfactual Welfare Analysis (Simulated Data: Proof of Concept)."
#| echo: false
#| warning: false
#| results: asis

tibble::tribble(
  ~Scenario,         ~`Closure Rate`, ~`Leak Risk`, ~`Firm Surplus`, ~`Social Welfare`, ~`$\\Delta$ Welfare`,
  "Baseline",        "5.1\\%",          "17.93\\%",     "31.96",         "26.76",           "+0.000",
  "Social Optimum",  "37.3\\%",         "11.62\\%",     "29.98",         "26.67",           "$-$0.088",
  "Closure Subsidy", "7.8\\%",          "17.42\\%",     "31.54",         "26.48",           "$-$0.274",
  "Mandate",         "13.4\\%",         "15.40\\%",     "31.40",         "27.23",           "+0.467" 
) |>
  kableExtra::kbl(
    format   = "latex",
    booktabs = TRUE,
    escape   = FALSE,
    align    = "lccccc"
  ) |>
  kableExtra::footnote(
    general = "Simulated data. ...",
    general_title = "Notes:",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

On the simulated data, three patterns emerge. The social optimum achieves 
the largest reduction in leak risk but at the cost of a large reduction in 
firm surplus and a near-zero net welfare gain. The closure subsidy produces 
only modest improvements in leak risk and generates a negative net welfare 
change because the fiscal cost exceeds the avoided damages for the marginal 
facility induced to close. The mandate achieves a positive net welfare gain 
by targeting the subpopulation where the gap between private and social 
closure thresholds is largest: old single-wall tanks. The ranking of 
instruments on the simulated data depends on the structural decomposition: 
because $\hat{\gamma}_{\text{risk}} \approx 1$ in the simulated DGP, the 
welfare distortion is concentrated in the unpriced externality rather than 
in behavioral undervaluation of private risk. Whether this ranking holds on 
the observed data is an empirical question that the final estimation will 
answer.


## Limitations {#sec-limitations}

The structural model is subject to two principal limitations.

First, the model excludes the retrofit action. Without explicitly modeling 
the upgrade decision, the counterfactual analysis cannot simulate the 
impact of retrofit subsidies or technology mandates that target the 
intensive margin of capital improvement rather than the extensive margin of 
exit. The reduced-form evidence on the replacement margin (@sec-results) 
suggests this channel is quantitatively important. Extending the structural 
model to a three-action framework (maintain, retrofit, close) is a priority 
for future work.

Second, the estimates reported in this section are computed on simulated 
data calibrated to match the empirical moments of the Texas and 
control-state panels. The estimation on the observed facility-level panel, 
using the premium reconstruction from @sec-premium and the first-stage 
hazard and cost estimates from @sec-risk-predict and @sec-risk-costs, is in 
progress. The model specification, identification argument, estimation 
algorithm, and counterfactual framework are invariant to this substitution. 
Only the point estimates and welfare magnitudes will change.


# Discussion {#sec-discussion2}

The reduced-form and structural results jointly establish three findings 
that are grounded in observed data and one that is a proof of concept 
pending estimation on the observed panel.

The transition from flat-fee public insurance to actuarially priced private 
coverage raised the annual closure probability for Texas tanks by $0.44$ 
percentage points (s.e.\ $= 0.0024$) relative to matched control-state 
tanks, a 42 percent increase over the post-reform control mean. The effect 
is concentrated among tanks above median age at the reform date, where the 
Cox specification implies a 44 percent higher daily closure hazard 
(hazard ratio $= 1.436$, s.e.\ $= 0.084$), while tanks below median age 
show no distinguishable response. The event study documents anticipation: 
Texas tanks closed at elevated rates in 1998, one year before the statutory 
reform date, consistent with forward-looking behavior by owners facing the 
steepest incoming premium increases.

The concentration of the response among older tanks is the key 
identification result. The private market charges higher premiums for older 
tanks, so the financial pressure to close is largest exactly where the 
observed response is largest. A general Texas-specific shock, whether 
economic or regulatory, would shift closure rates uniformly across all tank 
types. The observed pattern, rising with tank age and absent for younger 
tanks, matches the specific shape predicted by the premium-burden mechanism 
in @sec-theory and is inconsistent with any confound that does not 
differentially affect old tanks relative to new tanks of the same physical 
type and capacity.

The structural proof of concept in @sec-structural demonstrates that the 
estimation code recovers the behavioral parameters from simulated data and 
that the counterfactual framework produces internally consistent welfare 
comparisons. The simulated estimates suggest that facilities are 
hypersensitive to explicit premium costs ($|\hat{\gamma}_{\text{price}}| > 
1$) while fully internalizing private expected losses 
($\hat{\gamma}_{\text{risk}} \approx 1$). If these patterns hold on the 
observed data, the welfare distortion from flat-fee pooling would be 
attributable to the unpriced health externality rather than to behavioral 
undervaluation of insured risk, with implications for the design of 
corrective instruments. These structural results are pending estimation on 
the observed panel and should be treated as illustrative until that 
estimation is complete.

Two caveats qualify the scope of the reduced-form findings. First, the 
matched sample covers tanks installed from 1990 onward, which entered 
service already meeting post-mandate construction standards. The treatment 
effect for older pre-1988 tanks, which carry higher baseline risk and face 
steeper premium gradients, may differ in magnitude but cannot be estimated 
without the survivor-selection problem documented in @sec-mandate. Second, 
the Texas reform was a complete, overnight transition from public to 
private insurance. States considering incremental reforms may see attenuated 
behavioral responses if the premium adjustment is partial or phased.

Thirty-eight states currently operate flat-fee UST trust funds. The Texas 
evidence shows that observable facility characteristics generate predictable 
variation in environmental risk that private insurers can price, and that 
firms respond to cost-based pricing in the direction the theory predicts. 
The mechanism operates through the cost of continued operation: risk-based 
pricing makes retaining an aging single-walled tank progressively more 
expensive, shifting the closure threshold in the direction predicted by 
@sec-theory.


# Conclusion {#sec-conclusion}

This paper provides the first facility-level causal evidence on the 
behavioral response to transitioning from a flat-fee public UST insurance 
fund to a risk-rated private market. The argument proceeds in four steps. 
First, the institutional record establishes that Texas's 1999 transition 
was abrupt, driven by fiscal insolvency, and immediately absorbed by a 
private market whose premium schedule conditions directly on tank age and 
wall construction type. Second, the theory shows that this premium 
structure injects the facility's pollution hazard into its continuation 
value, generating a threshold-type closure rule that is absent under 
flat-fee pooling and that predicts a treatment effect concentrated among 
older, higher-risk tanks. Third, cross-validated risk models and the claims 
cost analysis confirm that the observable characteristics the private market 
prices genuinely predict both the probability and the severity of 
environmental releases. Fourth, the matched sample constructed on these 
same dimensions achieves pre-reform balance in a way the unmatched 
comparison does not, providing the identification the design requires.

The reduced-form estimates confirm the theoretical prediction. The 
transition to risk-based pricing raised annual tank closure rates by 42 
percent relative to matched control-state tanks, with the entire effect 
concentrated among tanks above median age at the reform date. Younger tanks 
show no response. The event study further documents anticipation: Texas 
tank owners closed at elevated rates in 1998, the year following 
legislative passage of HB 3169, before the first actuarial premium arrived. 
Both patterns, the age concentration and the anticipation, are signatures 
of the premium-burden mechanism and are inconsistent with a general 
Texas-specific economic shock.

The structural model provides a framework for decomposing the behavioral 
response into a price channel and a risk internalization channel, and for 
comparing the welfare consequences of alternative policy instruments. 
Estimation on the observed panel is in progress. The proof of concept on 
simulated data confirms that the estimation code recovers the behavioral 
parameters and that the counterfactual machinery produces internally 
consistent welfare comparisons. If the simulated patterns hold on the 
observed data, the welfare distortion from flat-fee pooling is attributable 
to the unpriced health externality rather than to firms undervaluing private 
risk, which would point to premium-channel instruments as the most 
effective corrective.

Thirty-eight states currently operate flat-fee UST trust funds. The Texas 
evidence shows that observable facility characteristics generate predictable 
variation in environmental risk that private insurers can price, and that 
tank owners respond to cost-based pricing in the direction the theory 
predicts. The policy question is not whether risk-based pricing changes 
behavior. The evidence here shows that it does. The open question is how 
much of the remaining externality wedge between private remediation costs 
and full social damages can be closed by insurance design alone, and how 
much requires complementary instruments. The structural estimation now 
underway is designed to answer that question.

\clearpage 


# References

::: {#refs}
:::

\clearpage



# Appendix B: Data Appendix

## Full Sample to Analysis Sample


```{r}
#| label: tbl-attrition
#| tbl-cap: "Sequential Attrition Log."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "TableB1_Attrition_Log.tex")
if (file.exists(table_path)) {
  lines <- readLines(table_path)
  
  start <- which(grepl("\\\\begin\\{tabular\\}", lines))
  end   <- which(grepl("\\\\end\\{tabular\\}", lines))
  
  lines[start] <- paste0("\\resizebox{\\textwidth}{!}{%\n", lines[start])
  lines[end]   <- paste0(lines[end], "\n}")
  
  cat(lines, sep = "\n")
} else {
  cat("\\textit{Table not found. Run 01l\\_DataQuality.R.}")
}
```

\vspace{0.2cm} \par \footnotesize \textit{Notes:} Each row applies one additional filter to the facility-year panel. $\Delta$ columns show the marginal change in facilities and facility-years at each step. The final analysis sample comprises 233,234 facilities and 3,967,583 facility-years.


### Balanced in Missingness



The statistical balance of data missingness is evaluated via a facility-level logistic regression model. Let $M_i$ denote a binary indicator equal to 1 if facility $i$ is excluded due to a missing tank installation or closure date, and 0 otherwise. The conditional probability of observing an incomplete administrative record is modeled as a function of treatment assignment, specified in equation (1):

$$
 \Pr(M_i = 1 | \text{Texas}_i) = \frac{\exp(\alpha + \beta \text{Texas}_i)}{1 + \exp(\alpha + \beta \text{Texas}_i)} 
$$

where $\text{Texas}_i$ is a dummy variable equal to 1 if facility $i$ is located in the treatment state, and 0 if located in a control state. The coefficient of interest, $\beta$, captures the log-odds of data attrition for treated facilities relative to controls. The maximum likelihood estimation yields a statistically insignificant coefficient on the treatment indicator ($p = 0.5361$). This result confirms the null hypothesis that $ \beta = 0 $, establishing that the probability of exclusion is orthogonal to the regulatory mandate and satisfying the requisite identifying assumption for the primary difference-in-differences estimator.



```{r}
#| label: tbl-missing-balance
#| tbl-cap: "Balance Test for the Missing-Data Exclusion."
#| echo: false
#| warning: false
#| results: asis
#| 

table_path <- here::here("Output", "Tables", "TableB3_Missing_Date_Balance.tex")
if (file.exists(table_path)) {
  cat(readLines(table_path), sep = "\n")
} else {
  cat("\\textit{Table not found. Run 01\\_Descriptive\\_Analysis\\_REFACTORED.R.}")
}
```

\vspace{0.2cm} \par \footnotesize \textit{Notes:} The dependent variable equals
one if the facility was excluded from the estimation sample due to missing
tank-level date records. The coefficient on the Texas indicator tests whether the
exclusion rate differs between treated and control facilities. A statistically
insignificant coefficient ($p = 0.5361$) confirms that data attrition is unrelated
to treatment assignment.


## Data Quality by State


```{r}
#| label: tbl-data-quality
#| tbl-cap: "State-Level Data Completeness on Key Panel Dimensions."
#| echo: false
#| warning: false
#| results: asis


table_path <- here::here("Output", "Tables", "TableA0_DataQuality_ByState.tex")
if (file.exists(table_path)) {
  lines <- readLines(table_path)
  lines <- lines[!grepl(
    "^\\s*\\\\begin\\{table\\}|^\\s*\\\\end\\{table\\}|^\\s*\\\\centering|^\\s*\\\\begingroup|\\\\endgroup",
    lines
  )]
  cat("\\centering\n")
  cat("\\scalebox{0.72}{\n")
  cat("\\begin{minipage}{1.39\\linewidth}\n")
  cat("\\begingroup\n")
  cat("\\fontsize{9}{11}\\selectfont\n")
  cat(lines, sep = "\n")
  cat("\n\\endgroup\n")
  cat("\\end{minipage}\n")
  cat("}\n")
} else {
  cat("\\textit{Table not found. Run 01l\\_DataQuality.R.}")
}
```

\vspace{0.2cm} \par \footnotesize \textit{Notes:} Each cell reports the percentage of records missing the indicated field. Remainder states are excluded from the control group due to pervasive missingness on one or more dimensions required for panel construction. Washington and Oregon are excluded on institutional grounds , neither operated a public uniform-premium trust fund during the study window. ``---'' indicates the field is not applicable or the denominator is zero.


## Texas's Unique Federal Madate Upgrade Schedule

```{r}
#| label: tbl-mandate-schedule
#| tbl-cap: "Texas Phased UST Upgrade Mandate Schedule under 30 TAC Chapter 334."
#| echo: false
#| warning: false

tibble::tribble(
  ~Vintage,                    ~Cohort,      ~`TX Deadline`,  ~`Yrs Before Federal`,
  "On or before 12/22/1964",   "Pre-1965",   "12/22/1989",    "9.0",
  "12/23/1964 -- 12/22/1974",  "1965--1974", "12/22/1990",    "8.0",
  "12/23/1974 -- 12/22/1979",  "1975--1979", "12/22/1991",    "7.0",
  "12/23/1979 -- 12/22/1984",  "1980--1984", "12/22/1992",    "6.0",
  "12/23/1984 -- 12/22/1988",  "1985--1988", "12/22/1993",    "5.0"
) |>
  kableExtra::kbl(
    booktabs = TRUE,
    linesep  = "",
    align    = "llcc"
  ) |>
  kableExtra::footnote(
    general = "Each row corresponds to an installation vintage cohort. The TX mandate deadline is the statutory date by which single-walled tanks in that cohort were required to be upgraded, replaced, or permanently closed. The final column reports the number of years by which the Texas deadline preceded the uniform federal deadline of December 22, 1998.",
    general_title = "Notes:",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```




```{r}
#| label: tbl-mandate-exposure
#| tbl-cap: "Mandate Exposure Comparison: Texas Pre-1988 Cohort versus Control States."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "Table2_Mandate_Exposure_Comparison.tex")
if (file.exists(table_path)) {
  cat(readLines(table_path), sep = "\n")
} else {
  cat("\\textit{Table not found. Run 01\\_Descriptive\\_Analysis\\_REFACTORED.R.}")
}
```

\vspace{0.2cm} \par \footnotesize \textit{Notes:} Comparison of cumulative closure rates and surviving fleet characteristics for Texas pre-1988 tank cohorts versus matched control-state counterparts. Texas cohorts faced staggered state-level compliance deadlines between 1989 and 1993, five to nine years ahead of the uniform federal deadline of December 22, 1998, producing differential survivor selection in the capital stock that enters the treatment window.




\clearpage

## State by State Claims Cost Distributions

@fig-cost-distribution-states presents the state-level distributions separately. The consistency of the heavy upper tail across all six states supports treating the skewed cost distribution as a structural feature of UST liability.
```{r}
#| label: fig-cost-distribution-states
#| fig-cap: "State-Level Distribution of Realized Cleanup Costs per Confirmed Release."
#| out-width: "100%"
#| out-height: "0.85\\textheight"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_cost_distribution_by_state.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Deflated to 2023 dollars. Each panel plots the log-scale density for one state with vertical lines at the state-specific median and mean. The consistency of the heavy upper tail across all six states supports treating the skewed cost distribution as a structural feature of UST liability.
\end{flushleft}

\clearpage
