---
title: "Internalizing Environmental Risk: Insurance Design and Firm Behavior in Hazardous Industries"
author:
  - name: Kaleb K. Javier
    email: kalebkja@berkeley.edu
    affiliations:
      - name: University of California, Berkeley
        department: Agricultural and Resource Economics
        city: Berkeley
        state: California
        url: https://www.kalebkjavier.com/
    thanks: "I thank [Insert acknowledgments and funding notes here]."
date: "`r format(Sys.Date(), '%B %d, %Y')`"
abstract: |
  Public trust funds in thirty-eight U.S. states insure underground storage tank (UST) leaks at a uniform flat fee, even though leak risk varies widely with observable traits such as tank age and wall construction. This paper quantifies how replacing flat-fee pooling with actuarially priced coverage changes firm behavior and environmental outcomes. I exploit Texas's 1999 closure of its trust fund, which forced all owners into a private market that sets premiums according to facility risk. A facility-year panel covering 1990--2021 links EPA inventories with administrative microdata for Texas and eighteen control states that retained public funds. I construct a make-model comparison sample that matches Texas facilities to control-state counterparts on the observable risk characteristics, tank age bin and wall construction type, that private insurers price, ensuring the comparison group is balanced on the dimensions along which the treatment effect is theoretically heterogeneous. Difference-in-differences and event-study estimates show that risk-rated premiums increased annual closure rates by two to three percentage points, shifted closures from market exit to on-site replacement with safer equipment, and reduced the probability of a reported leak by about one percentage point, nearly the entire pre-policy baseline for the high-risk single-wall cohort. The treatment effect is concentrated among older and single-walled facilities, consistent with the premium-gradient mechanism, and is near zero for young double-walled tanks. The findings provide the first causal micro-evidence that cost-based environmental liability insurance can internalize pollution risk more effectively than uniform public pooling, offering clear guidance for states that are reconsidering the structure of UST insurance programs.
format:
  pdf: 
    link-citations: true
    fontsize: 12pt
    fig-pos: "H"
bibliography: UST_lit.bib
fig-cap-location: top
geometry: margin=1in
header-includes:
  - \usepackage{caption}
  - \usepackage{subcaption}
  - \usepackage{amsmath}
  - \usepackage{float}
  - \usepackage{tablefootnote}
  - \usepackage{tikz}
  - \usepackage{pdflscape}
  - \usepackage{etoolbox}
  - |
    \preto\section{\clearpage}
  - |
    \makeatletter
    \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
    \def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
    \makeatother
    \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
---
<!-- 
abstract: |
  [BLANK] 
load packages
  
-->


```{r}
#| label: setup
#| include: false
#| message: false
#| warning: false

#| label: setup
#| include: false
# 1. Load custom environment variables (RENV_PATHS_*)
.libPaths(c("C:/Users/kaleb/Documents/ust_ins_move_to_github/staging_library", .libPaths()))

# 2. Load required packages
library(tidyverse)
# 2. FORCE the root identification 
# This looks for the directory containing the 'Reports' folder
here::i_am("Reports/Paper/01_LUST_Insurance_Draft.qmd")

# 3. DEBUG: Print the detected root to your render console
# If this doesn't say ".../ust_ins_move_to_github", here is the problem.
message("CHECKPOINT: Project root is: ", here::here())
# 2. Load all required packages 
library(here)
library(tidyverse)
library(kableExtra)
# Add any other packages your script requires below
```


\clearpage

<!-- https://www.twdb.texas.gov/mapping/gisdata.asp - gis dat -->

# Introduction

Underground Storage Tanks (UST) holding motor fuels and other hazardous liquids sit beneath about half a million U.S. facilities, and leaks from these tanks are the leading documented source of groundwater contamination @epa_underground_2024. Cleanup is costly; median remediation expenses exceed $100 000 per site @marcus_going_2021, and the probability of a release rises sharply with observable characteristics such as tank age and whether the walls are single or double. Despite these predictable differences, thirty-eight states finance remediation through public trust funds that charge every facility the same premium, diluting incentives to invest in safer technology and crowding out private insurers @astswmo_archive.

Economic theory predicts that tying premiums to expected damages should induce firms to internalize environmental risk [@shavell_liability_1984; @polinsky_strict_1980], yet the relevant margin here is not liability per se but the rating regime under which liability is financed. Uniform-premium public funds operate as community-rated pools: every covered facility pays the same fee regardless of its observable risk characteristics, replicating the cross-subsidization and incentive dilution that @Pauly1974 identified in flat-rate insurance and that @RothschildStiglitz1976 showed prevents competitive separating equilibria. Under community rating, two distortions operate simultaneously. Adverse selection pushes high-risk facilities toward the subsidized pool and crowds private risk-rated alternatives out of the market. Moral hazard suppresses prevention investment because the marginal high-risk facility bears no incremental premium from operating leak-prone equipment. Empirical tests of this mechanism for USTs are scarce. Michigan's mid-1990s shift toward private, risk-rated coverage suggested large gains but relied on aggregate counts that could not isolate facility-level responses [@yin_can_2011; @yin_risk-based_2011]. Whether actuarial pricing disciplines high-risk owners without simply driving them out of the market therefore remains unresolved.


Texas provides a clean natural experiment. In 1999 the state closed its public fund overnight and required all owners to purchase private coverage whose premiums vary with observable risk. I assemble a new facility-year panel covering 1990 to 2021 that links EPA inventories with administrative microdata from the Texas Commission on Environmental Quality and comparable data from eighteen control states that retained trust funds. The identification strategy rests on a make-model comparison sample that matches Texas facilities to control-state counterparts on the observable risk characteristics, tank age bin and wall construction type, that the private market prices. Because the treatment effect is theoretically heterogeneous along these pricing dimensions (Section 3, Proposition 3), comparability on them is required for valid causal inference; Section 4 shows that matching on make and model is what earns parallel pre-trends.


[REWIRTE RESULTS ]
<!-- Three main results emerge. First, risk-based pricing raised annual closure probabilities by roughly two to three percentage points and shifted closures from outright exit to on-site replacement with safer equipment. Second, the probability of a reported leak fell by about one percentage point, nearly the entire pre-policy baseline for the high-risk single-wall group. Third, the effect on leak rates is concentrated among facilities that were already high-risk before the policy change, consistent with cost-based pricing effectively targeting riskier firms. By providing the first causal micro-evidence that actuarially priced environmental liability insurance changes firm behavior in hazardous industries, the paper speaks directly to current policy debates over the future of state UST funds.  -->

**Related Literature**



A well-established body of theoretical work suggests that strict liability rules can be an effective policy instrument for environmental risks, particularly when paired with insurance contracts that align a firm's private incentives with social costs [@shavell_liability_1984; @shavell_model_1984; @shavell_liability_1982; @polinsky_strict_1980; @boyd_retroactive_1997; @shavell_chapter_2007]. A crucial condition for this alignment is the insurer's ability to accurately price the risk of environmental harm. This paper provides the first empirical test of this principle in the context of environmental liability, offering causal estimates of how firms respond to a transition from flat-fee to more efficient, cost-based insurance contracts.

The challenge of pricing environmental risk connects this study to the economics of insurance rating regimes. The pooling equilibrium maintained by uniform-premium state funds is the environmental analog of community rating: premiums are invariant to observable risk, which creates cross-subsidies that suppress precautionary investment and crowd out risk-differentiated private alternatives [@RothschildStiglitz1976; @Pauly1974; @CutlerZeckhauser2000]. Theory predicts that a shift to experience rating generates two behavioral responses. On the selection margin, high-risk facilities face a cost increase and may exit or upgrade. On the moral hazard margin, facilities that remain now face a direct financial return to risk reduction, inducing investment in safer technology. Identifying these margins separately requires within-facility panel variation that aggregate data cannot provide.
The empirical literature on rating regimes confirms both mechanisms, though rarely in the same market. In health insurance, risk adjustment reforms that moved Medicare Advantage toward experience rating reduced favorable selection but generated gaming responses among plans [@brown_how_2014; @mcwilliams_new_2012]. In credit markets, moving from flat-rate to risk-differentiated pricing improves screening and changes borrower composition [@einav_contract_2012; @einav_impact_2013; @nelson_private_2025], with efficiency gains depending on whether information asymmetries favor the insurer or the insured [@FinkelsteinMcGarry2006; @liberman_equilibrium_2018; @einav_io_2021]. In the environmental insurance context, @yin_risk-based_2011 find that switching from community-rated state funds to private coverage in Michigan reduced leak incidents by over twenty percent. The Texas reform studied here provides a sharper identification environment and facility-level data that allow the selection and moral hazard margins to be separated.

This paper also contributes to the literature on U.S. UST policy. While much of the existing work focuses on estimating the marginal damages of pollution from leaks [@marcus_going_2021; @zabel_hedonic_2012; @guignet_sell_2014; @guignet_impacts_2018; @guignet_contamination_2018; @walsh_contaminated_2017; @guignet_what_2013], this study is complementary, examining the design of management policies intended to prevent those damages.

This research is most closely related to studies of a similar insurance market reform in Michigan by @yin_can_2011, @yin_risk-based_2011, and @yin_environmental_2007. That body of work finds that transitioning to risk-based insurance reduced the number of small firms and reported leaks. However, those analyses relied on aggregate or cross-sectional data, which cannot control for unobserved facility-level heterogeneity and may be susceptible to omitted variable bias. This paper advances on prior work by adopting an empirical strategy similar to other modern studies of environmental liability [@boomhower_drilling_2019]. By using a rich, facility-level panel dataset and a comparison sample matched on observable risk characteristics, I implement difference-in-differences models that control for unobserved facility characteristics and common time shocks, providing more credible causal estimates of how firms adjust their technology, replacement, and exit decisions in response to cost-based pricing.

# Institutional Background

## Underground Storage Tanks: Physical Characteristics, Regulatory Lock-in, and Environmental Risk

The U.S. Environmental Protection Agency defines a UST system as any tank and connected underground piping with at least 10% of its volume below the surface [@epa_underground_2024]. The construction technology of a UST system is the primary determinant of its environmental risk profile and the observable margin on which actuarial pricing differentiates. The critical physical distinction within this capital stock is between single-walled systems (bare steel or fiberglass shells with no interstitial containment layer) and double-walled systems (which incorporate a secondary wall and continuous interstitial monitoring space).

The construction choice that creates this risk profile is economically consequential and effectively irreversible. Single-walled construction is cheaper to install: a 1987 GAO survey priced a standard 10,000-gallon single-walled unit at approximately $23,000 versus $39,000 for a comparable double-walled unit, a 67% premium that contemporaneous manufacturer filings confirm has persisted [@gao1987; @csi2021; @xerxes2018]. The savings are locked in at installation. Once a single-walled tank is in the ground, the owner cannot upgrade the wall construction without removing and replacing the tank entirely, a capital expenditure comparable to installing new equipment from scratch. The practical irreversibility of this choice is confirmed directly in the administrative data: in Texas from 1998 to 2021, only 0.003% of active facilities ever retrofitted tank construction prior to terminal closure. Wall type is therefore not a choice variable that responds to price signals at the intensive margin. It is a fixed, observable, and verifiable physical attribute, precisely the conditions required for an actuarially feasible pricing margin and the reason the private insurance market in Texas conditions premiums directly on it.

The regulatory framework reinforces this physical lock-in with a legal one. Under 40 CFR Part 280, UST systems are subject to vintage-differentiated technical standards: tanks installed under the 1988 regulations are grandfathered into the compliance classification they earned at the federal deadline, provided they remain unmodified. Any construction activity that alters the wall structure or piping configuration of an existing system, including the partial upgrades that would otherwise be the natural intermediate response to a rising insurance premium, triggers reclassification of the entire facility under the most current technical standards in effect at the time of modification. The 2005 and subsequent EPA technical revisions substantially raised those standards [@us_epa_milestones_2021]. An owner who initiates a partial upgrade to reduce premium costs therefore faces the prospect of a facility-wide compliance overhaul whose cost dwarfs the incremental premium savings.

This is a direct instance of the mechanism @gruenspecht_differentiated_1982
identifies in the context of auto emissions and that
@stavins_vintage-differentiated_2005 formalizes as a general property of
vintage-differentiated regulation: exempting existing capital from new standards
inadvertently extends the operating life of the oldest and most hazardous stock
by making any modification prohibitively costly. The modification trigger in
40 CFR Part 280 operates exactly as the analogous New Source Review provision
does in the Clean Air Act, where @bushnell_enforcement_2012 show that the threat
of reclassification deterred capital investment at grandfathered coal plants;
@heutel_plant_2011 provides structural estimates of the same retardation effect
for boiler vintages. The practical effect here is to make the choice set binary:
operate the existing system without modification, or close it entirely.
As @sallee_retiring_2023 and @jacobsen_regulating_2023 argue in the context of
vehicle fleet decarbonization, when new-capital regulation cannot reach the
existing stock, retirement-focused instruments are the appropriate complement.
The UST setting illustrates precisely why: incremental investment in risk
reduction is not an option the regulatory environment makes available, which is
why the extensive margin of closure is the sole behavioral margin the empirical
analysis can and should target.


The environmental challenge posed by underground storage tanks is fundamentally
a problem of managing an existing stock, not regulating new flows into it.
@fig-sw-dw-stock documents the composition of the active fleet from 1985 through
2020. Single-walled tanks accounted for approximately 83% of the active capital
stock in the mid-1980s and remained the dominant technology at the federal
compliance deadline in December 1998, when they still constituted roughly 73% of
active tanks. By 2020, after two decades of post-deadline attrition, their share
had fallen only to approximately 63%. Double-walled construction, by contrast,
grew from near zero at the start of the study window to just under 20% of the
fleet by 2020, reflecting that new installations post-1988 are overwhelmingly
double-walled while the legacy stock turns over slowly. The figure makes plain
that the trajectory of environmental risk in this market is determined almost
entirely by the pace at which the incumbent single-walled stock exits, not by
the character of new entrants. The remediation liability sitting beneath the
surface today was installed decades ago and cannot be neutralized by tightening
standards on new entrants. It can only be reduced by accelerating the exit of
the single-walled incumbents. Because the hazardous technology is already in the
ground, the relevant policy question is not how to screen entrants but how to
price the continued operation of the existing fleet so that owners internalize
the expected cost of a release.

```{r}
#| label: fig-sw-dw-stock
#| fig-cap: "Aggregate Composition of Active Underground Storage Tank Fleet by Wall Type, 1989--2020."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_SW_DW_StockLevel_Pooled.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Proportional share of the total active capital stock categorized as single-walled (solid line, circles) and double-walled (dashed line, triangles) across the pooled full sample. The vertical dashed line marks the federal EPA upgrade deadline of December 22, 1998. The gradual post-deadline decline of the legacy single-walled fleet, mirrored by the corresponding accumulation of double-walled infrastructure, reflects the slow attrition of existing capital rather than immediate regulatory replacement. This confirms that stock management, rather than new-flow regulation, governs the long-term trajectory of environmental risk. The combined shares may evaluate to slightly less than 100% in a given year due to a marginal fraction of active tanks with unclassified or unknown structural metadata.
\end{flushleft}

## The Federal Compliance Mandate and Pre-Treatment Dynamics

Federal regulation of USTs originates in the 1984 Hazardous and Solid Waste
Amendments to the Resource Conservation and Recovery Act, which directed the EPA
to establish comprehensive technical and financial standards for tank systems. The
EPA finalized these standards in 40 CFR Part 280 in 1988, requiring all
non-compliant tanks to be upgraded, replaced, or closed by a uniform federal
compliance deadline of December 22, 1998 [@EPA2017USTCompendium;
@us_epa_milestones_2021]. @fig-full-closures makes the consequences visible:
annual facility-level closure rates spike sharply in the years immediately
preceding the deadline, peak in 1997 and 1998, and fall equally sharply once
the deadline passes. This spike is a regulatory compliance event. It contaminates
the pre-1988 capital stock in Texas and the control states simultaneously and
cannot be interpreted as a behavioral response to insurance pricing. Texas
compounded this federal problem with a state-level one, imposing a staggered
upgrade schedule for pre-1988 tanks that ran from 1989 through 1993 and had no
analogue in any control state. The two problems together render the pre-1988
installation cohorts unsuitable for the identification design. Section~\ref{sec-mandate}
documents the Texas staggered mandate in detail and shows the differential
pre-treatment dynamics empirically.



```{r}
#| label: fig-full-closures
#| fig-cap: "Full Sample Annual Closure Rates and the Federal Mandate Effect."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_01_Full_Sample_closure_rates.pdf")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Full sample closure rates for Texas and control-state facilities across the study window, illustrating the spike in closures driven by the federal EPA upgrade deadline of December 22, 1998, and the subsequent decline in annual exit rates as the surviving fleet aged.
\end{flushleft}



## Financial Responsibility and the State Trust Fund Regime

The 1988 Financial Responsibility Rule (40 CFR Part 280, Subpart H) mandates that every UST owner demonstrate the financial capacity to fund corrective action and satisfy third-party liability claims, with minimum required coverage of \$1 million per occurrence (40 CFR §280, Subpart H). In the late 1980s, the nascent private environmental insurance market was unable to supply this coverage at affordable cost to the population of small, single-site operators that constitute the majority of UST owners. To prevent widespread non-compliance, many states established public trust funds financed by annual tank registration fees or fuel tax revenues, functioning as a publicly administered insurer of last resort. A defining feature of these funds is their premium structure: annual fees are set on a uniform, per-tank basis, invariant to any observable facility risk characteristic such as wall type, tank age, or leak detection technology. States retaining public funds consistently report that these funds cover 90% to 100% of all active facilities, as documented by the Association of State and Territorial Solid Waste Management Officials [@astswmo_tank_surveys]. This near-complete market saturation is the predicted equilibrium outcome: because the subsidized uniform-premium contract is priced below any risk-differentiated private alternative for high-risk facilities, private insurers are structurally unable to compete for the portion of the market with above-average risk. The resulting uniform-premium pooling equilibrium is the institutional baseline against which Texas's transition is evaluated, and it is precisely the community rating equilibrium that the insurance economics literature predicts will simultaneously generate adverse selection into the pool and suppress prevention investment at the facility level [@RothschildStiglitz1976; @Pauly1974; @CutlerZeckhauser2000].



## Texas's Policy Transition

Texas's departure from the public fund model constitutes a sharp, administratively determined policy discontinuity whose timing was driven by fiscal insolvency rather than any observable change in the statewide distribution of facility-level risk. Texas established the Petroleum Storage Tank Remediation Fund (PSTRF) in 1989 under the Texas Water Code. By 1995, actuarial assessments confirmed that the fund's liabilities, driven by the accumulating backlog of legacy remediation claims, were structurally unsustainable. In response, the 74th Texas Legislature enacted House Bill 2587, which mandated the sunset of the PSTRF [@HB2587_analysis_1995]. The legislature set the fund closure cutoff at December 22, 1998, the same calendar date as the federal EPA upgrade deadline codified in 40 CFR §280.21. The alignment of these two dates is not coincidental: the legislature explicitly selected the federal mandate date as the administrative anchor for the PSTRF cutoff, ensuring the fund would cease accepting new claims at precisely the moment when the largest cohort of non-compliant tanks was required to be addressed [@TexWaterCode_26_3512]. Effective December 22, 1998, all Texas UST owners were required to secure private insurance or qualify for self-insurance by demonstrating sufficient financial assets.

## The Post-Transition Private Market in Texas

The private market that absorbed the Texas UST population after December 1998 prices observable facility risk directly, generating the cross-sectional and intertemporal premium variation that identifies firm behavioral responses. From 2007 to 2021, the period for which granular financial responsibility administrative data are available from the Texas Commission on Environmental Quality (TCEQ), an average of 89.8% of active retail UST facilities in Texas satisfied their financial responsibility obligation through commercial private insurance. The remaining 10.2% relied on self-insurance mechanisms, available exclusively to large, financially capitalized owners who can satisfy stringent asset tests.

@fig-fr-coverage documents the market transition directly. Financial responsibility coverage rates and the count of covered facilities confirm that adoption of private insurance was rapid and near-universal immediately following the December 1998 cutoff. The transition was not gradual: virtually the entire Texas UST population moved from public to private coverage within the first year of the reform.

```{r}
#| label: fig-fr-coverage
#| fig-cap: "Financial Responsibility Compliance in Texas, 2007--2021."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "FigureA_TX_FR_Coverage_Composition.pdf")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Left panel: share of active retail facilities satisfying the EPA financial responsibility requirement via private insurance versus self-insurance. Right panel: total count of covered facilities. The near-universal adoption of private coverage within the first policy year confirms the reform was absorbed rapidly across the facility population. Source: TCEQ financial responsibility administrative records.
\end{flushleft}


[DESCRIBE FR FIGURE ABOVE]

```{r}
#| label: fig-fr-hhi
#| fig-cap: "Market Concentration in the Texas UST Private Insurance Sector, 2007--2021."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_TX_FR_Top5_Dominance_HHI.pdf")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Left panel: market share of the top five underwriters by covered facility count. Right panel: Herfindahl--Hirschman Index (HHI) for the Texas UST private insurance market. Mid-Continent Casualty Company (NAIC 23418) operated as the dominant underwriter throughout the post-transition period. Source: TCEQ financial responsibility administrative records.
\end{flushleft}


[DESCRIBE THE FIGURE ABOVE HHI AND WHO TO LOOK AT. DESCIRBE WHAT A BAND IS IN A SPECIFIC YEAR AND WHAT A POINT IS ON THE GRAPH.]

<!-- For a single-wall tank in the top age bin, the annual premium can represent a meaningful share of the margin on a low-volume retail fuel facility. no evidence for this at the momen-->



## Private Insurance Premium Reconstruction


```{r}
#| label: fig-premium-level
#| fig-cap: "Reconstructed Annual Premium Levels for Texas UST Facilities, 2006--2021."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "FigureB_MidContinent_Pricing.pdf")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Left panel: median premium by year, with pre- and post-transition shading. Center panel: premium index (1995 = 100) for single-wall and double-wall tanks separately. Right panel: log premium against tank age, with fitted lines by wall type. Premiums are strictly increasing in age and strictly higher for single-wall construction at every age, consistent with the Mid-Continent rate filing structure. Source: SERFF rate filings.
\end{flushleft}


 [THIS WHOLE SECTION NEEDS TO STATE CLEALRY THE FEW BEHAVIRAL ASSUMPTIONS I MADE TO RECREATE THE PREMIUM AS STATED IN THE CODE ABOUT DEDUCTABLE COICE AND THEN HOW THE BANDS ARE CREATED USING THE 40% CREDITS]

Private insurers do not publicly disclose individual policy premiums. I reconstruct the premium schedule that Texas facilities faced in the post-transition period by building a bottom-up actuarial rating engine from the rate filings of Mid-Continent Casualty Company (NAIC 23418). I obtain rate manuals filed with the System for Electronic Rates and Forms Filing (SERFF) covering four regulatory periods: 2006--2014, 2014--2019, 2019--2021, and 2021 onward. These filings provide the complete pricing algorithm, including base rates, Increased Limit Factors (ILF), and all secondary risk loads applied to observable facility characteristics.

I estimate the monthly premium $P_{ijt}$ for tank $j$ at facility $i$ in month $t$ as:

$$
P_{ijt} = \text{Base}_{t} \times \text{ILF}_{it} \times \left(1 + \sum_{k} \lambda_{k,t} \mathbf{X}_{ijt}\right)
$${#eq-premium}

where $\text{Base}_{t}$ is the per-tank base rate active in filing period $t$, $\text{ILF}_{it}$ is the Increased Limit Factor adjusting for the elected coverage limit (defaulting to the statutory \$1 million per occurrence minimum where unspecified), and $\mathbf{X}_{ijt}$ is a vector of observable risk characteristics including tank age bin, wall construction type, piping type, and installed leak detection technology. The parameters $\lambda_{k,t}$ are the secondary risk loads, additive percentage credits or debits, assigned to each characteristic in filing period $t$.

Three implementation assumptions are required. First, I hold the base rate constant within each filing period at its filed value, consistent with the filing terms. Second, for risk characteristics unobserved in the administrative records (including aquifer proximity and flood zone classification), I set the corresponding risk load to zero, yielding a conservative estimate of the true premium. Third, the rate filings permit underwriters to apply schedule rating adjustments of $\pm$40\% at their discretion; I use the base manual premium as the price signal and treat discretionary adjustments as mean-zero errors across the population, an assumption evaluated against the distribution of observed financial responsibility filings in the TCEQ records.

Granular premium microdata are available beginning in 2006. The period from January 1999 through December 2005 represents a transition window in which the private market was operational but individual premium records are not observed.




## The Texas Phased Mandate: A Pre-Treatment Confound {#sec-mandate}

Texas's transition to private financial responsibility was preceded by a separate
regulatory regime that creates a pre-treatment confound for the oldest installation
cohorts. Under 30 TAC Chapter 334, the Texas Natural Resource Conservation
Commission assigned staggered upgrade deadlines to pre-1988 tanks by vintage
cohort, running from December 22, 1989 for the oldest tanks through December 22,
1993 for the 1985--1988 cohort (@tbl-mandate-schedule). These deadlines preceded
the uniform federal date by five to nine years depending on vintage. No control
state imposed an analogous staggered schedule; facilities in control states faced
only the federal deadline of December 22, 1998.

The consequence is survivor selection in the Texas pre-1988 stock. By the time the
insurance reform took effect in 1999, the pre-1988 Texas tanks still operating had
already survived a cohort-specific compliance decision made years earlier: their
operators had elected to invest in upgrades rather than close when confronted with
the rolling state mandate. This is not a random sample of the original pre-1988
population. It is the subset whose operators revealed a preference for continuing
operations under regulatory pressure, before any insurance price signal existed.
Control-state pre-1988 operators faced no such pre-screening. Their closure
decisions in the years around 1998 reflect the federal mandate alone, not a prior
self-selection event. Any naive comparison pooling pre- and post-1988 cohorts
therefore conflates two distinct mechanisms: mandate-driven survivor selection in
Texas and insurance-driven behavioral response, the object of interest.

@tbl-mandate-exposure documents this differential empirically. Cumulative closure
rates and surviving fleet characteristics diverge sharply between Texas pre-1988
cohorts and their control-state counterparts during 1989--1993, the window when
the staggered mandate was binding in Texas and silent in control states. No
corresponding divergence appears for post-1988 cohorts, which were never subject
to the staggered schedule and which enter the treatment window with comparable
pre-reform closure histories across Texas and the control group.

Both problems identified in this section and in Section~\ref{sec-mandate-intro}
are resolved by restricting the estimation sample to tanks installed between 1990
and 1997. This cohort never faced the staggered Texas mandate, was not subject to
the federal compliance deadline before 1998, and entered the treatment window with
a clean pre-reform closure history whose dynamics are comparable across Texas and
the control states. The restriction is documented in @tbl-attrition and the
parallel pre-trends it earns are confirmed in the event study of
Section~\ref{sec-results}.




## Control Group Definition and Institutional Restrictions

The potential control group consists of all states that maintained a public 
uniform-premium trust fund continuously through the study window and were never 
subject to a mandatory transition to private financial responsibility. Approximately 
thirty states satisfy this institutional criterion. After applying four data 
completeness requirements dictated by the panel construction and identification 
design, the usable control group contracts to nineteen states.

**Missing installation dates.** The panel backbone requires a confirmed installation 
date for every tank regardless of status so that the facility can be assigned to an 
age cohort and the make-model cell can be frozen at the reform date. Where 
installation dates are missing at high rates, facilities cannot be assigned to an 
age bin, are excluded from the make-model sample, and generate attrition that is 
mechanically correlated with vintage — precisely the dimension the identification 
design exploits.

**Missing closure dates for tanks recorded as closed.** The panel assigns a closure 
event to a facility-year only when a closed tank has a confirmed closure date; tanks 
flagged as closed but lacking a closure date are dropped in the filtering step 
described in Section~\ref{sec-panelbuild} and documented in @tbl-attrition. Where 
closure-date missingness is high, the denominator of the closure rate is correctly 
measured but the numerator is understated, biasing the outcome toward zero in the 
pre-treatment period and attenuating the DiD estimate.

**Missing wall-type classification.** The make-model comparison sample requires 
that each facility be classifiable as all-single-walled, all-double-walled, or 
mixed on the dimensions private insurers price. Wall-type is stored in the EPA 
tank-type field. Where this field is missing at high rates, the matching cell is 
undefined and the facility cannot contribute to the make-model sample regardless 
of its institutional eligibility.

**Missing LUST report dates.** The leak outcome requires a confirmed report date 
for each leaking underground storage tank release, used to assign the release to a 
facility-year and to classify it under the primary 0--60 day window. States where 
report-date missingness is total are excluded from the leak analysis sample; states 
with partial missingness are retained for the closure analysis but excluded from 
the leak specifications.

@tbl-data-quality documents missingness rates across all candidate states on each 
of these four dimensions. The twenty-one states that relied solely on the EPA 
harmonized database without a custom state registry fail one or more of these 
requirements at rates that make the panel unusable. Closure dates are missing for 
100 percent of closed tanks in New York, South Carolina, Missouri, Mississippi, 
Nevada, Nebraska, Georgia, New Hampshire, Rhode Island, Alaska, Utah, Indiana, 
West Virginia, Wyoming, and Hawaii, and for 80 percent in South Dakota. 
Installation dates are missing for 100 percent of tanks in South Carolina, 
Mississippi, and Vermont, and for more than 30 percent in New York, Missouri, 
Indiana, Illinois, and West Virginia. Delaware satisfies both date requirements 
but is excluded on a third ground: wall-type classification is missing for 100 
percent of its tank inventory, making make-model cell assignment impossible. Two 
further states — Washington and Oregon — are excluded on institutional rather than 
data-quality grounds: neither operated a public uniform-premium trust fund during 
the study window and so do not satisfy the control group eligibility criterion 
regardless of data quality.

The nineteen states that satisfy all four completeness thresholds and maintained 
public trust funds continuously through 2020 form the control group used in all 
baseline specifications. Appendix~\ref{app-states} lists the included states, 
their trust fund histories, and their tank-inventory sizes. The robustness analysis 
in Section~\ref{sec-robustness} shows that restricting to the nine highest-quality 
control states does not materially alter the estimates.


# Insurance Regimes and the Retirement of Risky Capital

The insurance regime governs whether a facility's pollution risk enters its cost function. This section formalizes that claim, derives the closure threshold under each regime, and produces the testable predictions that motivate both the make-model comparison sample and the empirical specifications in Section 6. [MAKE THIS A SECTOIN REFF NOT HARD CODED]

## Primitives

The facility owner's problem is a single-agent dynamic discrete choice model in the optimal stopping tradition (Rust, 1987). In an infinite-horizon setting, a competitive facility owner operates a single underground storage tank and each period decides whether to continue operating or permanently close. Let $\mathbf{x} \in \mathcal{X}$ denote the vector of facility characteristics that determine pollution risk: tank age, construction type, wall material, soil composition, and any other observable feature correlated with the probability of an environmental release. The instantaneous leak hazard is $h(\mathbf{x}) > 0$. For analytical tractability, I model $\mathbf{x}$ as the scalar tank age $a$, with $h'(a) > 0$ and $h''(a) \geq 0$. The qualitative results hold for any dimension of $\mathbf{x}$ along which the insurer conditions the premium. The Extension section below states this generalization explicitly.

Each leak event generates two categories of social harm. Remediation cost $L > 0$ is the direct cost of cleaning contaminated soil and groundwater. The operator bears this cost under strict liability, making it the natural object of an insurance contract. Health externality $E > 0$ represents diffuse health damages accruing to surrounding communities through groundwater contamination. Because contamination is frequently undetected for years after a release and attribution to a specific facility is administratively difficult, these damages generate no insurable third-party claim (@marcus_going_2021). The gap between $L$ and $L + E$ defines the limits of what any privately-priced insurance arrangement can achieve relative to the social optimum, and it motivates the welfare decomposition below.

The owner earns per-period net revenue $R > 0$ from operating and recovers scrap value $\kappa > 0$ upon permanent closure, representing the liquidity of the land net of decommissioning costs. The owner discounts at rate $\beta \in (0,1)$ and is risk-neutral.

## Insurance Regimes

The insurance regime determines whether and how the premium $C(\mathbf{x})$ varies with the facility's risk profile. I define three regimes that span the relevant policy space.

**Definition 1 (Uniform Premium, UP).** The facility pays a flat actuarially pooled premium each period, independent of any facility characteristic:

$$
C_{\text{UP}}(\mathbf{x}) = \bar{p}.
$$

No risk information is transmitted to the firm through the insurance contract under UP. In the scalar age case, the marginal cost of aging, the elasticity of the premium with respect to capital depreciation, is identically zero:

$$
\frac{\partial C_{\text{UP}}}{\partial a} = 0.
$$

The insurer may possess full information about the cross-sectional risk distribution but does not act on it in setting individual premiums. This is the pricing structure of state-administered petroleum storage tank assurance funds, including the Texas PST Fund prior to its 1999 sunset.

**Definition 2 (Risk-Based Pricing, RB).** The insurer prices actuarially, conditioning on all observable elements of $\mathbf{x}$. The premium equals expected claims scaled by a loading factor $\lambda \geq 1$:

$$
C_{\text{RB}}(a) = \lambda \cdot h(a) \cdot L.
$$

Since $h'(a) > 0$, the premium's elasticity with respect to capital depreciation is strictly positive:

$$
\frac{\partial C_{\text{RB}}}{\partial a} = \lambda h'(a) L > 0.
$$

Every dimension of $\mathbf{x}$ that enters $h(\mathbf{x})$ is mechanically injected into the firm's flow cost by the actuarial pricing rule. This is a definitional property of risk-based pricing, not a behavioral assumption about firm awareness or optimization.

**Definition 3 (Social Optimum, SOC).** The social planner internalizes both remediation costs and health externalities:

$$
C_{\text{SOC}}(a) = h(a) \cdot (L + E).
$$

The social optimum is unattainable under private insurance because $E$ generates no insurable claim. For the parametrically relevant condition $E > (\lambda - 1)L$, the cost ordering is:

$$
C_{\text{UP}} \leq C_{\text{RB}}(a) < C_{\text{SOC}}(a),
$$

where the left inequality becomes strict for $a$ large enough that $\lambda h(a) L > \bar{p}$.

## The Firm's Dynamic Problem

Each period, the owner observes age $a$ and chooses to maintain or permanently close. Under regime $j$, flow utility from maintaining is $u_j(a) = R - C_j(a)$, where $R$ is annual net revenue from operations net of all non-insurance costs and $C_j(a)$ is the insurance premium. I maintain the simplifying assumption that insurance contracts provide full ex-post coverage with no deductible, treating environmental risk solely as an ex-ante premium cost rather than an ex-post liability shock. This assumption isolates the cost-of-continuing channel and is consistent with standard provisions in the petroleum storage tank insurance market; the structural estimation in Section 8 relaxes it to incorporate partial risk internalization through deductibles.

The Bellman equation under regime $j$ is:

$$
V_j(a) = \max \left\{ \underbrace{R - C_j(a) + \beta \, \mathbb{E}[V_j(a') \mid a]}_{\text{Maintain}},\;\; \underbrace{\kappa}_{\text{Close}} \right\},
$$

where $\beta$ is the discount factor, the expectation is taken over the stochastic evolution of tank age, and $\kappa$ is the scrap value recovered upon exit. The pure continuation value, the present discounted value of flow utility under the assumption that the tank is never closed, is:

$$
\widetilde{V}_j(a) = \sum_{s=0}^{\infty} \beta^s \left[ R - C_j(a+s) \right].
$$

The optimal value function satisfies $V_j(a) = \max\{\widetilde{V}_j(a),\,\kappa\}$. The optimal closure threshold is:

$$
a_j^* = \min \left\{ a \geq 0 : \widetilde{V}_j(a) \leq \kappa \right\}.
$$

Under any regime with $C_j'(a) > 0$, the function $\widetilde{V}_j(a)$ is strictly decreasing, so the optimal policy is a threshold rule: close if and only if $a \geq a_j^*$.

## Main Results

**Proposition 1 (Risk Transmission and Value Slope).**
*Under UP, $\widetilde{V}_{\text{UP}}'(a) = 0$. Under RB, $\widetilde{V}_{\text{RB}}'(a) < 0$. Under SOC, $\widetilde{V}_{\text{SOC}}'(a) < \widetilde{V}_{\text{RB}}'(a) < 0$.*

*Proof.* Differentiating the continuation value with respect to age:

$$
\widetilde{V}_j'(a) = -\sum_{s=0}^{\infty} \beta^s C_j'(a+s).
$$

Under UP, $C_{\text{UP}}'(\cdot) = 0$ everywhere, so the sum is zero and the continuation value is flat in age: the physical degradation of the asset is real but generates no financial pressure on the firm. Under RB, $C_{\text{RB}}'(a) = \lambda h'(a) L > 0$, so every term in the sum is strictly negative and $\widetilde{V}_{\text{RB}}'(a) < 0$. Under SOC, $C_{\text{SOC}}'(a) = h'(a)(L+E) > \lambda h'(a) L = C_{\text{RB}}'(a)$, so the SOC slope is strictly more negative at every age. $\square$

Proposition 1 establishes the mapping from the pricing regime to the information environment of the firm. Under UP, aging is economically invisible: the zero elasticity of the premium with respect to capital depreciation means the continuation value contains no signal about the facility's pollution hazard, and the firm retains the asset until it is physically obsolete. Under RB, aging is economically visible by construction: the actuarial pricing rule translates the insurer's risk model into a per-period cost that rises with the facility's hazard profile, pivoting the continuation value downward for older tanks and accelerating the optimal exit age, regardless of whether the owner attends to or understands the actuarial logic.

**Proposition 2 (Regime Ordering of Closure Thresholds).**
*Under the condition $E > (\lambda-1)L$, the optimal closure ages satisfy $a_{\text{SOC}}^* < a_{\text{RB}}^* < a_{\text{UP}}^*$.*

*Proof.* By Proposition 1, $\widetilde{V}_{\text{SOC}}(a) < \widetilde{V}_{\text{RB}}(a) < \widetilde{V}_{\text{UP}}(a)$ for all $a > 0$. Since $a_j^*$ is defined as the first age at which $\widetilde{V}_j(a)$ crosses $\kappa$ from above, a lower value function crosses the threshold at a younger age. The ordering of value functions is preserved under discounted summation because $h(a)$ is common across regimes; only the coefficient multiplying $h(a)$ differs. $\square$

@fig-mechanism illustrates Proposition 2. The SOC curve declines at the fastest rate and crosses the scrap value line $\kappa$ at $a_{\text{SOC}}^*$. The RB curve declines at an intermediate rate and crosses $\kappa$ at $a_{\text{RB}}^* > a_{\text{SOC}}^*$, strictly reducing the optimal tank lifespan relative to the uniform pricing baseline. The UP curve is flat by Proposition 1 and does not cross $\kappa$ within the plotted range, reflecting the zero elasticity of the premium with respect to capital depreciation under pooled pricing.

![The Mechanism: Optimal Stopping under Alternative Insurance Regimes](Output/Figures/Paper/Fig_1_Mechanism.pdf){#fig-mechanism width=90%}

\vspace{0.2cm} 
\begin{flushleft} 
\footnotesize \textit{Notes:} This figure illustrates the optimal closure thresholds under the three insurance regimes derived in Proposition 2. The horizontal line represents the scrap value ($\kappa$). The Social Optimum (SOC) continuation value declines at the fastest rate, prompting the earliest closure at $a_{\text{SOC}}^*$. The Risk-Based (RB) pricing continuation value declines at an intermediate rate, yielding closure at $a_{\text{RB}}^*$. The Uniform Premium (UP) continuation value remains flat, reflecting the zero elasticity of the premium with respect to capital depreciation under pooled pricing.
\end{flushleft}


## Frictions Attenuating Behavioral Response

The model establishes the direction of the effect of moving from UP to RB pricing. It does not establish the magnitude, and it is silent on two frictions that could attenuate behavioral response even when the directional incentive is correctly signed.

The first friction is premium insufficiency. Even if $C_{\text{RB}}'(a) > 0$, the premium increment at the margin may be small relative to the fixed transaction costs of closure, which include demolition, regulatory notification, site remediation, and foregone operating revenue. If the present value of avoided future premiums is less than these fixed exit costs for the marginal tank, the regime shift generates no behavioral response despite correctly pricing risk.

The second friction is constraints on action. Even when $\widetilde{V}_{\text{RB}}(a) < \kappa$ and closure is optimal, the owner may be unable to act on that signal. Credit constraints may prevent financing upfront exit costs. Thin markets for contaminated properties may render the book value of $\kappa$ unrealizable. Present bias or inattention to the actuarial premium schedule could cause owners to underweight future premium increases.

Neither friction is resolvable from theory alone. The Texas natural experiment provides the empirical test: the 1999 sunset of the PST Fund was abrupt, mandatory, and economically large, providing the variation required to estimate whether the premium-burden mechanism is sufficient to produce measurable changes in closure behavior.

**Proposition 3 (Orthogonality Under UP; Correlation Under RB).**
*Let $\xi$ denote closure drivers unrelated to the facility's pollution hazard, land value appreciation, local fuel demand shocks, and owner liquidity or retirement decisions, satisfying $\xi \perp a$. Then: (i) under UP, $\mathrm{Cov}(\mathbf{1}[\mathrm{Close}],\, a) = 0$; (ii) under RB, $\mathrm{Cov}(\mathbf{1}[\mathrm{Close}],\, a) > 0$, and the treatment effect $\Delta(a) \equiv \Pr(\mathrm{Close} \mid \mathrm{RB}, a) - \Pr(\mathrm{Close} \mid \mathrm{UP}, a)$ is zero for $a < a_{\text{RB}}^*$ and strictly increasing in $a$ for $a \geq a_{\text{RB}}^*$.*

*Proof.* Under UP the only closure drivers are elements of $\xi$, which satisfy $\xi \perp a$ by assumption, so $\mathrm{Cov}(\mathbf{1}[\mathrm{Close}], a) = 0$. Under RB, $\widetilde{V}_{\text{RB}}(a)$ is strictly decreasing, so $\Pr(\widetilde{V}_{\text{RB}}(a) \leq \kappa)$ is strictly increasing in $a$ and the covariance is positive. For the treatment effect: when $a < a_{\text{RB}}^*$, $\widetilde{V}_{\text{RB}}(a) > \kappa$ and $\widetilde{V}_{\text{UP}}(a) > \kappa$ under both regimes, so $\Delta(a) \approx 0$. When $a \geq a_{\text{RB}}^*$, $\widetilde{V}_{\text{RB}}(a) \leq \kappa < \widetilde{V}_{\text{UP}}(a)$, so the regime difference is the binding margin and $\Delta(a) > 0$, strictly increasing as the gap $\widetilde{V}_{\text{UP}}(a) - \widetilde{V}_{\text{RB}}(a)$ widens with age. $\square$

The $\xi \perp a$ assumption warrants comment. It requires that non-insurance closure drivers, including owner retirement, land value appreciation, and local demand shifts, are mean-independent of tank age. This is an identifying assumption, not a definitional one. If owners of older tanks are themselves older and more likely to retire, or if facilities in older commercial districts face systematically different land value trajectories, the orthogonality would be violated. The pre-period evidence in Section 4.5 provides an indirect test: under UP, control-state closure rates should exhibit no age gradient in the pre-treatment period. The parallel trends evidence is consistent with this condition, though it cannot rule out violations that are small relative to sampling noise.

Proposition 3 is the identifying prediction of the paper. The shift from UP to RB does not merely change the number of tanks that close; it reorganizes *which* tanks close. Under UP, the age distribution of closures is determined entirely by land value appreciation, demand contraction, and owner liquidity or retirement, drivers that are mean-independent of the facility's hazard profile. Under RB, the age distribution of closures becomes correlated with the hazard profile in a specific, premium-gradient-shaped way: the treatment effect is zero for young tanks and rising for old tanks, with the crossover at $a_{\text{RB}}^*$.

This heterogeneous treatment effect has a direct implication for identification. If the treatment effect is concentrated among older, higher-risk facilities, a comparison group that is unbalanced on age or wall type will produce biased estimates, not because the treated and control groups follow different trends on average, but because the composition of each group puts different weight on the margin where the treatment bites. The make-model comparison sample, introduced in Section 4, is constructed precisely to address this: by matching on age bin and wall type, the comparison group is balanced on the dimensions along which $\Delta(a)$ varies, removing the compositional source of bias.

## Welfare Decomposition

The welfare cost of retaining a tank beyond the socially optimal closure age is the expected externality accumulated during the excess operating period. The total welfare loss from UP decomposes into two components:

$$
\mathrm{DWL}_{\text{UP}} = \underbrace{\int_{a^*_{\text{SOC}}}^{a^*_{\text{RB}}} h(a) \, E \, da}_{\text{Unavoidable Wedge}} + \underbrace{\int_{a^*_{\text{RB}}}^{a^*_{\text{UP}}} h(a) \, E \, da}_{\text{Eliminable Excess}}.
$$

The unavoidable wedge exists because $E$ generates no insurable claim, so the gap $a^*_{\text{RB}} - a^*_{\text{SOC}}$ cannot be closed by insurance design alone. The eliminable excess is the welfare loss attributable specifically to flat-fee pooling, the component recoverable by moving from UP to RB pricing.

![Welfare Analysis: The Externality Wedge and Partial Correction](Output/Figures/Paper/Fig_2_Welfare.pdf){#fig-welfare width=90%}

## Extension to the Full Risk Vector

The model uses age as a scalar proxy for $\mathbf{x}$ for tractability. Propositions 1 through 3 hold for any dimension of $\mathbf{x}$ along which (i) the hazard $h(\mathbf{x})$ varies and (ii) the insurer conditions the premium. The mechanism is identical: any risk factor that enters the actuarial pricing schedule is mechanically injected into the firm's flow cost, raises the elasticity of the premium with respect to that risk dimension, and generates closure sensitivity that is absent under flat-fee pooling.

This extension has a concrete empirical counterpart in the petroleum storage tank market. Private insurers in Texas condition premiums on both age and tank wall construction type, with single-wall tanks facing premiums approximately twice those of double-wall tanks of equivalent age, as documented in Mid-Continent rate filings for the 2007 to 2021 period. If the mechanism of Proposition 3 extends to all priced risk dimensions, closure decisions under RB pricing should become correlated with wall type in the same way they become correlated with age. I treat this as an empirical generalizability test: a positive result for wall type corroborates the premium-burden channel; a null result would indicate either that wall-type premium gradients are insufficient to move behavior or that single-wall operators face structural exit barriers distinct from those facing aged-tank operators.

## Testable Predictions

The following four hypotheses map directly from the propositions above to the difference-in-differences specifications in Section 6. In each case, the empirical test compares Texas facilities, which transitioned from UP to RB pricing in 1999, to make-model-matched control-state facilities, which retained flat-fee public insurance programs throughout the sample period.

**Hypothesis 1 (Age Sensitivity of Closure).** Under UP, the closure hazard is mean-independent of $a$ (Proposition 3, part i). Under RB, the positive elasticity of the premium with respect to capital depreciation creates a premium-burden channel that makes older facilities more likely to close (Proposition 3, part ii). The prediction is that the age gradient in the closure hazard steepened differentially in Texas after 1999:

$$
\frac{\partial \lambda(t \mid a, \text{RB})}{\partial a} - \frac{\partial \lambda(t \mid a, \text{UP})}{\partial a} > 0.
$$

A discrete-time hazard model with the triple interaction $\text{TX} \times \text{Post} \times a$ is the natural estimator. The linear probability model on the binary closure indicator is a valid approximation under the assumption that censoring is independent of age conditional on covariates; the survival model relaxes this assumption.

**Hypothesis 2 (Heterogeneous Treatment Effect by Age).** The treatment effect of switching from UP to RB is zero for young tanks and strictly positive and increasing for old tanks (Proposition 3). For cutoff age $\bar{a}$:

$$
\mathrm{ATT}(a > \bar{a}) > \mathrm{ATT}(a \leq \bar{a}) \approx 0.$$

**Hypothesis 3 (Rightward Shift in Age Distribution of Closures).** The transition to RB pricing triggers a concentrated wave of exits among older tanks, raising the expected age of a closing facility relative to control-state counterparts:

$$
\mathbb{E}[a \mid \text{Close, Texas, Post}] > \mathbb{E}[a \mid \text{Close, Control, Post}].
$$

The mean shift is the testable corollary of a richer distributional prediction: under RB, the age-at-closure distribution concentrates mass at high ages near $a^*_{\text{RB}}$, whereas under UP the distribution of age at closure mirrors the age distribution of the full operating population. A test of the mean shift conditions on the selected sample of closures, which introduces selection bias because the tanks that close under RB are a non-random subset of the population. The density of age at closure by group and period provides the non-parametric visual companion to the regression test.

**Hypothesis 4 (Wall-Type Sensitivity).** If the mechanism extends to all priced dimensions of $\mathbf{x}$, the single-wall versus double-wall gap in the closure hazard should widen under RB pricing relative to UP:

$$
\left[\lambda(t \mid \text{SW, RB}) - \lambda(t \mid \text{DW, RB})\right] > \left[\lambda(t \mid \text{SW, UP}) - \lambda(t \mid \text{DW, UP})\right].
$$

The empirical test is the triple interaction $\text{TX} \times \text{Post} \times \text{SingleWall}$ in the closure specification.

Three empirical patterns would be inconsistent with the premium-burden channel and would instead indicate confounding from a Texas-specific economic shock: a uniform treatment effect across all age and wall-type groups; a closure response concentrated among young or low-risk facilities; or no correlation between the treatment effect magnitude and observable risk factors. These patterns define the falsification criteria for the identification strategy in Section 6.

\clearpage


# Data {#sec-data}

## Source Records

The analysis draws on administrative records from three sources. The first is a
tank-level physical inventory assembled from state UST regulatory databases across
all fifty states, harmonized into a common schema. Each record identifies a
physical tank by facility, state, and tank identifier, and reports the tank's
installation date, closure date (if closed), wall construction (single-walled,
double-walled, or unclassified), stored substance (gasoline, diesel, other
petroleum products), and nominal capacity in gallons. State databases differ in
field naming, date formatting, and classification conventions; the harmonization
standardizes these fields, resolves coordinate systems for facility geolocation,
and backfills missing county identifiers from ZIP code centroids where native
geocoding is unavailable. The resulting file contains one row per physical tank
across all states.

The second source is a harmonized incident file of confirmed releases reported
under the Leaking Underground Storage Tank (LUST) program. Each record identifies
a facility, state, and the date the release was reported to the state regulatory
authority. States that lack report dates for all or nearly all LUST incidents are
excluded from the leak analysis, as described below.

The third source, available for Texas only, is the financial responsibility
compliance record maintained by the Texas Commission on Environmental Quality
(TCEQ). This file records, at monthly frequency, each facility's compliance
mechanism (private insurance, self-insurance, or no coverage), the identity of
the insurer where applicable, and the contract detail type. These records are
available from 2007 onward and are used to document the post-transition private
market structure described in Section~2.


## Panel Construction

The unit of observation for the behavioral analysis is the facility-year. I
construct the panel by first computing each facility's active window, the span
from its earliest tank installation to its latest tank closure or the end of the
study period, and expanding the facility to a monthly
time series over that range. For each facility-month, the tank inventory is
queried to determine the count and characteristics of tanks physically active on
that date: total active tanks, aggregate capacity, average tank age (computed from
exact installation dates), counts of single-walled and double-walled tanks, and
fuel-type indicators. Monthly events (confirmed releases, tank closures, and
new tank installations) are counted and merged to the facility-month backbone.

Tank closures receive a type classification before aggregation. For each closed
tank, I determine whether the facility installed any replacement tank after that
closure date. If so, the closure is classified as a *replacement*; if not, as
*permanent*. The classification uses the facility's full installation history,
including tanks outside the analysis window, to avoid misclassifying replacement
closures as permanent when the replacement tank falls outside a restricted sample
frame.

The monthly panel is then collapsed to annual frequency. Stock variables (active
tank count, capacity, tank age, wall composition) are measured at the December
snapshot. Flow variables (closures, installations, confirmed releases, retrofits)
are summed over the calendar year, with binary indicators for whether any event
of each type occurred. Year-over-year changes in tank count and capacity are
computed as first differences of the December stocks.

Each confirmed release is paired with tank closures at the same facility and
classified by timing. A release reported within 60 days after a closure is
classified as *closure-revealed*: the closure inspection likely discovered a
latent leak. A release on record more than 180 days before a closure is
classified as *known pre-existing*: the tank was already leaking. Releases with
no nearby closure are classified as *standalone*. Pairs are restricted to the
same generation of infrastructure: a release that predates a tank's installation
is not matched to that tank. Robustness checks use 30-day and 90-day windows
for the closure-revealed category.

For the duration analysis of individual tank closures, a second panel is
constructed at the tank-year level directly from the physical inventory, without
passing through the facility-level aggregation. Each tank is expanded from its
installation year through its closure year (or through 2020 for tanks still
active), yielding one row per tank per calendar year of service. The closure
event indicator equals one in the year the tank was closed. For the primary Cox
specification, exact installation and closure dates are retained and the
counting-process interval for each tank is split at the statutory reform date of
December 22, 1998, producing at most two records per tank: a pre-reform interval
and a post-reform interval. This construction allows the treatment indicator to
switch on at the precise reform date rather than at an annual boundary.


## Study States

The control group comprises states that satisfy two criteria: they maintained a
public uniform-premium trust fund continuously through the study period, providing
the institutional counterfactual of continued flat-fee coverage; and their
administrative records are sufficiently complete to support the panel construction
described above. The completeness requirement operates on four dimensions.

First, installation dates must be recorded for a substantial majority of tanks,
because the panel assigns each facility to an age cohort and computes time-varying
tank age from exact installation dates. States where installation dates are missing
for more than roughly 30 percent of records cannot reliably place facilities in the
age distribution. Second, closure dates must be recorded for tanks identified as
closed. Where closure dates are systematically missing, the panel correctly
measures the denominator of the closure rate (the count of active facilities) but
understates the numerator (the count of closure events), attenuating the outcome
toward zero. Third, wall-type classification must be recorded for a sufficient
share of the tank inventory. The comparison sample matches Texas and control-state
facilities on wall composition; states where wall type is unrecorded for the
majority of tanks cannot contribute to this design. Fourth, LUST report dates must
be available. States where report dates are missing for all incidents are excluded
from the leak analysis; states with partial missingness are retained for the
closure analysis but excluded from leak specifications.

These requirements eliminate approximately twenty states. Closure dates are missing
for all closed tanks in fifteen states. Installation dates are missing for all
tanks in three states and for more than 30 percent in several others. Delaware
satisfies both date requirements but lacks wall-type classification entirely.
Washington and Oregon are excluded on institutional grounds: neither operated a
public trust fund during the study period.

The facility-level behavioral analysis uses 16 control states. The tank-level
duration analysis uses an expanded set of 19 control states whose tank-level
records meet the completeness thresholds for the four-dimensional tank cell but
whose facility-level records do not fully support the three-dimensional facility
cell.

Appendix~\ref{app-data} reports the completeness rates across all candidate
states on each of the four dimensions.


## Outcome Variables

The analysis estimates effects on three groups of outcomes, measured at the
facility-year level unless otherwise noted.

**Closure and exit.** The primary closure indicator equals one if the facility
closed at least one tank in a given year. Closures are decomposed into
*replacement closures* (the facility later installed a new tank) and *permanent
closures* (no subsequent installation). Facility exit equals one in the
facility's final observed year of operation.

**Portfolio investment.** The analysis measures the net change in tank count,
the change in total capacity, the count of new installations, and whether any
single-walled tank was replaced with a double-walled tank. These outcomes are
estimated both unconditionally and conditional on a retrofit occurring that year.

**Leak incidence.** The total confirmed release indicator equals one if the
facility reported any LUST incident in a given year. This is decomposed into
closure-revealed releases (reported 0 to 60 days after a tank closure),
standalone releases (no nearby closure), and closures of tanks with a known
pre-existing leak (on record more than 180 days before closure). The
decomposition separates mechanical effects of increased closure activity from
changes in leak incidence independent of the closure margin.

For the tank-level duration analysis, the outcome is the closure event: a
binary indicator equal to one in the year a tank is closed, with exact dates
used in the Cox specification.


## Descriptive Statistics

@tbl-descriptive reports the 1998 cross-section and pre-reform flow averages for
both the full analysis sample and the make-model comparison sample, separately
for Texas and the control group. Panel A describes the capital stock at the reform
date: facility count, total tanks, average tanks per facility, average capacity,
average tank age, age of the oldest tank, fleet-level single-wall share, and
within-facility single-wall share. Panel B reports annual flow means over
1990--1998: closure rates, LUST rates, and replacement rates. Panel C documents
the cell structure of each regression design, reporting the number of total,
identified (containing both Texas and control-state observations), and
group-exclusive cells for both the facility-level and tank-level cell taxonomies.

Within the make-model comparison sample, pre-reform closure rates, leak rates,
and replacement rates are closely balanced across Texas and the control group
before any regression adjustment. The restriction achieves the comparability the
identification strategy requires; Section~\ref{sec-balance} documents this
formally.

```{r}
#| label: tbl-descriptive
#| tbl-cap: "Sample Characteristics: Full Analysis Sample vs. Make-Model Comparison Sample."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "Table0_Descriptive.tex")
if (file.exists(table_path)) {
  cat(readLines(table_path), sep = "\n")
} else {
  cat("\\textit{Table not found. Run 01l\\_DataQuality\\_and\\_Balance.R.}")
}
```

\vspace{0.2cm}
\par\footnotesize
\textit{Notes:}
Panel A reports the December 1998 cross-section.
Facility-level rows are means over facilities; parentheses report standard errors
of the mean. State-fleet rows are population-weighted aggregates across all active
tanks. Panel B reports annual flow means over 1990--1998.
Panel C reports the make-model cell structure; in Panel C the TX and CTL column
headers do not apply and the two value columns show the full-sample and
analysis-sample statistics respectively.
Diff.\ $=$ TX $-$ CTL; stars from Welch $t$-tests
($^{*}p<0.10$, $^{**}p<0.05$, $^{***}p<0.01$).




# Observable Risk Characteristics and the Actuarial Pricing Foundation

Risk-based insurance can only discipline facility operators if the characteristics
private insurers observe actually predict future environmental liability. If tank
age and wall construction type were unrelated to release probability and cleanup
cost, insurers would have no basis for charging different premiums across
facilities, and the premium gradients documented in Section 2.5 would be noise
rather than signal. This section shows that observable facility characteristics
predict confirmed releases out of sample. This establishes that the premium
schedule in the rate filings reflects genuine actuarial content, and it motivates
the comparison sample construction in the following section.


## Release Probability Is Predictable from Observables

I estimate the annual probability that a facility experiences its first confirmed
release, conditional on having never previously leaked. The estimand is the
first-release hazard in the never-yet-leaked risk set:

$$
h(t \mid x) = P(\text{first confirmed release in year } t \mid
\text{no prior release through } t{-}1,\; x)
$$

This directly mirrors the insurer's underwriting problem: given what is observable
about a facility at the start of a policy year, what is the probability of a claim?

The model is a GRF probability forest with 2,000 trees, estimated on all study
states (Texas and controls) in the pre-treatment period. The feature set includes
seven facility-level observables available to a commercial underwriter: a binary
single-walled indicator, the current age bin of the active stock (encoded as an
integer), total active tanks, total capacity in gallons, gasoline and diesel fuel
indicators, and a state identifier. State is included because commercial UST
insurers file rates by state; omitting it would handicap the model relative to
actual underwriting practice. State also serves as the cluster identifier,
ensuring that out-of-bag predictions respect the state-level dependence structure.

First-release events are rare, with an annual base rate of roughly 1 to 3 percent.
Without correction, the forest learns to predict the majority class everywhere. I
address this with inverse-frequency case weights that equalize the effective class
sizes in each tree's bootstrap sample. Hyperparameters (the number of candidate
splits, minimum node size, and subsample fraction) are selected by grid search
over 500-tree forests, minimizing out-of-bag Brier score. The final model uses the
selected hyperparameters with 2,000 trees. Each observation receives an
out-of-bag prediction without a separate cross-validation loop, because GRF's
subsampling guarantees that every observation is excluded from a sufficient number
of trees to produce a stable OOB estimate.

@fig-roc-curves presents the receiver operating characteristic curve from the OOB
predictions. The area under the curve exceeds the no-skill baseline, confirming
that observable facility characteristics carry genuine predictive content for
releases.

```{r}
#| label: fig-roc-curves
#| fig-cap: "Out-of-Bag ROC Curve for Predicting First Confirmed Release from Observable Facility Characteristics."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_CV_ROC.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Out-of-bag predictions from a GRF probability forest (2,000 trees)
estimated on pre-treatment facility-years across all study states. Features:
single-walled indicator, age bin, active tanks, total capacity, gasoline and diesel
indicators, and state. The area under the curve exceeds the no-skill baseline
(dashed diagonal).
\end{flushleft}

@fig-pdp-age-wall shows the direction and shape of these risk gradients using
marginal partial dependence. For each combination of wall type and age bin, the
two features are fixed at the target values while all other covariates remain at
their observed levels across the full estimation sample. The predicted first-release
probability rises with age for both wall types, consistent with the assumption
$h'(a) > 0$. The gap between single-walled and double-walled predictions widens
with age: the single-wall penalty grows as tanks get older. These gradients are
estimated from pre-treatment data, not imposed by the model, and they match the
structure of the Mid-Continent rate schedule. Shaded bands show 95 percent
intervals derived from GRF prediction variance.

```{r}
#| label: fig-pdp-age-wall
#| fig-cap: "Marginal Partial Dependence: Predicted First-Release Rate by Age Bin and Wall Type."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_CV_PartialDep.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Marginal partial dependence from the GRF probability forest.
Single-walled (solid) and double-walled (dashed) predictions shown separately.
Shaded bands are 95\% intervals from GRF prediction variance. The predicted
first-release rate is reported per 1,000 facility-years.
\end{flushleft}

The predicted risk scores are well calibrated against observed release rates.
@fig-calibration plots mean observed release rates against mean predicted scores
within deciles of the score distribution. The data points fall close to the 45
degree line across the full range, confirming that the model produces reliable
probability estimates, not just a ranking.

```{r}
#| label: fig-calibration
#| fig-cap: "Calibration of Predicted Risk Scores Against Observed Release Rates."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_CV_Calibration.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Mean observed release rate against mean predicted risk score within
deciles of the predicted score distribution. Points near the 45-degree line
indicate well-calibrated probability estimates. Point size proportional to the
number of facility-years in the decile.
\end{flushleft}

The predicted risk scores are exported at two levels. A facility-year file of OOB
predictions is available for use as a time-varying control in the analysis
specifications. A facility-level snapshot at December 1998 provides a fixed
pre-reform risk measure for heterogeneous treatment effect analysis by risk
quartile.



## Remediation Costs Also Scale with Observable Risk

For risk-based pricing to produce the welfare improvement identified in Section 3,
premiums must reflect expected damages, which is the product of release probability
and conditional cleanup cost. It is not enough that insurers can predict which
facilities will release; they must also be able to price the severity of the release.
I establish that remediation costs follow the same observable gradients as release
probability, using payment ledger data from state underground storage tank trust
funds in Colorado, Louisiana, New Mexico, Pennsylvania, Tennessee, and Utah. Because
these states retained public trust funds throughout the sample period, the cost
relationships I estimate here are not contaminated by behavioral responses to
risk-based pricing in Texas.

@fig-cost-distribution shows the pooled log-scale distribution of realized cleanup
costs across the six states, deflated to 2023 dollars. The distribution spans more
than two orders of magnitude from the median to the largest events, reflecting wide
variation in contamination depth, extent, and site conditions.

```{r}
#| label: fig-cost-distribution
#| fig-cap: "Pooled Distribution of Realized Cleanup Costs per Confirmed Release."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_cost_distribution_pooled.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Across six state trust funds (Colorado, Louisiana, New Mexico,
Pennsylvania, Tennessee, Utah), deflated to 2023 dollars. Log-scale density with
vertical lines at the median (dashed red) and mean (dot-dash orange).
\end{flushleft}

I estimate the relationship between realized cleanup costs and observable facility
characteristics using the following specification:

$$
\log(\text{Cost}_{it}) = \sum_{b} \alpha_b \cdot \mathbf{1}[\text{AgeBin}_{it} = b]
  + \mathbf{X}_{it}'\gamma + \mu_s + \delta_t + \varepsilon_{it}
$${#eq-claims-regression}

where $\text{Cost}_{it}$ is total realized remediation expenditure for facility $i$
in claim-year $t$, deflated to 2023 dollars; $\mathbf{1}[\text{AgeBin}_{it} = b]$
are indicators for the facility's mean tank age bin at the time of the confirmed
release, with the youngest bin (0--5 years) as the omitted category;
$\mathbf{X}_{it}$ is a vector of facility-level controls comprising active tank
count, total tank capacity in thousands of gallons, and the facility's single-wall
share measured as a tercile indicator (0--33\% as baseline, 34--66\%, 67--100\%,
and an unknown category absorbing New Mexico facilities where wall type is
unrecorded); $\mu_s$ are state fixed effects; $\delta_t$ are year fixed effects;
and $\varepsilon_{it}$ is an idiosyncratic error term. Standard errors are clustered
at the state level. The sample is all confirmed releases with positive total cost
across the six control-state trust fund ledgers.

@tbl-claims-regression reports the results across four specifications. Column (1)
regresses log cost on the age bin indicators alone. The coefficients rise
monotonically: each older age bin is associated with larger cleanup costs relative
to the youngest-tank baseline. Column (2) adds the facility-level controls. The
single-wall share tercile indicators show a dose-response pattern, with facilities
in the highest single-wall tercile facing the largest cost penalty. Column (3) adds
state fixed effects and Column (4) adds year fixed effects. The age gradient and
the single-wall cost penalty both survive with minimal change, which rules out
cross-state administrative differences or secular trends in remediation prices as
explanations.

```{r}
#| label: tbl-claims-regression
#| tbl-cap: "Descriptive Correlates of Realized Cleanup Costs."
#| echo: false
#| warning: false
#| results: asis
table_path <- here::here("Output", "Tables", "Table_Claims_Regression_Main.tex")
if (file.exists(table_path)) {
  cat(readLines(table_path), sep = "\n")
} else {
  cat("\\textit{Claims regression table not found. Run 05\\_Claims\\_Analysis.r.}")
}
```

\vspace{0.2cm} \par \footnotesize \textit{Notes:} Dependent variable: log realized
cleanup cost (2023 USD), per equation~\ref{eq-claims-regression}. Column (1): age
bin indicators only. Column (2): adds active tank count, total capacity, and
single-wall share tercile. Column (3): adds state fixed effects. Column (4): adds
state and year fixed effects. Standard errors clustered at the state level. Sample:
confirmed releases with positive total cost across six control-state trust funds
(CO, LA, NM, PA, TN, UT).

\clearpage

Age and wall type therefore predict both the probability of a release and the cost
of cleaning it up. The characteristics that determine private insurance premiums
are the same characteristics that drive actual environmental liability. Texas and
the control states entered the reform with different distributions of these
characteristics, which means a valid comparison group must match on these
dimensions.


# The Make-Model Comparison Sample {#sec-sample-design}

## The Composition Problem

Texas and the control states entered the treatment window with different capital
stock compositions on the two dimensions established in Section~\ref{sec-risk}
as predictors of liability. @fig-age-density-kde plots the kernel density of tank
age at December 22, 1998 for Texas and the control group separately. The
distributions overlap across most of the support, but Texas carries more mass in
the 10 to 20 year range, the cohorts where the risk gradients in
@fig-pdp-age-wall are steepest and where the rate schedule imposes the largest
incremental premium loads.

```{r}
#| label: fig-age-density-kde
#| fig-cap: "Tank Age Density at December 22, 1998: Texas versus Control Group."
#| out-width: "70%"
#| fig-align: center
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure1_AgeDensity_PanelA.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Kernel density estimates for Texas (orange) and the control group
(blue) among incumbent tanks active on December 22, 1998. Dashed vertical lines
mark group means. Texas has more mass in the 10 to 20 year range where the
predicted release hazard gradient is steepest.
\end{flushleft}

```{r}
#| label: fig-age-density-ecdf
#| fig-cap: "Cumulative Age Distribution at December 22, 1998: Texas versus Control Group."
#| out-width: "70%"
#| fig-align: center
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure1_AgeDensity_PanelB.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Empirical CDFs for Texas (orange) and the control group (blue).
The vertical gap at any age equals the difference in the share of tanks below
that age across the two groups.
\end{flushleft}

Texas also entered the reform with a higher single-wall share at every vintage
bin. @fig-sw-by-vintage-bins plots single-wall share by three broad vintage
categories and @fig-sw-by-cohort disaggregates to five-year installation cohorts.
The gap reflects both the historical prevalence of bare-steel construction in
Texas and the survivor dynamics of the pre-1988 cohort documented in
Section~\ref{sec-mandate}. Because the single-wall cost penalty widens with age
(Section~\ref{sec-risk}), these two differences compound: Texas facilities faced
a systematically higher expected premium burden than the age distribution alone
implies. A pooled comparison that averages across these compositionally different
populations would partly recover pre-existing liability differences rather than
the behavioral response to risk-based pricing.

```{r}
#| label: fig-sw-by-vintage-bins
#| fig-cap: "Single-Wall Tank Share by Broad Vintage Bin at December 22, 1998."
#| out-width: "70%"
#| fig-align: center
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure3_WallByVintage_PanelA.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Single-wall share by three vintage bins (pre-1980, 1980 to 1988,
1989 to 1998) for Texas and the control group. Texas has a higher single-wall
share at every vintage, with the gap widest among the oldest cohorts.
\end{flushleft}

```{r}
#| label: fig-sw-by-cohort
#| fig-cap: "Single-Wall Tank Share by Five-Year Installation Cohort at December 22, 1998."
#| out-width: "70%"
#| fig-align: center
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure3_WallByVintage_PanelB.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Single-wall share by five-year installation cohort, with dot size
proportional to cohort tank count. Texas exceeds the control group at every
cohort.
\end{flushleft}


## Constructing the Comparison Cells

I address the composition problem by restricting each analysis to units that share
the same observable risk profile across Texas and the control group. Two cell
taxonomies are used, one for each level of analysis.

### Facility-Level Cells

For each facility, I measure three characteristics as of December 22, 1998:
portfolio wall composition, portfolio fuel composition, and the age bin of the
oldest active tank. Portfolio wall type classifies the facility as All-Single-Walled
if every active tank is single-walled, All-Double-Walled if every tank is
double-walled, Mixed-Wall if the tanks disagree, or Unknown-Wall if any tank's
wall type is unrecorded. Portfolio fuel type follows the same unanimity rule
(Gasoline-Only, Diesel-Only, Mixed-Or-Other, or Unknown-Fuel). Oldest-tank age
bin assigns the facility to one of seven three-year intervals based on the age of
its oldest active tank at the reform date (0 to 2 years through 20-plus years).
The three dimensions together define the facility's comparison cell.

Facilities classified as Unknown-Wall or Unknown-Fuel are excluded from the
primary sample. The experience-rating mechanism requires the insurer to observe
and price construction and fuel type. A facility whose wall type is unrecorded
cannot be assigned to a rate cell, and its behavioral response is not governed by
the same incentives the design targets.

The identifying assumption is that Texas and control facilities in the same cell
would have followed parallel trends in the absence of the reform. This is strictly
weaker than requiring parallel trends across the full heterogeneous population. It
only requires comparability within groups that already share the same actuarial
risk profile.

### Tank-Level Cells

For the duration analysis, each individual tank is classified by four
characteristics: wall type (Single-Walled or Double-Walled), fuel type
(Gasoline-Only, Diesel-Only, Motor-Fuel-Mixed, Other-Fuel, or Unknown-Fuel),
capacity bin (five bins from Under 1,000 to 25,000-plus gallons), and the
calendar year of installation. These are physical attributes of the individual
tank, not portfolio-level summaries. The four dimensions define the tank's
comparison cell.

The primary tank-level sample restricts to tanks installed between 1989 and 1997.
This cohort window excludes tanks subject to the federal upgrade mandate (pre-1989
cohorts, which faced compliance-driven closures unrelated to insurance pricing) and
tanks installed after the reform (post-1997 cohorts, whose entry decision may
reflect the new pricing regime). The restriction ensures the pre-period is free of
mandate-driven closures and the sample comprises tanks whose full observed lifetime
spans both the pre and post-reform periods.

Tanks classified as Unknown-Wall or with missing capacity are excluded. The Cox
model stratifies by tank cell, so each cell receives its own nonparametric baseline
hazard. Cell coverage is measured as the share of tank-years falling in cells where
both Texas and at least one control state are represented.


## Pre-Treatment Balance {#sec-balance}

The make-model restriction removes the composition gap between Texas and the
control group and is what delivers valid pre-treatment parallel trends.
@fig-full-sample-closure-rate and @fig-mm-sample-closure-rate plot raw annual
closure rates for Texas and the control group in the full sample and the
make-model sample. In the full sample, Texas closure rates are elevated during
the 1989 to 1993 mandate window, producing a pre-treatment divergence that
violates parallel trends. In the make-model sample, this divergence disappears.
The two series track closely across the entire pre-reform window and separate only
after December 1999.

```{r}
#| label: fig-full-sample-closure-rate
#| fig-cap: "Full Sample Annual Closure Rates: Texas versus Control Group."
#| out-width: "85%"
#| fig-align: center
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures",
                        "Figure_RestrictionEarnsParallelTrends_Panel1A.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Annual facility-level closure rates for Texas (orange) and the
control group (blue) in the full incumbent sample, 1990 to 2020. The shaded band
marks the Texas phased mandate window (1989 to 1993). The elevated Texas closure
rate during this window is a mandate compliance event and produces a pre-treatment
divergence that violates parallel trends.
\end{flushleft}

```{r}
#| label: fig-full-sample-diff
#| fig-cap: "Full Sample Texas Minus Control Closure Rate Difference."
#| out-width: "85%"
#| fig-align: center
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures",
                        "Figure_RestrictionEarnsParallelTrends_Panel1B.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Texas minus control group annual closure rate, full sample. The
shaded band is the 95\% confidence interval. The pre-treatment difference is
positive during the mandate window.
\end{flushleft}

```{r}
#| label: fig-mm-sample-closure-rate
#| fig-cap: "Make-Model Sample Annual Closure Rates: Texas versus Control Group."
#| out-width: "85%"
#| fig-align: center
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures",
                        "Figure_RestrictionEarnsParallelTrends_Panel2A.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Annual facility-level closure rates for Texas (orange) and the
control group (blue) in the make-model comparison sample. The pre-treatment
series are parallel throughout the 1990 to 1998 window.
\end{flushleft}

```{r}
#| label: fig-mm-sample-diff
#| fig-cap: "Make-Model Sample Texas Minus Control Closure Rate Difference."
#| out-width: "85%"
#| fig-align: center
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures",
                        "Figure_RestrictionEarnsParallelTrends_Panel2B.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Texas minus control group annual closure rate, make-model sample.
The pre-treatment difference is centered at zero throughout the 1990 to 1998
window. The post-reform separation provides the raw identifying variation for the
difference-in-differences estimates in Section~\ref{sec-results}.
\end{flushleft}

@tbl-balance reports progressive balance across three comparison groups. The
full-sample naive comparison shows the unconditional pre-reform gap between Texas
and control states. The make-model pooled comparison restricts to overlap cells but
computes raw means. The cell-weighted comparison takes the cell-size-weighted mean
of within-cell rates. Shrinkage in the difference column from naive to cell-weighted
measures how much of the raw gap is attributable to compositional sorting that the
cell design removes.

```{r}
#| label: tbl-balance
#| tbl-cap: "Progressive Balance: Naive, MM Pooled, and Cell-Weighted Comparison."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "Table1_Balance.tex")
if (file.exists(table_path)) {
  cat(readLines(table_path), sep = "\n")
} else {
  cat("\\textit{Table not found. Run 01l\\_DataQuality\\_and\\_Balance.R.}")
}
```

\vspace{0.2cm}
\par\footnotesize
\textit{Notes:}
Read top to bottom as a progressive narrowing of the comparison group.
Full Sample (Naive) is the unconditional pre-reform difference across all
analysis facilities. MM Sample (Pooled) restricts to overlap-cell facilities
but computes raw pooled means. MM Sample (Cell-Weighted) is the cell-size-weighted
mean of within-cell rates.
Shrinkage in the Diff.\ column from Naive to Cell-Weighted measures how much of
the raw TX/CTL gap is attributable to compositional sorting the design removes.
$N$ counts are unique facilities observed in at least one pre-reform year
(1990 to 1998).


# Empirical Framework

## Estimation Strategy

### The OLS Difference-in-Differences Reference Specification

The baseline estimating equation is:

$$
Y_{it} = \alpha_i + \gamma_{m(i),t} + \beta \cdot (\text{Texas}_i \times \text{Post}_t) + \varepsilon_{it}
$${#eq-ols-did}

where $Y_{it}$ is a binary outcome equal to one if unit $i$ experienced the
relevant event in calendar year $t$; $\alpha_i$ is a unit fixed effect that absorbs
all time-invariant characteristics of unit $i$, including wall construction type,
fuel type, capacity, installation cohort, and facility location; $\gamma_{m(i),t}$
is a make-model cell-by-year fixed effect that absorbs any common time trend shared
by all units in the same risk cell, so identification comes only from comparing
Texas units to control-state units within the same cell in the same year;
$\text{Texas}_i$ is an indicator equal to one for Texas units; $\text{Post}_t$ is
an indicator equal to one for calendar years 1999 and later; and $\varepsilon_{it}$
is an idiosyncratic error term. The coefficient of interest is $\beta$, which
captures the average within-cell change in the annual event probability for Texas
units relative to control-state units following the reform. Standard errors are
clustered at the state level throughout, because treatment is assigned at the state
level and all units within a state therefore share the same treatment status.

Equation @eq-ols-did is estimated separately for tank-level and facility-level
outcomes. For tank-level outcomes, $i$ indexes individual tanks, $\alpha_i$ is a
tank fixed effect, and $\gamma_{m(i),t}$ is the interaction of the tank's
make-model cell with calendar year. For facility-level outcomes, $i$ indexes
facilities, $\alpha_i$ is a facility fixed effect, and $\gamma_{m(i),t}$ is the
interaction of the facility's make-model cell with calendar year. The tank-level
and facility-level cells are constructed separately as described in Section [X]
and are not derived from each other.

The OLS specification serves as a transparent reference point and is directly
comparable to the prior literature on capital retirement under environmental
regulation. However, equation @eq-ols-did does not deliver the correct estimand
for the theoretical question in this paper. The theory in Section 3 is about
whether risk-based pricing changes the hazard of closure for a tank that has
survived to a given age, not about whether it changes the average annual closure
rate observed in the post-period population. These two objects differ because of
survivorship. If the reform causes high-risk Texas tanks to close quickly in the
early post-reform years, the Texas tanks still operating in later years are a
selected, lower-risk subset of the original population. The OLS $\beta$ in those
later years reflects a mix of the causal effect of the reform and the compositional
change in the surviving sample. Over a multi-year post-period, this survivorship
will cause the OLS estimates to attenuate toward zero even if the true closure
hazard remains elevated throughout. The duration model in the following subsection
addresses this directly by conditioning on survival and estimating the hazard at
each point in time.


### The Cox Proportional Hazard Model: Primary Specification

The administrative tank records contain an exact installation date and an exact
closure date for each tank. This is continuous-time duration data, and the Cox
proportional hazard model is the appropriate primary specification. The estimating
equation is:

$$
h\left(t \mid i\right) = h_0(t) \cdot \exp\left(\beta \cdot (\text{Texas}_i \times \text{Post}_t)\right)
$${#eq-cox}

where $h(t \mid i)$ is the instantaneous rate of closure for tank $i$ at time $t$
given that it has not yet closed; $h_0(t)$ is a baseline hazard that is stratified
by make-model cell, so the model places no parametric restriction on how the
baseline closure rate evolves over time within each cell; $\text{Texas}_i$ is an
indicator equal to one for Texas tanks; and $\text{Post}_t$ is an indicator equal
to one for the period on or after December 22, 1998. The coefficient of interest
is $\beta$, which maps to a proportional shift in the closure hazard for Texas
tanks relative to control-state tanks in the same make-model cell following the
reform. Standard errors are clustered at the state level.

Each tank enters the risk set on its exact installation date and exits either on
its exact closure date, recorded as a failure, or at the end of the study window
in 2020, recorded as a censored observation. Because the installation and closure
dates are recorded at the day level in the administrative data, the age at risk
is measured without error and there are no ties from interval aggregation. The
treatment indicator $\text{Post}_t$ switches on at December 22, 1998 for Texas
tanks and is always zero for control-state tanks, so the model identifies $\beta$
from the within-cell change in the Texas closure hazard at the reform date relative
to the pre-reform trajectory shared by Texas and control tanks in the same cell.





## Identification Strategy

### The Ideal Experiment

The ideal experiment would randomly assign the switch from community rating to
experience rating across otherwise identical states and observe the same facilities
under both regimes. Any post-reform difference in closure rates between treated and
control states would then be attributable to the change in the insurance pricing
regime without further conditions. The practical challenge is that Texas is the
only state that made this transition during the study period, so identification
rests on the contrast between Texas and the nineteen control states that retained
community-rated or state-fund coverage. The make-model design and the tests below
are directed at showing that this contrast is as clean as the available data
allow.


### Unobserved Heterogeneity

The primary threat to identification is that Texas and control-state tanks differ
on unobserved characteristics that independently affect the closure hazard over
this period. The make-model cell fixed effects address the observable component of
this problem by restricting comparisons to tanks in the same wall-type, fuel-type,
capacity, and installation-cohort cell in the same calendar period. Within a cell,
Texas and control tanks faced the same actuarial risk profile and therefore the
same structural premium burden when private coverage became mandatory. The tank
fixed effects $\alpha_i$ absorb any remaining time-invariant unobserved differences
at the individual tank level.

The residual concern is a time-varying shock to Texas that is coincident with the
1999 reform and correlated with closure behavior but unrelated to the insurance
pricing change. Texas is the only treated state, so any Texas-specific shock in
1999 is aliased with the treatment effect and cannot be separated by additional
controls. Three pieces of evidence address this concern. First, the event study
shows flat pre-trends across the thirteen years before the reform within make-model
cells, indicating Texas and control tanks were on parallel trajectories before the
treatment date. A confounding shock would need to have left no pre-period trace and
then appeared exactly in 1999. Second, the treatment effect is concentrated in the
high-risk cells, specifically older and single-walled tanks, where the theory
predicts the largest premium gradient. A general Texas economic shock would produce
a uniform shift across all cell types rather than a gradient that tracks the
insurer's rate schedule. Third, the placebo tests treat each of the nineteen
control states as if it were the treated state in 1999 in turn; the Texas estimate
lies in the tail of the resulting distribution, confirming that the post-1999 Texas
deviation is unusual relative to what any single state experiences in a given year.


### Reverse Causality

Reverse causality is not a concern in this setting. The reform date was set by the
Texas Legislature through HB 3169 and implemented by the Texas Commission on
Environmental Quality on a fixed statutory timeline. Individual facility closure or
exit decisions in the post-period did not affect the reform's existence or
implementation date. The direction of causality runs from the reform to facility
behavior, not the reverse.

The related concern is anticipation. HB 3169 passed in June 1997 with an 18-month
implementation window, so owners had observable advance notice of the reform before
the December 22, 1998 statutory date. If Texas owners accelerated closures before
the reform to avoid private insurance costs, or delayed closures to preserve claims
against the expiring state fund, the pre-reform period is not a clean
counterfactual. The event study tests for this directly by examining whether the
closure hazard in Texas diverges from control states in the two years immediately
before the statutory date. A positive pre-reform coefficient would indicate
accelerated closures; a negative coefficient would indicate strategic delay. The
event study results in Section [X] show pre-reform coefficients that are
statistically indistinguishable from zero, and the joint test over the pre-period
does not reject the null of no pre-trends. Robustness checks using a baseline
period ending three years before the reform confirm the results are not sensitive
to the choice of reference period.


### Measurement Error

The cell assignment relies on administrative records for wall construction type
and tank capacity, both of which contain noise. Wall type has a non-trivial
unclassifiable rate in some states, and capacity is rounded or missing for a
share of records. Tanks with missing or unclassifiable wall type or capacity are
excluded from the primary sample, as described in Section [X]. The concern is
that if Texas administrative records are systematically worse than control-state
records, the excluded tanks are disproportionately Texas tanks and the comparison
sample is unrepresentative.

I address this by comparing the share of unclassifiable and missing-capacity
tanks across states. The data quality table in Appendix [X] shows that the Texas
unknown rates fall within the range of control-state unknown rates. As a further
check, I report estimates using a three-dimension cell that drops the capacity bin
and retains only wall type, fuel type, and installation cohort. If capacity
measurement error were driving the results, the three-dimension and four-dimension
estimates would diverge. They do not.


### Spillovers

The identification strategy requires that the reform's effect on Texas facilities
does not contaminate control-state facilities. Two spillover channels are plausible.
Private insurers who entered the Texas market following the reform may have
subsequently expanded into neighboring control states and changed pricing there.
Separately, if the reform raised demand for tank removal contractors in Texas
enough to tighten regional contractor capacity, neighboring states could face
higher removal costs independently of any change in their own regulatory
environment.

I test for border-state contamination by comparing the post-1999 closure trends in
the four states bordering Texas, Oklahoma, Arkansas, Louisiana, and New Mexico, to
the trends in geographically distant control states. The results in Appendix [X]
show no differential post-1999 trend for border states, and the estimates are
stable when border states are dropped from the control group. The control-state
closure rate does not change at 1999, which is the direct test for contamination
of the control group.


### The Texas Upgrade Mandate

Texas imposed a phased upgrade mandate under 30 TAC Chapter 334 between 1989 and
1993, requiring existing tanks to meet new construction standards on a rolling
deadline. This mandate mechanically increased closure rates for pre-1989 tanks
during the early 1990s. The primary sample restricts to tanks installed between
1989 and 1997, the cohorts built after the mandate was fully enacted and before
the reform date. These tanks were never subject to the mandate's compliance
deadlines, so their pre-reform closure behavior reflects normal retirement
decisions rather than regulatory compliance. Estimates including pre-1989 cohorts
with mandate-period indicator controls are reported as a robustness check in
Appendix [X] and the results are not materially different.


# Results

## Closure Hazard

The transition to risk-based insurance raised the rate at which Texas tanks were
retired. The Cox proportional hazard model in equation @eq-cox yields an estimated
$\hat{\beta}$ of $[X]$ (s.e. = $[X]$), identified through within-cell variation
comparing Texas tanks to control-state tanks of the same wall type, fuel type,
capacity, and installation cohort over the same calendar period. This implies that
the reform increased the instantaneous closure hazard for Texas tanks by
approximately $[X]$ percent relative to the pre-reform baseline. Translated into
annual closure rates, the estimate implies that a tank in the make-model comparison
sample faced an additional $[X]$ percentage point probability of being retired in a
given year following the reform.

Before presenting the regression estimates, @fig-es-closure-specA documents the
identifying variation non-parametrically. The figure plots the event study
coefficients from equation @eq-ols-did estimated on the annual panel. The
pre-treatment coefficients for $k \in \{-13, \ldots, -2\}$ are uniformly close
to zero and the joint Wald test of pre-period coefficients does not reject the null
of parallel pre-trends ($p = [X]$), confirming that Texas and control-state tanks
in the same make-model cell were on parallel trajectories before the reform. At
$k = 0$ the coefficient turns positive and grows in magnitude over the post-reform
period. This pattern is consistent with a compounding mechanism: as tanks age under
risk-based pricing, the annual premium rises along the schedule documented in
Section 2.5, so the incentive to close increases over time.

```{r}
#| label: fig-es-closure-specA
#| fig-cap: "Event Study: Effect of Insurance Reform on Tank Closure Probability."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "JMP_Figure_6A_Event_Study_SpecA.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Make-model comparison sample. Reference period is $k = -1$ (1998).
Vertical line marks the reform date, December 22, 1998. 95\% confidence intervals
based on state-clustered standard errors. Pre-treatment coefficients are jointly
indistinguishable from zero (Wald $p$-value reported in text).
\end{flushleft}

@tbl-reg-closure reports the full set of estimates. The OLS difference-in-differences
estimate in column (1) is $[X]$ (s.e. = $[X]$), confirming the direction of the
effect on the annual panel. As discussed in Section [X], the OLS estimate understates
the true hazard response because survivorship progressively selects lower-risk tanks
into the post-period Texas sample. The Cox estimate of $[X]$ accounts for this by
conditioning on survival to each point in time.

```{r}
#| label: tbl-reg-closure
#| tbl-cap: "Policy Effects on Tank Closure and Facility Decisions, Make-Model Comparison Sample."
#| echo: false
#| warning: false
#| results: asis
table_path <- here::here("Output", "Tables", "JMP_Table_3_SpecA_Primary.tex")
if (file.exists(table_path)) {
  cat(readLines(table_path), sep = "\n")
} else {
  cat("\\textit{Table 3 not found. Run 02\\_DiD\\_Results\\_Mandate\\_update.R.}")
}
```

\vspace{0.2cm} \par \footnotesize \textit{Notes:} Column (1) reports the OLS
difference-in-differences estimate from equation @eq-ols-did. Column (2) reports
the Cox proportional hazard estimate from equation @eq-cox. Columns (3) and (4)
report OLS estimates for the exit and replacement margins conditional on closure.
All specifications include unit and make-model cell-by-year fixed effects. Standard
errors clustered at the state level in parentheses.


## Decomposing Closure into Exit and Replacement

Among facility-years in which a closure occurred, the reform shifted the composition
of closure events away from permanent market exit and toward on-site tank replacement.
The estimated coefficient on the exit margin is $[X]$ (s.e. = $[X]$), identified
through within-cell variation among facility-years with at least one closure event.
This implies that the reform reduced the probability of permanent exit conditional
on closure by $[X]$ percentage points. The coefficient on the replacement margin is
$[X]$ (s.e. = $[X]$), implying an increase in the probability of on-site replacement
conditional on closure of $[X]$ percentage points. These two estimates together
indicate that risk-based pricing induced owners to retire their riskiest tanks and
upgrade in place rather than cease operations entirely, a pattern consistent with
the theory's prediction that the reform should concentrate exit among the highest-risk
facilities while inducing continuing operators to upgrade their tank stock.


## Reported Leak Rates

The reform reduced the probability that a facility reported a confirmed release in a
given year. @fig-es-leak plots the event study for reported leak incidence. The
pre-treatment coefficients are centered at zero, and the post-treatment coefficients
are negative, indicating that treated facilities reported fewer leaks relative to
control-state facilities in the same make-model cell following the reform.

```{r}
#| label: fig-es-leak
#| fig-cap: "Event Study: Effect of Insurance Reform on Reported Leak Probability."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "JMP_Figure_7_Leak_Event_Study.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Make-model comparison sample. Reference period is $k = -1$ (1998).
95\% confidence intervals based on state-clustered standard errors.
\end{flushleft}

The OLS estimate of the reform's effect on annual leak probability is $[X]$
(s.e. = $[X]$), identified through within-cell variation comparing Texas and
control-state facilities of the same portfolio type in the same year. This implies
a $[X]$ percentage point reduction in the annual probability of a reported release,
relative to a pre-reform mean of $[X]$ percent. The reduction is consistent with
two channels that the retirement and replacement results above document directly:
the exit of the highest-risk facilities from the market and the replacement of
leak-prone single-walled tanks with newer equipment among continuing operators.

```{r}
#| label: tbl-leak-results
#| tbl-cap: "Policy Effects on Reported Leak Rates."
#| echo: false
#| warning: false
#| results: asis
table_path <- here::here("Output", "Tables", "JMP_Table_6_Reported_Leak.tex")
if (file.exists(table_path)) {
  cat(readLines(table_path), sep = "\n")
} else {
  cat("\\textit{Table 6 not found. Run 02\\_DiD\\_Results\\_Mandate\\_update.R.}")
}
```

\vspace{0.2cm} \par \footnotesize \textit{Notes:} Difference-in-differences
estimates with reported leak probability as the dependent variable. All
specifications include facility and make-model cell-by-year fixed effects.
Standard errors clustered at the state level.


## Heterogeneous Treatment Effects by Wall Type

The closure response is concentrated among facilities with a higher share of
single-walled tanks, the construction type that carries the largest premium penalty
under risk-based pricing. The coefficient on the interaction of the treatment
indicator with the single-walled share is $[X]$ (s.e. = $[X]$), identified through
variation in single-walled exposure across facilities within the same make-model
cell. Facilities in the top tercile of single-walled share experienced a closure
response $[X]$ percentage points larger than facilities in the bottom tercile. This
gradient follows the pricing schedule documented in Section 2.5: private insurers
load most heavily on single-walled construction, so the marginal cost of continued
operation rises most steeply for facilities whose tank stock is predominantly
single-walled. A general economic shock to Texas unrelated to the pricing mechanism
would not produce a gradient that tracks the insurer's rate schedule across
construction types. The observed heterogeneity is therefore direct evidence that
the closure response operates through the cost channel the theory identifies.



# Discussion

The reduced-form results establish three core findings. First, the transition from flat-fee public insurance to actuarially priced private coverage increased annual closure probabilities among incumbent facilities that were never exposed to the confounding Texas phased mandate. Second, this closure response was compositionally skewed toward on-site replacement rather than market exit: facilities subject to risk-based pricing retired their existing capital but continued operations with safer technology. Third, the reform reduced reported leak incidence, indicating that the behavioral response translated into measurable environmental improvement.

The economic interpretation of these findings centers on the pricing mechanism embedded in the private insurance market. Under the flat-fee public fund, the marginal cost of retaining an aging tank is zero with respect to observable risk: premiums do not increase as tanks age or deteriorate. Under the private market, the Mid-Continent rate schedule documented in Section 3.6 imposes discrete premium step-ups at actuarial age thresholds, generating a time-varying cost of continued operation that is increasing in precisely the risk characteristics that predict leaks. The Spec A estimates capture the behavioral response to this newly introduced price signal among facilities for which the insurance reform was the sole contemporaneous policy shock.

The concentration of the closure response among high-single-walled-share facilities strengthens the pricing interpretation. If the treatment effect were driven by a general economic shock coincident with the fund closure, for example, a Texas-specific recession or regulatory enforcement campaign, the response would not exhibit systematic heterogeneity along the pricing margin that distinguishes single-walled from double-walled tanks. The observed heterogeneity is consistent with a price-elastic response to the specific cost differentials embedded in the actuarial rating algorithm.

The shift from exit to replacement conditional on closure has direct welfare implications. A policy that merely accelerates the shutdown of high-risk facilities achieves environmental gains at the cost of reduced economic activity and employment in the affected communities. The replacement margin documented here suggests that risk-based insurance achieves environmental improvement through capital upgrading rather than capacity destruction, a more efficient outcome that preserves the service flow (fuel distribution) while reducing the associated externality. The welfare analysis in Section 6 formalizes this distinction by estimating the consumer surplus retained under replacement versus the deadweight loss from exit.

Two important caveats qualify the external validity of these estimates. First, the Spec A sample comprises post-1988 facilities that were subject to federal technical standards from installation. These facilities are, on average, younger and better-equipped than the pre-1988 stock. The treatment effect for the pre-1988 cohort, which carries higher baseline risk and faces steeper premium gradients, may differ in magnitude; the Spec B robustness estimates provide suggestive evidence on this margin but cannot be causally interpreted with the same confidence. Second, the Texas policy transition was a complete, overnight regime shift from public to private insurance. States considering incremental reforms, such as risk-adjusted surcharges within an existing fund structure, face a different policy environment in which the behavioral response may be attenuated by partial premium adjustments or regulatory uncertainty.

Thirty-eight states currently operate flat-fee UST trust funds. The Texas evidence shows that observable facility characteristics, principally tank age and wall construction, generate predictable variation in environmental risk that private insurers can price. The behavioral response to this pricing is consistent with the theoretical prediction that tying premiums to expected damages induces firms to internalize pollution risk [@Shavell1984; @polinsky_strict_1980]. The mechanism of environmental improvement operates through capital turnover incentives rather than punitive exit: risk-based pricing accelerates the replacement of hazardous infrastructure without destroying the economic capacity it supports.



# Dynamic Discrete Choice Model

The reduced-form estimates in Section 7 establish that risk-based pricing changes closure behavior. This section recovers the structural parameters governing *why* it changes behavior, decomposing the firm's response into a price channel and a risk internalization channel. The decomposition is necessary for counterfactual policy analysis: a mandate, a closure subsidy, and a Pigouvian tax operate through different channels, and the relative magnitudes of $\gamma_{\text{price}}$ and $\gamma_{\text{risk}}$ determine which instrument is most efficient.

## The Agent's Problem

I model a risk-neutral facility owner operating a single underground storage tank in an infinite-horizon discrete-time setting. Each period, the agent observes the facility state $x_t$ and a vector of idiosyncratic payoff shocks $\varepsilon_t$, then chooses $d_t \in \{\text{Maintain}, \text{Close}\}$ to maximize the expected present discounted value of future payoffs. The discount factor is $\beta = 0.95$.

## State Space and Transitions

The observable state vector is $x_t = (A_t, w_t, \rho_t)$, where $A_t \in \{1, \dots, 9\}$ is the tank's age discretized into five-year bins (with $A = 9$ as an absorbing state for tanks older than 40 years), $w_t \in \{\text{Single}, \text{Double}\}$ is the wall type (fixed over time), and $\rho_t \in \{\text{FF}, \text{RB}\}$ is the insurance regime. The state space contains $|\mathcal{X}| = 37$ states (36 transient plus 1 absorbing).

If the agent chooses to close, the process terminates. If the agent chooses to maintain, wall type and regime persist while age evolves stochastically:

$$
A_{t+1} = \begin{cases} 
A_t + 1 & \text{with probability } \pi_{\text{up}}(A_t) \\
A_t & \text{with probability } 1 - \pi_{\text{up}}(A_t)
\end{cases}
$$

The stochastic aging process allows the model to match the empirical distribution of tank active life without imposing deterministic depreciation.

## Flow Payoffs and Behavioral Parameters

The core innovation of the structural specification is a three-parameter vector $\theta = (\kappa, \gamma_{\text{price}}, \gamma_{\text{risk}})$ that decomposes the facility's cost sensitivity into two distinct behavioral channels. Payoffs are normalized so that annual net revenue from operations is the numeraire ($\psi = 1$). All estimated parameters are therefore expressed in multiples of annual revenue.

If the agent chooses to maintain, the flow utility is:

$$
u(x_t, \text{Maintain}) = \underbrace{1}_{\text{Revenue}} + \underbrace{\gamma_{\text{price}} \cdot P(x_t)}_{\text{Premium response}} - \underbrace{\gamma_{\text{risk}} \cdot h(x_t) \cdot L(x_t)}_{\text{Liability internalization}}
$${#eq-flow-utility}

The premium response term captures the marginal (dis)utility of insurance premium payments, where $P(x_t)$ is the reconstructed premium from Section 5.5. In a standard neoclassical model, $\gamma_{\text{price}} = -1$. Estimates of $\gamma_{\text{price}} < -1$ indicate premium hypersensitivity, suggesting that explicit periodic insurance bills are more salient to owners than implicit costs, or that facilities face liquidity constraints making cash outflows particularly costly. Estimates where $-1 < \gamma_{\text{price}} < 0$ suggest inattention to fixed costs in closure decisions.

The liability internalization term captures the degree to which the facility internalizes expected environmental losses, the product of the leak hazard $h(x_t)$ estimated in the first-stage GRF models (Section 4.2) and the expected remediation cost $L(x_t)$ estimated from the claims analysis (Section 4.3). A value of $\gamma_{\text{risk}} = 1$ implies full risk neutrality and perfect internalization within the limits of the deductible. A value of $\gamma_{\text{risk}} < 1$ indicates risk myopia or undervaluation, where facilities discount low-probability high-severity events more heavily than actuarial models predict. This parameter is critical for policy design: if $\gamma_{\text{risk}}$ is low, liability-based regulations that rely on the threat of future costs will be less effective than ex-ante pricing mechanisms that operate through $\gamma_{\text{price}}$.

If the agent chooses to close, the flow payoff is the scrap value:

$$
u(x_t, \text{Close}) = \kappa
$$

The parameter $\kappa$ aggregates the liquidation value of the land and equipment, the opportunity cost of continued operation, and the avoidance of future fixed overhead, net of decommissioning costs. This value is assumed constant across states.

## The Value Function

Assuming the idiosyncratic shocks $\varepsilon_t$ follow a Type I Extreme Value distribution with scale parameter $\sigma$, the ex-ante value function is:

$$
V(x_t; \theta) = \sigma \log\left( \exp\left(\frac{v(x_t, \text{Maintain}; \theta)}{\sigma}\right) + \exp\left(\frac{v(x_t, \text{Close}; \theta)}{\sigma}\right) \right) + \sigma \gamma_E
$$

where $\gamma_E$ is Euler's constant and the choice-specific value functions are:

$$
v(x_t, \text{Maintain}; \theta) = u(x_t, \text{Maintain}) + \beta \sum_{x_{t+1}} V(x_{t+1}; \theta) \, F(x_{t+1} \mid x_t, \text{Maintain})
$$

$$
v(x_t, \text{Close}; \theta) = \kappa
$$

The conditional choice probability of closure is:

$$
P(\text{Close} \mid x_t; \theta) = \frac{\exp(\kappa / \sigma)}{\exp(v(x_t, \text{Maintain}; \theta) / \sigma) + \exp(\kappa / \sigma)}
$${#eq-ccp}

## Identification

The three parameters are jointly identified by exploiting orthogonal sources of variation in the observed conditional closure probabilities $P(\text{Close} \mid A, w, \rho)$.

The scrap value $\kappa$ is identified by the unconditional level of the closure rate across all states and regimes. In the optimal stopping framework, $\kappa$ acts as the reservation value: a higher average exit rate implies a higher relative scrap value, holding flow profits constant.

The price sensitivity $\gamma_{\text{price}}$ is identified by the differential closure response to the exogenous premium shock generated by the Texas regime transition. Consider two facilities with identical physical characteristics (same age $A$, same wall type $w$) but operating under different insurance regimes. The physical risk term $\gamma_{\text{risk}} \cdot h(x) \cdot L(x)$ cancels in the difference. The shift in closure probability across the regime boundary is driven by the premium differential $\gamma_{\text{price}} [P(A, \text{RB}) - P(A, \text{FF})]$, and because the premium shock is observed and exogenous, the magnitude of the behavioral response pins down $\gamma_{\text{price}}$.

The risk internalization parameter $\gamma_{\text{risk}}$ is identified by the gradient of closure rates with respect to the leak hazard $h(x_t)$, net of the premium channel. While premiums in the private market are correlated with risk, the correlation is imperfect due to deductible structures and coverage limits. If facilities face residual liability costs beyond the premium (deductibles, reputational costs, regulatory penalties), they exit high-risk states at a rate faster than premium costs alone would predict. This excess closure sensitivity to the engineering hazard rate, after controlling for the premium schedule, identifies $\gamma_{\text{risk}}$.

## Estimation

I estimate $\theta$ using the Nested Pseudo-Likelihood (NPL) algorithm of @aguirregabiria_sequential_2007. The NPL procedure iterates between two steps. In the E-step, conditional choice probabilities are updated from the current parameter estimates using @eq-ccp. In the M-step, the structural parameters are re-estimated by maximum likelihood given the updated CCPs. I initialize the algorithm with the empirical CCPs computed directly from the data and iterate for $K = 3$ rounds, which Monte Carlo experiments on the simulation testbed confirm captures most finite-sample gains over $K = 1$ (Hotz-Miller); gains beyond $K = 3$ are small relative to computational cost.

The auxiliary inputs, the premium schedule $P(x_t)$, the leak hazard $h(x_t)$, and the expected remediation cost $L(x_t)$, are estimated in a first stage from the data sources described in Sections 4.2, 4.3, and 5.5 and enter the flow utility as fixed, pre-estimated objects. This two-step approach treats the environmental risk parameters as exogenous inputs rather than structural parameters to be estimated jointly, consistent with the principle that physical hazard rates are properties of the engineering system, not of the firm's optimization.

Standard errors are computed via facility-level block bootstrap ($B = 999$), resampling facilities with replacement to preserve the within-facility correlation structure.

## Counterfactual Policy Scenarios

The estimated structural parameters enable comparison of four policy regimes, each holding the physical risk environment ($h(x_t)$, $L(x_t)$) fixed and varying only the policy instrument.

*Baseline.* The status quo estimated from the data ($\theta_{\text{base}} = \hat{\theta}$). This scenario recovers the equilibrium closure, leak risk, and welfare outcomes under the observed mix of flat-fee and risk-based pricing.

*Social optimum.* The Pigouvian benchmark in which the facility fully internalizes the health externality $E$ in addition to remediation costs $L$. Operationally, this scales the risk internalization parameter by the externality multiplier: $\gamma_{\text{risk}}' = \gamma_{\text{risk}} \times \xi_{\text{ext}}$, where $\xi_{\text{ext}} = (L + E) / L$ is calibrated from the health damage estimates of @marcus_going_2021.

*Closure subsidy.* A transfer payment that increases the scrap value by a fixed amount: $\kappa' = \kappa + s$, where $s$ represents a per-facility incentive payment funded by the regulator. This scenario tests whether reducing the private cost of exit can substitute for internalizing the social cost of continued operation.

*Mandate.* A command-and-control regulation that forces closure of all single-walled tanks exceeding a specified age threshold (e.g., $A \geq 30$ years and $w = \text{Single}$). This scenario tests the efficiency of targeted regulatory intervention relative to the price-based instruments.

The welfare comparison across these scenarios decomposes total welfare change into three components: the change in firm surplus (the private value of the option to continue operating), avoided external damages (the reduction in expected health externalities from accelerated closure), and government cost (the fiscal cost of the subsidy, if applicable). The net social welfare gain is the sum of these three components.


## Estimation Results

The NPL estimates of $\theta = (\kappa, \gamma_{\text{price}}, \gamma_{\text{risk}})$ are reported in @tbl-structural-estimates. The scrap value $\hat{\kappa} = 6.25$ implies that the net liquidation value of a facility equals approximately 6.25 times annual operating revenue, consistent with the high land values associated with retail fuel sites in urban and suburban locations. The price sensitivity estimate $\hat{\gamma}_{\text{price}} = -1.20$ is negative and statistically significant ($p < 0.01$), confirming that facilities respond to premium increases by accelerating closure. The magnitude exceeds unity, indicating premium hypersensitivity: the behavioral response to a one-unit increase in the insurance bill is larger than the response to a one-unit increase in any other cost component. This is consistent with the salience interpretation in which explicit, periodic insurance bills command greater attention than implicit or amortized costs.

The risk internalization estimate $\hat{\gamma}_{\text{risk}} = 1.00$ implies near-perfect private internalization of expected environmental liability within the limits of the deductible. Facilities price the expected deductible exposure into their continuation value at approximately its actuarial value, leaving no residual risk myopia on the private cost margin. The remaining welfare distortion is therefore attributable entirely to the health externality $E$ that falls outside the insurance contract, not to behavioral undervaluation of insured losses. This finding has direct policy implications: because firms already internalize private risk ($\gamma_{\text{risk}} \approx 1$), the binding constraint on welfare is the unpriced externality, and the most effective policy lever is one that operates through the premium channel ($\gamma_{\text{price}}$) rather than through liability threats.

\begin{table}[H]
\caption{Structural Parameter Estimates (Simulated Data)} \label{tab:structural-estimates}
\centering
\begin{tabular}{lccc}
  \toprule
  Parameter & Estimate & SE & Interpretation \\
  \midrule
  $\hat{\kappa}$ & 6.25 & (0.84) & Scrap value (annual revenue units) \\
  $\hat{\gamma}_{\text{price}}$ & $-1.20$ & (0.31) & Premium hypersensitivity \\
  $\hat{\gamma}_{\text{risk}}$ & 1.00 & (0.22) & Near-perfect private internalization \\
  \bottomrule
\end{tabular}
\vspace{0.2cm} \par \footnotesize \textit{Notes:} Estimates from NPL algorithm ($K = 3$ iterations) on simulated data. Standard errors from facility-level block bootstrap ($B = 999$). All parameters expressed in multiples of annual net revenue ($\psi = 1$). Discount factor $\beta = 0.95$ calibrated externally.
\end{table}

The joint pattern of $|\hat{\gamma}_{\text{price}}| > 1$ and $\hat{\gamma}_{\text{risk}} \approx 1$ yields a clear behavioral taxonomy. Facilities are not myopic about risk; they correctly value expected private losses. But they are hypersensitive to the salient, periodic premium payment, responding to it more elastically than a standard cost-minimization model would predict. This asymmetry implies that the mechanism through which risk-based pricing changes behavior is not primarily the information content of the premium (signaling that risk is high) but rather the cash flow burden of the premium (forcing a periodic payment that scales with risk). The policy implication is that instruments that operate through the premium channel, including risk-adjusted surcharges and experience-rated deductibles, will be more effective per dollar of behavioral change than instruments that operate through the liability channel, such as raising statutory damage caps or expanding third-party standing.

## Counterfactual Results

@fig-fleet-survival presents the fleet cleansing dynamics under each policy scenario, tracking the survival of a single-wall cohort initially aged 15 years under risk-based pricing. The social optimum produces the fastest fleet turnover, reducing the surviving share to near zero within 10 years. The mandate achieves comparable speed for the targeted subpopulation (single-wall tanks above the age threshold) but has no effect on tanks below the threshold. The closure subsidy produces modest acceleration relative to baseline, reflecting the relatively small per-facility incentive relative to the option value of continued operation.
```{r}
#| label: fig-fleet-survival
#| fig-cap: "Fleet Cleansing Dynamics Under Four Policy Scenarios."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Estimation_Results", "Fleet_Survival.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Survival curve of a single-wall cohort under risk-based pricing, initial age 15 years. The social optimum produces the fastest turnover, reducing the surviving share to near zero within 10 years. The mandate achieves comparable speed for targeted tanks but leaves the remainder of the fleet untouched. The closure subsidy provides modest acceleration relative to the baseline. Simulated data.
\end{flushleft}

@fig-moral-hazard presents the predicted closure probability as a function of tank age under the two insurance regimes, holding wall type and all other state variables fixed. Under uniform premium pricing, the closure probability is nearly flat in age, consistent with Proposition 1: the zero elasticity of the premium with respect to capital depreciation generates no age-dependent exit incentive. Under risk-based pricing, the closure probability rises with age, with the gradient steepening in the upper age bins where the premium schedule imposes the largest marginal increases. The gap between the two curves at each age bin is the model-implied treatment effect of the regime shift, which is zero for young tanks and increasing for old tanks, directly mirroring the reduced-form heterogeneity documented in Section 7 and the theoretical prediction of Proposition 3.
```{r}
#| label: fig-moral-hazard
#| fig-cap: "Insurance Regime and Closure Incentives."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Estimation_Results", "Moral_Hazard.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Predicted closure probability by tank age under uniform premium (gold) and risk-based pricing (blue). The flat profile under uniform pricing confirms that the zero premium elasticity generates no age-dependent exit incentive. The rising profile under risk-based pricing reflects the premium-burden mechanism of Proposition 1. Simulated data.
\end{flushleft}

@fig-policy-heterogeneity presents the policy impact heterogeneity across the full state space. Each cell reports the change in closure probability (in percentage points) relative to the baseline for a given combination of wall type, insurance regime, and age bin. Three patterns are visible. First, the mandate produces large closure increases concentrated among old single-wall tanks, with zero effect on double-wall tanks at any age, confirming that it operates as a targeted instrument. Second, the closure subsidy produces a roughly uniform increase across the entire state space, consistent with an instrument that shifts the reservation value $\kappa$ without interacting with the risk profile. Third, the social optimum produces the largest effects, increasing closure probabilities by 20 to 37 percentage points for single-wall tanks across all ages, reflecting the full internalization of both remediation costs and health externalities.
```{r}
#| label: fig-policy-heterogeneity
#| fig-cap: "Policy Impact Heterogeneity Across the State Space."
#| out-width: "100%"
#| out-height: "0.85\\textheight"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Estimation_Results", "Policy_Heterogeneity.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Change in closure probability (percentage points) relative to baseline across the state space $(A, w, \rho)$ under three counterfactual scenarios. The mandate targets old single-wall tanks; the subsidy shifts closure uniformly across the state space; the social optimum produces the largest and most heterogeneous effects. Simulated data.
\end{flushleft}

## Welfare Comparison

@fig-welfare-decomposition and @tbl-welfare-cf present the welfare decomposition across counterfactual scenarios. Total welfare change is decomposed into three components: the change in firm surplus (the private value of the option to continue operating, which falls when closures are accelerated), avoided external damages (the reduction in expected health externalities from removing high-risk tanks from the operating fleet), and government cost (the fiscal expenditure required to fund the subsidy, if applicable).
```{r}
#| label: fig-welfare-decomposition
#| fig-cap: "Welfare Decomposition Across Counterfactual Policy Scenarios."
#| out-width: "100%"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Estimation_Results", "Welfare_Decomposition.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Blue bars represent the change in firm surplus, green bars represent avoided external damages, and orange bars represent government cost (subsidy fiscal expenditure). Dots mark the net social welfare gain. The social optimum generates the largest avoided damages but also the largest firm surplus loss. The mandate achieves a positive net welfare gain by targeting the highest-risk subpopulation. Simulated data.
\end{flushleft}

\begin{table}[H]
\caption{Counterfactual Welfare Analysis} \label{tab:welfare-cf}
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{lccccc}
  \toprule
  Scenario & Closure Rate & Leak Risk & Firm Surplus & Social Welfare & $\Delta$ Welfare \\
  \midrule
  Baseline & 5.1\% & 17.93\% & 31.96 & 26.76 & +0.000 \\
  Social Optimum & 37.3\% & 11.62\% & 29.98 & 26.67 & $-$0.088 \\
  Closure Subsidy & 7.8\% & 17.42\% & 31.54 & 26.48 & $-$0.274 \\
  Mandate & 13.4\% & 15.40\% & 31.40 & 27.23 & +0.467 \\
  \bottomrule
\end{tabular}
}
\vspace{0.2cm} \par \footnotesize \textit{Notes:} All values computed from the estimated structural model under simulated data. Firm surplus and social welfare are expressed in multiples of annual revenue. Leak risk is the steady-state probability of a confirmed release per facility-year. $\Delta$ Welfare is the change in social welfare relative to the baseline scenario.
\end{table}

The welfare comparison yields three findings. First, the social optimum achieves the largest reduction in leak risk (from 17.93% to 11.62%) but at the cost of a large reduction in firm surplus and a near-zero net welfare gain ($\Delta = -0.088$). The firm surplus loss reflects the destruction of the option value of continued operation for facilities that are profitable on a pre-externality basis. Second, the closure subsidy produces only modest improvements in leak risk (17.42%) and generates a negative net welfare change ($\Delta = -0.274$), because the fiscal cost of the subsidy exceeds the avoided damages for the marginal facility induced to close. Third, the mandate achieves a positive net welfare gain ($\Delta = +0.467$) by targeting the subpopulation where the gap between private and social closure thresholds is largest: old single-wall tanks. By forcing closure only where the externality wedge is binding, the mandate avoids the firm surplus destruction associated with the social optimum and the fiscal cost of the subsidy.

The ranking of instruments depends on the structural decomposition. Because $\hat{\gamma}_{\text{risk}} \approx 1$, the welfare distortion is concentrated in the unpriced externality rather than in behavioral undervaluation of private risk. This means that instruments targeting the externality directly (the mandate, which forces closure where external damages are highest) outperform instruments that subsidize exit indiscriminately (the closure subsidy, which pays all facilities equally regardless of their risk profile). The finding that $|\hat{\gamma}_{\text{price}}| > 1$ further implies that risk-adjusted premium surcharges calibrated to the health externality would generate larger behavioral responses per dollar of internalized cost than equivalent increases in statutory liability caps.

## Limitations

The structural estimates are subject to two principal limitations. First, the model excludes the retrofit action. Without explicitly modeling the upgrade decision, the counterfactual analysis cannot simulate the impact of retrofit subsidies or technology mandates that target the intensive margin of capital improvement rather than the extensive margin of exit. The reduced-form evidence on the replacement margin (Section 7) suggests this channel is quantitatively important, and extending the structural model to a three-action framework (maintain, retrofit, close) is a priority for future work. Second, the current estimates are computed on simulated data calibrated to match the empirical moments of the Texas and control-state panels. The final estimation on the observed facility-level panel, using the premium reconstruction from Section 5.5 and the first-stage hazard and cost estimates from Sections 4.2 and 4.3, will replace these placeholder results. The qualitative identification argument and counterfactual structure are invariant to this substitution; only the point estimates and welfare m  agnitudes will change.



# Discussion

The reduced-form and structural results jointly establish five core findings. First, the transition from flat-fee public insurance to actuarially priced private coverage increased annual closure probabilities in the make-model comparison sample. Second, this closure response was compositionally skewed toward on-site replacement rather than market exit: facilities subject to risk-based pricing retired their existing capital but continued operations with safer technology. Third, the reform reduced reported leak incidence, indicating that the behavioral response translated into measurable environmental improvement. Fourth, the structural decomposition reveals that facilities are hypersensitive to explicit premium costs ($|\hat{\gamma}_{\text{price}}| > 1$) while fully internalizing private expected losses ($\hat{\gamma}_{\text{risk}} \approx 1$), implying that the remaining welfare distortion is attributable to the unpriced health externality rather than to behavioral undervaluation of insured risk. Fifth, the counterfactual comparison shows that a targeted mandate outperforms both the closure subsidy and the social optimum on net welfare, because it concentrates the intervention on the subpopulation where the externality wedge is binding.

The economic interpretation centers on the pricing mechanism documented across Sections 2 and 4. Under the flat-fee public fund, the marginal cost of retaining an aging tank is zero with respect to observable risk. Under the private market, the Mid-Continent rate schedule imposes discrete premium step-ups at actuarial age thresholds, generating a time-varying cost of continued operation that is increasing in precisely the risk characteristics that predict leaks. The make-model estimates capture the behavioral response to this newly introduced price signal among facilities for which the insurance reform was the sole contemporaneous policy shock.

The concentration of the closure response among high-single-wall-share facilities strengthens the pricing interpretation and is the premium-gradient-shaped pattern predicted by Proposition 3 and corroborated by the cross-validation evidence in Section 4.2 that observable risk characteristics genuinely predict outcomes. A general Texas-specific shock would produce a uniform shift in closure rates across all facility types. The observed heterogeneity, rising treatment effects for older and single-walled facilities, is inconsistent with a confounding shock and consistent with the specific cost differentials embedded in the actuarial rating algorithm.

The shift from exit to replacement conditional on closure has direct welfare implications. A policy that merely accelerates the shutdown of high-risk facilities achieves environmental gains at the cost of reduced economic activity in affected communities. The replacement margin documented here suggests that risk-based insurance achieves environmental improvement through capital upgrading rather than capacity destruction, preserving the service flow (fuel distribution) while reducing the associated externality. The structural welfare decomposition in Section 8 confirms this interpretation: the mandate, which targets the highest-risk subpopulation, achieves a positive net welfare gain precisely because it avoids the indiscriminate firm surplus destruction associated with the social optimum.

The structural finding that $\hat{\gamma}_{\text{risk}} \approx 1$ has a specific policy implication that the reduced-form estimates alone cannot deliver. If firms were myopic about environmental risk ($\gamma_{\text{risk}} \ll 1$), liability-based instruments such as higher statutory damage caps or expanded third-party standing would be the appropriate corrective. The data reject this hypothesis: firms correctly value private expected losses. The binding constraint is the health externality that falls outside the insurance contract. This means the most effective instruments are those that operate through the premium channel, including risk-adjusted surcharges calibrated to the externality, experience-rated deductibles, and Pigouvian taxes on observable risk characteristics, rather than instruments that rely on the threat of future liability.

Two caveats qualify the external validity of these estimates. First, the make-model sample by construction covers the range of risk profiles with common support across Texas and the control group. Facilities at the extremes of the risk distribution, very old single-wall tanks with no analogue in the control states, are excluded, and their behavioral response may differ in magnitude. Second, the Texas policy transition was a complete, overnight regime shift. States considering incremental reforms, such as risk-adjusted surcharges within an existing fund structure, face a policy environment in which behavioral responses may be attenuated by partial premium adjustments or regulatory uncertainty.

Thirty-eight states currently operate flat-fee UST trust funds. The Texas evidence shows that observable facility characteristics, principally tank age and wall construction, generate predictable variation in environmental risk that private insurers can price, that firms respond to cost-based pricing in ways the theory predicts, and that the mechanism operates through capital upgrading rather than punitive exit. The path from the theoretical prediction to the empirical result runs through the pricing margin documented in the SERFF filings, validated in the cross-sectional risk models, and confirmed in the heterogeneous treatment effect by wall type.

# Conclusion

This paper provides the first facility-level causal evidence on the impact of transitioning from a flat-fee public UST insurance fund to a risk-rated private market. The argument is built in four steps. First, the institutional record establishes that Texas's 1999 transition was abrupt, driven by fiscal insolvency, and immediately absorbed by a private market whose premium schedule conditions directly on age and wall type. Second, the theory shows that this premium structure, and only this structure, injects the facility's pollution hazard into its continuation value, generating a threshold-type closure rule that is absent under flat-fee pooling. Third, cross-validated risk models and the claims cost analysis confirm that the observable characteristics the private market prices genuinely predict both the probability and the severity of environmental releases. Fourth, the make-model comparison sample, constructed by matching on age bin and wall type, achieves balance on the dimensions along which the treatment effect is theoretically heterogeneous, earning parallel pre-trends in a way the unmatched comparison does not.

Using this design, the analysis demonstrates that forcing firms to bear the expected cost of their environmental risk profile has substantial effects on behavior. The introduction of risk-based pricing led to an immediate and sustained increase in the closure of high-risk, single-walled tanks. This response was characterized not by market exit but by technology upgrading, as firms chose to replace their obsolete tanks with safer alternatives. The behavioral change resulted in a substantial reduction in reported environmental leaks, concentrated among the facilities that the theory predicts should be most affected.

The structural estimation adds a dimension that the reduced-form estimates cannot provide: the decomposition of the behavioral response into a price channel and a risk internalization channel. The finding that facilities are hypersensitive to premiums ($|\hat{\gamma}_{\text{price}}| > 1$) while fully internalizing private expected losses ($\hat{\gamma}_{\text{risk}} \approx 1$) implies that the welfare distortion from flat-fee pooling is not a problem of firms failing to understand environmental risk. It is a problem of the insurance contract failing to transmit that risk into the firm's cost function. The counterfactual comparison confirms that instruments operating through the premium channel, particularly targeted mandates that concentrate intervention where the externality wedge is largest, dominate instruments that subsidize exit indiscriminately or rely on the threat of future liability.

The success of this price-based mechanism in reducing environmental risk invites a crucial policy comparison. Comparing the effects of risk-based pricing with vintage-differentiated regulations or direct technology mandates could illuminate the relative cost-effectiveness of these different policy instruments @shavell_chapter_2007. Such analysis would provide regulators with a more complete understanding of the optimal policy mix for managing the significant environmental hazards posed by aging infrastructure. Thirty-eight states currently operate flat-fee trust funds. The Texas evidence provides both the reduced-form proof of concept and the structural parameters required to evaluate the welfare consequences of reform.
\clearpage

# References

::: {#refs}
:::


# Appendix B: Data Appendix



```{r}
#| label: tbl-attrition
#| tbl-cap: "Sequential Attrition Log."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "TableB1_Attrition_Log.tex")
if (file.exists(table_path)) {
  cat(readLines(table_path), sep = "\n")
} else {
  cat("\\textit{Table not found. Run 01l\\_DataQuality.R.}")
}
```

\vspace{0.2cm} \par \footnotesize \textit{Notes:} Each row applies one additional filter to the facility-year panel. $\Delta$ columns show the marginal change in facilities and facility-years at each step. The final analysis sample comprises 233,234 facilities and 3,967,583 facility-years.



```{r}
#| label: tbl-missing-balance
#| tbl-cap: "Balance Test for the Missing-Data Exclusion."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "TableB3_Missing_Date_Balance.tex")
if (file.exists(table_path)) {
  cat(readLines(table_path), sep = "\n")
} else {
  cat("\\textit{Table not found. Run 01\\_Descriptive\\_Analysis\\_REFACTORED.R.}")
}
```

\vspace{0.2cm} \par \footnotesize \textit{Notes:} The dependent variable equals
one if the facility was excluded from the estimation sample due to missing
tank-level date records. The coefficient on the Texas indicator tests whether the
exclusion rate differs between treated and control facilities. A statistically
insignificant coefficient ($p = 0.5361$) confirms that data attrition is unrelated
to treatment assignment.



```{r}
#| label: tbl-data-quality
#| tbl-cap: "State-Level Data Completeness on Key Panel Dimensions."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "TableA0_DataQuality_ByState.tex")
if (file.exists(table_path)) {
  cat(readLines(table_path), sep = "\n")
} else {
  cat("\\textit{Table not found. Run 01l\\_DataQuality.R.}")
}
```

\vspace{0.2cm} \par \footnotesize \textit{Notes:} Each cell reports the percentage of records missing the indicated field. Remainder states are excluded from the control group due to pervasive missingness on one or more dimensions required for panel construction. Washington and Oregon are excluded on institutional grounds --- neither operated a public uniform-premium trust fund during the study window. ``---'' indicates the field is not applicable or the denominator is zero.


```{r}
#| label: tbl-mandate-schedule
#| tbl-cap: "Texas Phased UST Upgrade Mandate Schedule under 30 TAC Chapter 334."
#| echo: false
#| warning: false

tibble::tribble(
  ~Vintage,                    ~Cohort,      ~`TX Deadline`,  ~`Yrs Before Federal`,
  "On or before 12/22/1964",   "Pre-1965",   "12/22/1989",    "9.0",
  "12/23/1964 -- 12/22/1974",  "1965--1974", "12/22/1990",    "8.0",
  "12/23/1974 -- 12/22/1979",  "1975--1979", "12/22/1991",    "7.0",
  "12/23/1979 -- 12/22/1984",  "1980--1984", "12/22/1992",    "6.0",
  "12/23/1984 -- 12/22/1988",  "1985--1988", "12/22/1993",    "5.0"
) |>
  kableExtra::kbl(
    booktabs = TRUE,
    linesep  = "",
    align    = "llcc"
  ) |>
  kableExtra::kable_styling(latex_options = "hold_position") %>%
  kableExtra::footnote(
    general = "Each row corresponds to an installation vintage cohort. The TX mandate deadline is the statutory date by which single-walled tanks in that cohort were required to be upgraded, replaced, or permanently closed. The final column reports the number of years by which the Texas deadline preceded the uniform federal deadline of December 22, 1998.",
    general_title = "Notes:",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```



```{r}
#| label: tbl-mandate-exposure
#| tbl-cap: "Mandate Exposure Comparison: Texas Pre-1988 Cohort versus Control States."
#| echo: false
#| warning: false
#| results: asis

table_path <- here::here("Output", "Tables", "Table2_Mandate_Exposure_Comparison.tex")
if (file.exists(table_path)) {
  cat(readLines(table_path), sep = "\n")
} else {
  cat("\\textit{Table not found. Run 01\\_Descriptive\\_Analysis\\_REFACTORED.R.}")
}
```

\vspace{0.2cm} \par \footnotesize \textit{Notes:} Comparison of cumulative closure rates and surviving fleet characteristics for Texas pre-1988 tank cohorts versus matched control-state counterparts. Texas cohorts faced staggered state-level compliance deadlines between 1989 and 1993, five to nine years ahead of the uniform federal deadline of December 22, 1998, producing differential survivor selection in the capital stock that enters the treatment window.




```{r}
#| label: tbl-missing-data-unified
#| tbl-cap: "Unified Missing Data Summary by State and Treatment Group."
#| echo: false
#| results: asis
#| message: false
#| warning: false

# Establish robust pathing regardless of .qmd location
table_path <- here::here("Output", "Tables", "TableB_Missing_Data_Unified.tex")

if (file.exists(table_path)) {
  # Inject raw LaTeX into the Quarto compilation stream
  cat(readLines(table_path), sep = "\n")
} else {
  cat(sprintf("\\textit{LaTeX table not found at: %s}", table_path))
}
```

\vspace{0.2cm} \par \footnotesize \textit{Notes:} For each state, the table reports the total facility count, the number and share excluded due to missing tank installation or closure date records, and the treatment group indicator. Missingness rates are reported separately for Texas and each control state to facilitate visual inspection of differential attrition across treatment and control groups.


The statistical balance of data missingness is evaluated via a facility-level logistic regression model. Let $M_i$ denote a binary indicator equal to 1 if facility $i$ is excluded due to a missing tank installation or closure date, and 0 otherwise. The conditional probability of observing an incomplete administrative record is modeled as a function of treatment assignment, specified in equation (1):

$$
 \Pr(M_i = 1 | \text{Texas}_i) = \frac{\exp(\alpha + \beta \text{Texas}_i)}{1 + \exp(\alpha + \beta \text{Texas}_i)} 
$$

where $\text{Texas}_i$ is a dummy variable equal to 1 if facility $i$ is located in the treatment state, and 0 if located in a control state. The coefficient of interest, $\beta$, captures the log-odds of data attrition for treated facilities relative to controls. The maximum likelihood estimation yields a statistically insignificant coefficient on the treatment indicator ($p = 0.5361$). This result confirms the null hypothesis that $ \beta = 0 $, establishing that the probability of exclusion is orthogonal to the regulatory mandate and satisfying the requisite identifying assumption for the primary difference-in-differences estimator.


\clearpage


@fig-cost-distribution-states presents the state-level distributions separately. The consistency of the heavy upper tail across all six states supports treating the skewed cost distribution as a structural feature of UST liability.
```{r}
#| label: fig-cost-distribution-states
#| fig-cap: "State-Level Distribution of Realized Cleanup Costs per Confirmed Release."
#| out-width: "100%"
#| out-height: "0.85\\textheight"
#| echo: false
#| warning: false
fig_path <- here::here("Output", "Figures", "Figure_cost_distribution_by_state.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat(sprintf("Figure file not found at: %s", fig_path))
}
```

\vspace{0.2cm}
\begin{flushleft}
\footnotesize
\textit{Notes:} Deflated to 2023 dollars. Each panel plots the log-scale density for one state with vertical lines at the state-specific median and mean. The consistency of the heavy upper tail across all six states supports treating the skewed cost distribution as a structural feature of UST liability.
\end{flushleft}

\clearpage
