---
title: "Econometric Identification of the Six-Parameter Replacement Model with Stay Fixed Effects"
subtitle: "What is normalized, what is assumed, and what the data identifies"
author: "Kaleb Javier"
date: "`r Sys.Date()`"
format:
  pdf:
    documentclass: article
    geometry: margin=1in
    fontsize: 11pt
    linestretch: 1.15
    toc: true
    number-sections: true
    keep-tex: true
    fig-cap-location: top
    fig-pos: "H"
    include-in-header:
      text: |
        \usepackage{amsmath}
        \usepackage{amssymb}
        \usepackage{booktabs}
        \usepackage{float}
        \usepackage{longtable}
        \usepackage{graphicx}
        \usepackage{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.pos = "H", fig.width = 6.5, fig.height = 4.0, dpi = 150
)
library(data.table)
library(ggplot2)
library(knitr)
library(kableExtra)
library(here)

theme_set(theme_minimal(base_size = 10) +
          theme(legend.position = "top",
                plot.title = element_text(face = "bold"),
                plot.subtitle = element_text(color = "grey30"),
                strip.text = element_text(face = "bold")))
```

```{r load-fit, include=FALSE}
# Canonical clean (T007 Phase 3) 6p+stayFE fit; fall back to the dropSDMO fit.
.fit_candidates <- c(
  here("Output","Estimation_Results","Model_Replacement_6paramFE_profile_clean_observed.rds"),
  here("Output","Estimation_Results","Model_Replacement_6paramFE_profile_observed_dropSDMO.rds")
)
.fit_path <- .fit_candidates[file.exists(.fit_candidates)][1]
fit_ok <- !is.na(.fit_path) && length(.fit_path) == 1L
if (fit_ok) {
  fit <- readRDS(.fit_path)
  ca  <- fit$cache
  lut <- ca$state_lut
  cell_dt <- data.table(
    s_idx  = lut$s_idx,
    A_bin  = lut$A_bin,
    wall   = ifelse(lut$w_state   == 1L, "SW", "DW"),
    regime = ifelse(lut$rho_state == 1L, "FF", "RB"),
    P      = ca$P_vec       * 10000,   # $/yr per tank
    h      = ca$h_vec,                 # annual leak probability
    haz    = ca$hazard_loss * 10000    # $/yr expected loss
  )
}

# Estimation sample (empirical CCPs for the identification figures)
.obs_path <- here("Data","Analysis","dcm_obs_panel_observed.csv")
obs_ok <- file.exists(.obs_path)
if (obs_ok) {
  obs <- fread(.obs_path)
  emp <- obs[, .(
    n_cell = .N,
    P_M    = mean(y_it == 0L),
    P_E    = mean(y_it == 1L & (is.na(I_replace) | I_replace == 0L)),
    P_R    = mean(y_it == 1L & !is.na(I_replace) & I_replace == 1L),
    P_close= mean(y_it == 1L)
  ), by = s_idx]
  emp[, P_R_cond := P_R / pmax(P_close, 1e-12)]
  emp[, P_E_cond := P_E / pmax(P_close, 1e-12)]
  if (fit_ok) emp <- merge(cell_dt, emp, by = "s_idx", all.x = TRUE)
}
```

# Purpose and scope

This appendix states, in econometric terms, exactly what the six-parameter
replacement model with geographic stay fixed effects (the "6p+stayFE"
specification, `Model_Replacement_6paramFE_profile_clean_observed.rds`)
assumes, normalizes, and identifies. It separates three categories that are
easily conflated: (i) *normalizations* that are required for any discrete-choice
model to be identified; (ii) *sign and support restrictions* that the
researcher imposes by parameterization and which are testable; and (iii) the
*identifying variation* in the data that pins each structural parameter.

The model is a single-agent dynamic discrete-choice model over three actions —
maintain, exit, replace — on a state space of 32 cells (8 age bins $\times$ 2
wall types $\times$ 2 pricing regimes). Estimation is by nested
pseudo-likelihood (NPL); the geographic fixed effects are profiled out at each
NPL iteration.

# The flow utility and choice probabilities

The per-period flow utilities at state $s$ are

$$
u_M(s) = 1 + \gamma_{\text{price}}\, P(s) - \gamma_{\text{risk}}\, \ell(s),
\qquad
u_E(s) = \kappa_{w(s)},
\qquad
u_R(s) = -K_{w(s)} = -\exp\!\big(\widetilde{K}_{w(s)}\big),
$$

where $P(s)$ is the annual premium, $\ell(s) = h(s)\cdot L(s)$ is the annual
expected leak loss, $w(s)\in\{\text{SW},\text{DW}\}$ is the wall type, and the
leading constant $1$ in $u_M$ is the normalized annual net revenue
$R=\$10{,}000$. Premium and hazard enter the **maintain** utility only.

The choice-specific values combine flow utility with discounted continuation:

$$
v_M(s) = u_M(s) + \beta\, F_M(s,\cdot)\, V,\qquad
v_E(s) = u_E(s),\qquad
v_R(s) = u_R(s) + \beta\, F_R(s,\cdot)\, V,
$$

with $V$ the ex-ante value function solved from the Aguirregabiria–Mira linear
inversion $V = (I-\beta M)^{-1} R$ (exit is absorbing, so $F_E$ contributes
nothing to $M$). The conditional choice probabilities are logit:
$P_j(s) = \exp(v_j(s)/\sigma)\big/\sum_k \exp(v_k(s)/\sigma)$ with $\sigma=1$.

The geographic stay fixed effect $\alpha_g$ (one per control state $g$, with
$\alpha_{\text{TX}}\equiv 0$) enters at the choice-probability stage and is
added to the **maintain and replace** values but not exit:

$$
P_j(s,g) = \frac{\exp\big(v_j(s) + \alpha_g\,\mathbf{1}[j\in\{M,R\}]\big)}
                {\sum_k \exp\big(v_k(s) + \alpha_g\,\mathbf{1}[k\in\{M,R\}]\big)}.
$$

This is the literal "stay" interpretation: $\alpha_g$ shifts the value of
remaining in the industry (maintain or replace) relative to leaving it (exit).
Crucially, $\alpha_g$ is **not** inside the value-function inversion: $V$ is
computed from the structural utilities alone. The fixed effect therefore shifts
contemporaneous stay-vs-exit odds but does not propagate through continuation
values, and it is dropped entirely in counterfactual equilibrium re-solves
(the "Semantic 2 / nuisance" convention). Section\ \ref{sec-fe} returns to the
implications of this choice for the Texas counterfactual.

# What is normalized, assumed, and estimated {#sec-assume}

```{r tbl-assume}
at <- data.table(
  quantity = c("$\\sigma^2$ (logit scale)",
               "$u_M$ constant $=1$ ($=R$)",
               "$\\beta$ (discount)",
               "$\\alpha_{\\text{TX}}\\equiv 0$",
               "$\\kappa_{\\text{SW}}, \\kappa_{\\text{DW}}$",
               "$K_w = \\exp(\\widetilde{K}_w)$",
               "$\\gamma_{\\text{price}}$",
               "$\\gamma_{\\text{risk}}$"),
  status = c("Normalization (required)",
             "Normalization (required)",
             "Fixed a priori",
             "Normalization (FE reference)",
             "Estimated, sign free",
             "Estimated, sign imposed $>0$",
             "Estimated within bound",
             "Estimated within bound"),
  detail = c("Scale not separately identified from utility magnitudes.",
             "Only utility differences identified; one level pinned.",
             "Not identified from CCPs alone; set to 0.95.",
             "Other $\\alpha_g$ are deviations from Texas.",
             "Enters raw (no transform): data reveals the sign.",
             "$\\exp(\\cdot)$ forces replacement to be a flow cost.",
             "Bound $[-3,-1]$; sign $<0$ is definitional (premium is a cost).",
             "Bound $[0,0.15]$; sign $>0$ is definitional (loss is a bad).")
)
kable(at, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("Quantity","Status","Detail")) |>
  kable_styling(font_size = 9, latex_options = "hold_position")
```

\vspace{0.1cm}
\begin{flushleft}\footnotesize
\textit{Notes:} Three categories of restriction. Normalizations and $\beta$ are
required or conventional and carry no economic content beyond the choice of
units. The two genuine researcher-imposed restrictions are the positivity of
the replacement cost (via the $\exp$ parameterization) and the box constraints
on $(\gamma_{\text{price}},\gamma_{\text{risk}})$. The exit value $\kappa_w$ is
unrestricted in sign. The signs of $\gamma$ follow from the meaning of the
regressors; the only substantive prior is the floor on $\gamma_{\text{price}}$
away from zero.
\end{flushleft}

## Robustness: are the imposed restrictions binding?

The two imposed restrictions are testable by re-estimation. (i) Replace
$\widetilde{K}_w$ with a free $K_w$ (no $\exp$) and verify $\widehat{K}_w>0$ at
the optimum; if so, the positivity restriction does not bind. (ii) Widen the
$\gamma$ bounds to $[-6,0]$ and $[0,1]$ and verify the estimates remain interior
to the original $[-3,-1]$ and $[0,0.15]$; if so, the box constraints do not
bind. Reporting both turns an assumption into a verified-innocuous restriction.
These re-estimations are out of scope for this appendix but should accompany the
main results table.

# Identification of $\gamma_{\text{price}}$ {#sec-gp}

$\gamma_{\text{price}}$ is identified from variation in the premium $P(s)$ that
is independent of the hazard $h(s)$. Under flat-fee (FF) pricing the premium is
set administratively and is (within regime) uncorrelated with the hazard; the FF
cells therefore trace out the maintain response to price holding risk fixed.
Under risk-based (RB) pricing the premium covaries with the hazard, so the RB
cells contribute fit but cannot separate $\gamma_{\text{price}}$ from
$\gamma_{\text{risk}}$.

Formally, in the static-utility limit the log odds of maintaining against the
better closure option shift linearly in $P$:
$\partial \log[P_M/P_{\text{close}}] / \partial P = \gamma_{\text{price}}/\sigma$
at fixed $\ell$. The FF cells provide the required $P$-variation orthogonal to
$h$.

```{r fig-Ph}
#| fig-cap: "Cell-level joint distribution of annual premium and hazard. FF cells span a range of hazards at roughly constant premium (the source of $\\gamma_{\\text{price}}$ identification); RB cells show premium covarying with hazard (fit only)."
if (exists("cell_dt")) {
  ggplot(cell_dt, aes(x = h, y = P, color = regime, shape = wall)) +
    geom_point(size = 3, alpha = 0.85) +
    scale_color_manual(values = c(FF = "#E76F51", RB = "#2A9D8F")) +
    scale_y_continuous(labels = scales::dollar) +
    labs(x = "Annual leak probability  h(s)",
         y = "Annual premium  P(s)  ($/tank-year)",
         color = "Regime", shape = "Wall",
         title = "Price-vs-risk identification geometry",
         subtitle = "FF cells: vertical slices (h varies at fixed P). RB cells: sloped locus.")
} else cat("*[fit not available — render after the 6p+FE fit exists]*")
```

```{r fig-logodds-P}
#| fig-cap: "Empirical maintain log-odds against premium within FF cells, colored by hazard tercile. A common negative slope across hazard terciles is the reduced-form signature of $\\gamma_{\\text{price}}<0$ identified independently of $h$."
if (exists("emp") && "P" %in% names(emp)) {
  d <- emp[regime == "FF" & !is.na(P_M) & n_cell > 0]
  d[, logodds_M := log(pmax(P_M,1e-6) / pmax(1 - P_M,1e-6))]
  d[, h_tercile := cut(h, breaks = quantile(h, c(0,1/3,2/3,1), na.rm=TRUE),
                       include.lowest = TRUE,
                       labels = c("low h","mid h","high h"))]
  ggplot(d, aes(x = P, y = logodds_M, color = h_tercile)) +
    geom_point(aes(size = n_cell), alpha = 0.8) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.7) +
    scale_x_continuous(labels = scales::dollar) +
    scale_size_continuous(guide = "none") +
    labs(x = "Annual premium  P(s)  ($/tank-year)",
         y = "Empirical maintain log-odds  log[P_M/(1-P_M)]",
         color = "Hazard tercile",
         title = "Maintain log-odds vs premium, FF cells",
         subtitle = "Slope holding hazard fixed identifies gamma_price")
} else cat("*[estimation sample not available]*")
```

# Identification of $\gamma_{\text{risk}}$ {#sec-gr}

$\gamma_{\text{risk}}$ is identified from variation in the expected loss
$\ell(s)=h(s)L(s)$. Two honest caveats apply. First, the hazard is strongly
correlated with age, which also drives the continuation value; the structural
separation relies on the modeling restriction that $\ell$ enters the *current*
maintain payoff while age enters only through $V$. Second, within FF (where $P$
is flat) the cross-cell hazard variation traces the risk response, but the
cleanest joint variation in $(P,\ell)$ comes from the RB locus where price and
risk move together. Identification of $\gamma_{\text{risk}}$ is therefore
weaker than that of $\gamma_{\text{price}}$ and leans more on functional form;
the figure in Section\ \ref{sec-gp} (the FF vertical slices and the RB locus)
is the relevant geometry.

# Identification of the exit values $\kappa_{\text{SW}}, \kappa_{\text{DW}}$ {#sec-kappa}

The exit value $\kappa_w$ is identified by the frequency of full exit relative to
maintaining, separately by wall type. With $u_E$ constant within wall, $\kappa_w$
is pinned by how often facilities choose exit given the (state-varying) maintain
value. The sign is not restricted: a positive $\kappa_w$ means exit carries a
net flow benefit (scrap value, avoided future premium and risk), a negative one
means exit is costly.

```{r fig-exit}
#| fig-cap: "Empirical exit share among closures, P(E | closure), by age and wall. The level and age-profile of full exit identify the exit values $\\kappa_w$."
if (exists("emp") && "P_E_cond" %in% names(emp)) {
  d <- emp[!is.na(P_E_cond)]
  ggplot(d, aes(x = A_bin, y = P_E_cond, color = wall, linetype = regime)) +
    geom_line(linewidth = 0.8) + geom_point(aes(size = n_cell), alpha = 0.7) +
    scale_x_continuous(breaks = 1:8,
      labels = c("0-5","5-10","10-15","15-20","20-25","25-30","30-35","35+")) +
    scale_size_continuous(guide = "none") +
    labs(x = "Age bin", y = "P(Exit | closure)", color = "Wall", linetype = "Regime",
         title = "Exit share among closures, by age and wall",
         subtitle = "Identifying variation for kappa_SW, kappa_DW")
} else cat("*[estimation sample not available]*")
```

# Identification of the replacement costs $K_{\text{SW}}, K_{\text{DW}}$ {#sec-K}

The replacement cost $K_w$ is identified by the conditional replace share among
closures, $P(R\mid\text{closure})$, by age and wall: a higher conditional
replace share implies a smaller $K_w$ relative to $\kappa_w$. Under the earlier
*end-of-year* stamping this collapsed for double-walled cells --- every DW exit
was routed out of the DW cells (Section\ \ref{sec-worsttank}), leaving DW cells
with only replaces, so $P(R\mid\text{cl})=1$ mechanically and $K_{\text{DW}}$ was
identified only off the DW *close rate* under the maintained assumption that
every close is a replace. **Beginning-of-year (decision-time) stamping
(Ticket 013) resolves this:** the DW cells now retain their exits, all-DW
closures split into genuine exits and replaces ($P(R\mid\text{cl})\approx 0.17$
on the corrected panel, no longer pinned at 1), and $K_{\text{DW}}$ is now
identified off actual within-DW exit-vs-replace variation rather than functional
form alone. The residual caveat is no longer wall-routing but the single-tank
*average-age* reset on replacement (the replace off-support of
@tbl-audit-d1): the rich partial-closure/replacement margin still awaits the
portfolio extension before its price responsiveness should be pushed hard.

```{r fig-replace}
#| fig-cap: "Conditional replace share among closures, P(R | closure), by age, wall, and regime, on the beginning-of-year-stamped panel. With decision-time stamping the DW series is no longer pinned at 1 --- DW cells retain their exits --- so both wall types now carry exit-vs-replace variation identifying $K_{\\text{SW}}$ and $K_{\\text{DW}}$."
if (exists("emp") && "P_R_cond" %in% names(emp)) {
  d <- emp[!is.na(P_R_cond)]
  ggplot(d, aes(x = A_bin, y = P_R_cond, color = wall, linetype = regime)) +
    geom_line(linewidth = 0.8) + geom_point(aes(size = n_cell), alpha = 0.7) +
    scale_x_continuous(breaks = 1:8,
      labels = c("0-5","5-10","10-15","15-20","20-25","25-30","30-35","35+")) +
    scale_size_continuous(guide = "none") +
    labs(x = "Age bin", y = "P(Replace | closure)", color = "Wall", linetype = "Regime",
         title = "Conditional replace share among closures",
         subtitle = "Beginning-of-year stamping: DW no longer pinned at 1; both walls identify K")
} else cat("*[estimation sample not available]*")
```

# The stay fixed effect: identification, normalization, and the Texas counterfactual {#sec-fe}

The geographic stay FE $\alpha_g$ is identified off the residual stay-vs-exit
margin in state $g$ relative to the Texas baseline ($\alpha_{\text{TX}}\equiv0$).
Because exit is absorbing and carries no FE, a common shift to the two stay
actions does **not** cancel from the within-state softmax; the FE is point
identified off the exit frequency net of the structural prediction.

Two design facts matter for counterfactuals. First, $\alpha_g$ enters the
choice probabilities but **not** the value-function inversion, so it does not
alter dynamic incentives; it is a measurement intercept on the stay-vs-exit
margin. Second, $\alpha_g$ is dropped in counterfactual equilibrium re-solves,
which use the structural $\theta$ only. For a **Texas** counterfactual this is
not a mis-leveling, because Texas *is* the reference category: there is no
Texas-specific intercept to add back. The welfare change is, in addition,
internally consistent because the FE is absent from both the baseline and the
counterfactual equilibrium and cancels in the difference.

The one residual concern is that the counterfactual *baseline* is a fresh
$\alpha=0$ equilibrium re-solve, whereas the estimation-time Texas value
function was inverted against the geo-weighted equilibrium probabilities. These
need not coincide exactly. The check is direct: the baseline re-solve should
reproduce the fitted choice probabilities on the Texas cells.

```{r fe-validation}
#| fig-cap: "Validation: counterfactual baseline re-solve vs fitted CCPs on Texas cells (s_idx 17-32). Points on the 45-degree line confirm the alpha=0 re-solve reproduces Texas's fitted behavior, i.e. Texas is at the correct level in the counterfactual."
.cf_path <- here("Output","Estimation_Results","CF_TX_FlatFee_results.rds")
if (file.exists(.cf_path) && exists("fit") && !is.null(fit$P_hat)) {
  cfr <- readRDS(.cf_path)
  tx_s <- 17:32
  dd <- data.table(
    fitted   = as.numeric(fit$P_hat[tx_s, ]),
    resolved = as.numeric(cfr$P_baseline[tx_s, ]),
    action   = rep(c("maintain","exit","replace"), each = length(tx_s))
  )
  max_abs <- dd[, max(abs(fitted - resolved))]
  ggplot(dd, aes(x = fitted, y = resolved, color = action)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(size = 2.4, alpha = 0.85) +
    labs(x = "Fitted CCP (fit$P_hat, TX cells)",
         y = "CF baseline re-solve (alpha = 0)",
         color = "Action",
         title = "Texas level check: re-solve vs fit",
         subtitle = sprintf("Max abs discrepancy across TX cells = %.4g", max_abs))
} else {
  cat("*[Run 04o_CF_TX_FlatFee.R and ensure fit$P_hat exists to render this check]*")
}
```

\vspace{0.1cm}
\begin{flushleft}\footnotesize
\textit{Notes:} If the maximum absolute discrepancy is at the convergence
tolerance ($\sim10^{-7}$), the $\alpha=0$ re-solve reproduces Texas's fitted
behavior and the counterfactual baseline is correctly leveled for Texas. A
visible departure from the 45-degree line would indicate that the geo-weighted
estimation equilibrium and the pure-structural re-solve disagree on the Texas
cells, in which case welfare *levels* (not the sign of the change) require
re-examination.
\end{flushleft}

# Caveat carried throughout: the worst-tank wall classification {#sec-worsttank}

The wall state is assigned by the worst-tank rule (`04b`): a facility is coded
double-walled only if *all* of its tanks are double-walled; any single
single-walled tank makes the facility "Mixed" and the facility is coded
single-walled. Mixed facilities are systematically larger, multi-tank
facilities that are mid-retrofit. Two consequences propagate into the
identification above and must be stated wherever SW results are reported:

1. **The SW pool is a composition blend.** It mixes genuine all-SW facilities
   with large, mostly-double-walled facilities carrying one straggler SW tank.
   The estimated $\kappa_{\text{SW}}$ and $K_{\text{SW}}$ are therefore averages
   over a heterogeneous pool, and apparent size-sensitivity of single-walled
   behavior (for example, the partial-closure rate rising with facility size) is
   *partly a composition artifact* of the worst-tank rule rather than a pure
   wall effect.

2. **The apparent $P(R\mid\text{cl})=1$ for double-walled cells was a routing
   artifact, not a behavioral fact.** Grouping closures by *beginning-of-year*
   (decision-time) wall composition shows that all-double-walled facilities exit
   on 83.5\% of their closures --- essentially the same rate as single-walled
   facilities (86.4\%). DW facilities are not less likely to walk away. The
   degeneracy arose only because the state had been stamped from the
   *end-of-year* portfolio: a DW facility that exits has zero tanks at year-end,
   the worst-tank rule defaulted it to SW, and the exit landed in an SW cell ---
   so every DW exit was routed out of the DW cells (Section\ \ref{sec-audit}),
   leaving the DW cells with only replaces ($P(R\mid\text{cl})=1$) and
   contaminating the SW exit margin with DW-origin exits. Beginning-of-year
   stamping (Ticket 013) corrects the routing: all-DW closures now split into
   exits and replaces within the DW cells, the manufactured degeneracy of
   Section\ \ref{sec-K} is removed, and $\kappa_{\text{SW}}$ no longer absorbs
   DW-origin exits. The remaining single-tank limitation is the portfolio
   \emph{average-age} reset on replacement, deferred to the scale extension ---
   not this wall-routing channel.

A robustness re-slice under a majority-rule wall assignment
($w=\text{DW}$ iff more than half of tanks are DW) or an all-DW rule, reporting
how $\widehat{\theta}$ and the empirical $P(\cdot\mid\text{cl})$ move, would
characterize the sensitivity of the structural estimates to this rule.

# What is not identified

For completeness, three objects are not identified by this design and should not
be interpreted structurally:

1. **Common across-action level shifts.** A constant added to all three action
   values at a state cancels from the choice probabilities. Only differences
   (and the stay-vs-exit shift carried by $\alpha_g$) are identified.

2. **The discount factor $\beta$.** Not identified from choice probabilities
   alone; fixed at 0.95.

3. **The portfolio average-age reset on replacement.** Under the corrected
   beginning-of-year stamping the double-walled exit/replace split *is* now
   identified (Section\ \ref{sec-K}: DW cells retain their exits), so this is no
   longer the binding non-identification. What remains outside this
   specification is the multi-tank replacement dynamic itself: $F_{\text{R}}$
   resets a replacing facility to age bin 1, but a one-tank swap at a multi-tank
   facility does not reset the portfolio average age (the replace off-support of
   @tbl-audit-d1). Separating a genuine like-for-like replacement from
   within-facility tank churn requires the portfolio state and is deferred to the
   scale-incorporation extension; see Sections\ \ref{sec-K} and
   \ref{sec-worsttank}.

# State-coding audit: pipeline invariants underlying identification {#sec-audit}

Every identification claim above rests on a data-generating invariant: that the
state conditioning a choice is the state the firm actually faced when it chose.
This section displays the invariants directly, so that a coding error in the
02b$\rightarrow$04b$\rightarrow$structural pipeline appears as a *visibly broken
invariant* rather than a plausible-looking estimate. The diagnostics are
read-only (`T012`, `T013` in `Code/Analysis/`); they are reproduced here because
they materially qualify the identification of $\kappa_w$ and $K_w$.

The unifying finding: the 32-cell state treats a facility as a single tank (one
age, one wall, full reset on replace), but a facility is a *portfolio*, and the
single-tank coding leaks precisely where portfolio dynamics live --- which is
exactly the wall-specific and replacement margins. The diagnostics originally
surfaced a *state-stamping* defect: the decision row's state was read from the
*end-of-year* (post-action) portfolio, when the correct conditioning state is the
*beginning-of-year* (decision-time) portfolio. That defect has since been
corrected (beginning-of-year stamping, Ticket 013), and the tables and figures
below are computed on the corrected panel. They serve two purposes now. First,
they confirm the fix: the wall-cell leakage that contaminated $\kappa_w$ and
$K_w$ is eliminated (closure mis-celling falls from 16.2\% to 0\%; the genuine
all-DW share of the DW replace cell rises from 16.6\% to 100\%). Second --- and
this is what we carry forward to discuss --- they isolate the *residual* mismatch
that beginning-of-year stamping does **not** remove, because it is not a coding
choice but a *modelling assumption*: representing a multi-tank facility by a
single age, a single wall, and a full age reset on replacement. Two known
limitations remain visible after the fix, and both are intrinsic to that
single-tank abstraction rather than to the pipeline:

- **Backward age under Maintain (a real limitation).** Because the state is the
  portfolio *average* age, a partial closure that retires an old tank *lowers*
  the facility's average age --- a move the monotone $F_{\text{M}}$ kernel
  (age stays or advances by one) cannot represent. This accounts for essentially
  all of the residual maintain off-support and is removed only by a portfolio
  (histogram-of-tanks) state, not by re-timing the stamp.
- **Replacement off-support (an open data-support question).** $F_{\text{R}}$
  resets a replacing facility to age bin 1, but a one-tank swap at a multi-tank
  facility does not reset the portfolio average. Beginning-of-year stamping
  *un-masks* the true magnitude of this mismatch (the end-of-year coding had
  understated it). Whether the data can identify a richer scale model of the
  partial-closure and replacement margins --- two-part costs, a size gradient in
  exit value --- is an open identification question, examined in the
  scale-incorporation scoping (Ticket 014), not a defect to be patched here.

## Transition realization versus the model's $F$ {#sec-audit-trans}

The model's entire dynamic structure is the transition matrices $F_{\text{M}}$
(maintain: age advances within wall and regime) and $F_{\text{R}}$ (replace:
reset to age bin 1). Any state-coding error appears as empirical transition mass
where the model assigns zero probability (``off-support mass'').

```{r audit-d1}
#| label: tbl-audit-d1
#| tbl-cap: "Transition realization versus the model's F matrices: off-support mass and model-forbidden transitions."
.f <- here::here("Output","Tables","T013_D1_TransitionRealization_Summary.csv")
if (file.exists(.f)) {
  dt <- data.table::fread(.f)
  knitr::kable(dt, format = "latex", booktabs = TRUE,
    col.names = c("Diagnostic", "Value")) |>
    kableExtra::kable_styling(font_size = 9, latex_options = "hold_position")
} else { cat("*[run T013 to generate T013\\_D1\\_TransitionRealization\\_Summary.csv]*") }
```

\vspace{0.1cm}
\begin{flushleft}\footnotesize
\textit{Notes:} Empirical $(s_t,a_t,s_{t+1})$ transitions from the
beginning-of-year-stamped estimation panel compared to the model's $F$.
``Off-support mass'' is the share of transitions landing where $F=0$. After the
stamping fix, the maintain off-support mass is 0.69\%, of which the wall
component (DW$\rightarrow$SW plus SW$\rightarrow$DW) is only 0.09\% --- the
remainder are backward age moves: a partial closure retiring an old tank lowers
the portfolio \emph{average} age, which the monotone $F_{\text{M}}$ kernel cannot
represent. Replace off-support is 98\%: $F_{\text{R}}$ sends every replace to age
bin 1, but a one-tank swap at a multi-tank facility does not reset the facility's
average age (beginning-of-year stamping \emph{un-masks} the true magnitude, which
the earlier end-of-year coding had understated at 30\%). Regime never changes and
exit is never followed by re-entry, confirming those two invariants hold. Both
residuals are structural limitations of the single-tank state abstraction, not
coding bugs, and are the subject of the portfolio (scale-incorporation) extension.
\end{flushleft}

```{r audit-d1-fig}
#| label: fig-audit-d1
#| fig-cap: "Model-forbidden Maintain transitions remaining after beginning-of-year stamping, by frequency. These are now dominated by backward age moves --- partial closures pruning old tanks from a portfolio average --- the residual single-tank-abstraction limitation the portfolio model resolves."
.f <- here::here("Output","Figures","T013_D1_TopOffSupportFlows.png")
if (file.exists(.f)) knitr::include_graphics(.f) else
  cat("*[run T013 to generate T013\\_D1\\_TopOffSupportFlows.png]*")
```

\vspace{0.1cm}
\begin{flushleft}\footnotesize
\textit{Notes:} The twelve most frequent model-forbidden Maintain transitions
\emph{after} beginning-of-year stamping. States are labelled age/wall/regime.
They are now overwhelmingly backward age transitions within the same wall (e.g.\
$\text{A}3\rightarrow\text{A}1$): $A_{\text{bin}}$ is a portfolio \emph{average}
that falls when an old tank is pruned by a partial closure, which the monotone
maintain kernel cannot represent. Wall-flip transitions (DW$\rightarrow$SW),
which the end-of-year coding produced in bulk, are now a negligible residual ---
the leakage the stamping fix removed.
\end{flushleft}

## Closure cell-assignment leakage {#sec-audit-leak}

A closure's wall cell should be the firm's decision-time (beginning-of-year)
wall. @tbl-audit-routing tracks, for closures grouped by their beginning-of-year
wall composition, which cell the end-of-year coding actually assigns them to.

```{r audit-routing}
#| label: tbl-audit-routing
#| tbl-cap: "Closure routing: decision-time (beginning-of-year) wall composition versus the assigned (end-of-year) wall cell."
.f <- here::here("Output","Tables","T012_C4b_ClosureRouting.csv")
if (file.exists(.f)) {
  dt <- data.table::fread(.f)
  knitr::kable(dt, format = "latex", booktabs = TRUE,
    col.names = c("BOY wall composition", "Action", "Assigned SW cell", "Assigned DW cell")) |>
    kableExtra::kable_styling(font_size = 9, latex_options = "hold_position")
} else { cat("*[run T012 to generate T012\\_C4b\\_ClosureRouting.csv]*") }
```

\vspace{0.1cm}
\begin{flushleft}\footnotesize
\textit{Notes:} Closures grouped by beginning-of-year wall composition (rows) and
tabulated by the assigned wall cell (columns), on the corrected panel. Each
closure is now filed under its decision-time wall: all-DW closures route to the
DW column, all-SW and mixed closures to the SW column, so the off-diagonal is
zero. This is the fix. Under the prior end-of-year coding the routing was
inverted --- every all-DW \emph{exit} was filed in an SW cell (a DW facility has
zero tanks at year-end) and SW-origin \emph{replaces} (single- to double-wall
retrofits) were filed in DW cells --- so 16.2\% of all closures landed in the
wrong decision-time wall: the SW exit margin (identifying $\kappa_{\text{SW}}$)
was 11.0\% contaminated by DW-origin exits, and only 16.6\% of the DW replace
cell (identifying $K_{\text{DW}}$) was genuine all-DW behavior, the remaining
83.4\% being mis-celled SW$\rightarrow$DW retrofits. Beginning-of-year stamping
(Ticket 013) eliminates this leakage; what remains (the backward-age and
replacement off-support of @tbl-audit-d1) is the single-tank \emph{modelling}
limitation, not a cell-assignment error.
\end{flushleft}

## Facility size: stickiness and within-cell variation {#sec-audit-size}

Whether facility size belongs in the state space turns on two facts: is size
sticky enough to be a fixed type (so no transition kernel is needed), and does it
shift choice probabilities *within* an existing cell (so it carries information
the 32-cell state omits).

```{r audit-size-fig}
#| label: fig-audit-size
#| fig-cap: "Facility size transition matrix (end-of-year tank stock, active-to-active facility-years). Near-diagonal mass indicates size is a sticky, near-permanent attribute."
.f <- here::here("Output","Figures","T012_C1_SizeTransition_Heatmap.png")
if (file.exists(.f)) knitr::include_graphics(.f) else
  cat("*[run T012 to generate T012\\_C1\\_SizeTransition\\_Heatmap.png]*")
```

\vspace{0.1cm}
\begin{flushleft}\footnotesize
\textit{Notes:} Year-over-year transition of the facility size bin (tank count),
among facility-years active in both years. Diagonal probabilities are
0.984--0.995 and only 1.17\% of active pairs change bin, so facility size is
well approximated as a fixed (time-invariant) facility type --- it can enter the
state space without a size-transition kernel. The complementary check (within-cell
variation in choice probabilities across size bins, `T012_C2_WithinCell_Spread.csv`)
finds an observation-weighted mean spread in $P(\text{close})$ of 0.019 and a
maximum of 0.082, concentrated in the risk-based cells, so size carries
information the 32-cell state omits. Together these motivate adding size as a
fixed-type state dimension (Tickets 014--015).
\end{flushleft}

The one-year transition above *understates* what a fixed type requires: the size
bin must be constant over a facility's entire operating spell, and a 99\% one-year
stay rate compounds over a 16-year mean spell. @fig-fixedtype-spell and
@tbl-fixedtype make that the test.

```{r fig-fixedtype-spell}
#| label: fig-fixedtype-spell
#| fig-cap: "Fraction of facilities still in their original size bin, by years since first observed: tank count versus total capacity. Capacity stays higher --- consolidation drops a tank but preserves capacity --- so capacity is the better fixed-type size state."
.f <- here::here("Output","Figures","04ag_FixedType_Survival.png")
if (file.exists(.f)) knitr::include_graphics(.f) else
  cat("*[run 04ag\\_Size\\_Stickiness\\_Diagnostic.R]*")
```

```{r tbl-fixedtype}
#| label: tbl-fixedtype
#| tbl-cap: "Fixed-type test: one-year stay rate versus spell-level constancy, for tank count and total capacity."
.t <- data.table::fread(here::here("Output","Tables","04ag_FixedType_Summary.csv"))
.t <- .t[, .(measure,
             one_year_stay        = sprintf("%.1f%%", 100 * one_year_stay),
             implied_if_iid       = sprintf("%.1f%%", 100 * implied_if_iid),
             actual_never_changed = sprintf("%.1f%%", 100 * actual_never_changed))]
knitr::kable(.t, booktabs = TRUE, escape = TRUE,
  col.names = c("Size measure", "One-year stay", "Implied over spell", "Actual never changed"))
```

\vspace{0.1cm}
\begin{flushleft}\footnotesize
\textit{Notes:} A fixed type must be constant over the facility's whole operating
spell (mean 16.4 years), so the one-year stay rate ($\sim$\,99.5\% for both count
and capacity) is not the relevant statistic --- it compounds. The spell-level test: 91.8\% of
facilities never change their tank-count bin and 93.5\% never change their
capacity quartile over their full spell (close to, and slightly above, the
i.i.d.-compounded benchmark --- i.e.\ mild extra persistence). Of the 0.55\% of
operating-year transitions that do change the count bin, 70\% are reductions
(shrinking/consolidation) and 100\% coincide with a coded install or closure, so
size never drifts without an action: a fixed-type state with action-driven resize
transitions loses no within-spell variation. Stickiness weakens with size
(single-tank $\sim$\,98\%, four-plus-tank $\sim$\,87\% by count but $\sim$\,89--94\%
by capacity), so the approximation is tightest where most facility-years sit and
loosest at large facilities --- where capacity holds up markedly better than count.
With the within-cell $P(\text{close})$ spread above, this supports \textbf{total
capacity} as the fixed-type size state. Source:
\texttt{04ag\_Size\_Stickiness\_Diagnostic.R}.
\end{flushleft}

Stickiness says size does not *drift*; the complementary question is whether one
representative tank is a faithful *summary* of the portfolio at a point in time ---
i.e.\ whether facilities are internally homogeneous on the risk dimensions.
@fig-homog-size shows they largely are: across all facility-years, 81.5\% have all
active tanks in a single wall type and a single five-year age bin (the exact support
of the representative-tank state), and wall homogeneity exceeds 88\% at every size.
The residual heterogeneity is an *age-spread* phenomenon concentrated at large
facilities --- the both-homogeneous share falls from 93\% (single-tank) to 62\%
(four-plus) --- so the single-representative-tank abstraction is exact for the bulk
of facility-years and degrades only in the large-facility tail, exactly where the
portfolio/histogram extension (or an oldest-two-ages summary state) earns its keep.

```{r fig-homog-size}
#| label: fig-homog-size
#| fig-cap: "Within-facility homogeneity by facility size: share of facility-years whose active tanks share one wall type, one five-year age bin, or both. Single-tank facilities are homogeneous by construction; heterogeneity is an age-spread phenomenon that emerges only at four-plus-tank facilities."
.f <- here::here("Output","Figures","04ah_Homogeneity_bySize.png")
if (file.exists(.f)) knitr::include_graphics(.f) else
  cat("*[run 04ah\\_Within\\_Facility\\_Homogeneity.R]*")
```

\vspace{0.1cm}
\begin{flushleft}\footnotesize
\textit{Notes:} Facility-year weighted, estimation sample. ``Age one bin'' is the
share whose oldest and youngest active tank fall in the same five-year state bin;
``Both'' adds the single-wall requirement and is the share the 32-cell
representative-tank state represents exactly. The median age range is zero at every
size --- most multi-tank facilities install their tanks together --- so heterogeneity
lives in the tail, not the typical facility. Single-tank facilities (23.2\% of
facility-years) are homogeneous by construction; the $\sim$\,7\% of single-tank
facility-years flagged not wall-homogeneous carry an unknown wall code, not genuine
mixing. Source: \texttt{04ah\_Within\_Facility\_Homogeneity.R}.
\end{flushleft}

Where a facility *is* internally heterogeneous (the large-facility tail), the
representative-tank abstraction is rescued by a further regularity: the tank that
gets acted on is *predictable*. @fig-which-tank shows that when a multi-tank facility
closes exactly one tank, that tank is the single-walled one in 79\% of mixed-wall
facilities (versus a 52\% base rate), the smallest-capacity tank 70\% of the time
(versus 32\% under a random draw), and the oldest 74\% of the time (versus 32\%) ---
the shed tank is the small, single-walled, older one, not a random draw. This is the
capacity-preserving consolidation margin (Section\ \ref{sec-audit-trans}) seen from
the tank side, and it means the heterogeneous tail can be carried by a marginal-tank
representation keyed on wall and size with a step-down transition, rather than the
full portfolio histogram.

```{r fig-which-tank}
#| label: fig-which-tank
#| fig-cap: "When a multi-tank facility closes exactly one tank, the kind of tank it sheds, observed versus a random-tank benchmark. The shed tank is disproportionately single-walled, smallest-capacity, and (more weakly) oldest."
.f <- here::here("Output","Figures","04ai_WhichTank_Shed.png")
if (file.exists(.f)) knitr::include_graphics(.f) else
  cat("*[run 04ai\\_Which\\_Tank\\_Diagnostic.R]*")
```

\vspace{0.1cm}
\begin{flushleft}\footnotesize
\textit{Notes:} Tank-level panel (\texttt{panel\_dt}); single-closure events in
multi-tank facility-years ($n=19{,}744$). ``If random'' is the share expected if the
closed tank were drawn uniformly --- the base single-wall share in mixed facilities
for the wall row, and $\mathbb{E}[1/n]$ for the oldest and smallest rows. The age
signal is weaker than wall and size because most facilities install their tanks
together, so the oldest tank is frequently tied. Source:
\texttt{04ai\_Which\_Tank\_Diagnostic.R}.
\end{flushleft}
