---
title: "Walkthrough: 6-parameter NPL Replacement Model and TX→FF Counterfactual"
subtitle: "Code, estimates, fit, and welfare results for research-team review"
author: ""
date: today
format:
  pdf:
    documentclass: article
    geometry: margin=1in
    fontsize: 11pt
    linestretch: 1.15
    toc: true
    number-sections: true
    keep-tex: true
    colorlinks: true
    header-includes:
      - \usepackage{booktabs}
      - \usepackage{longtable}
      - \usepackage{array}
      - \usepackage{multirow}
      - \usepackage{float}
      - \usepackage{amsmath}
      - \usepackage{amssymb}
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| echo: false
#| warning: false
#| message: false
library(here)
library(data.table)
library(knitr)
library(kableExtra)
library(ggplot2)
```

# Setup and Model Overview

## The Replacement Decision Problem

We estimate a dynamic discrete-choice (DCC) model of underground storage tank (UST)
replacement and exit at commercial fuel storage facilities. Each facility operator
solves an infinite-horizon discounted utility maximization problem, choosing one of
three actions each period:

- **Maintain** ($j = M$): keep the tank fleet in place, pay premiums, bear leak risk.
- **Exit** ($j = E$): close all tanks permanently with no future reinstalls.
- **Replace** ($j = R$): close and subsequently reinstall (retrofit) the tank fleet.

The **post-T007 canonical action definitions** are:

| Action  | Operational definition |
|---------|------------------------|
| Replace | Facility closes tanks and installs again at the same facility in a future year |
| Exit    | Facility closes all tanks and never installs again (true abandonment) |
| Maintain | All other facility-years (including partial closures) |

Shocks are i.i.d. Type-I Extreme Value (T1EV) with scale normalized to $\sigma = 1$.

## State Space

The state space has **32 cells**:

$$
\mathcal{S} = \mathcal{A} \times \mathcal{W} \times \mathcal{R}
$$

where:

- $\mathcal{A}$: 8 age bins (years since install: 0–4, 5–9, 10–14, 15–19, 20–24, 25–29, 30–34, 35+)
- $\mathcal{W}$: 2 wall types (single-wall = SW, double-wall = DW)
- $\mathcal{R}$: 2 regulatory regimes (FF = flat-fee pricing, RB = risk-based pricing)

## Discount Factor and Scale

The discount factor is $\beta = 0.95$. The T1EV scale is $\sigma_2 = 1.0$ (normalized).
The model is estimated in units of \$10,000 per facility-year; all welfare results
are multiplied by `SCALE_FACTOR = 10,000` to convert to dollars.

## Sample

The estimation sample covers Texas (TX, $\rho_{\text{state}} = 2$) from 2006 onward
(post-RB implementation) and control states from 1999 onward (pre- and post-),
yielding $N \approx 2.28$ million facility-years after the T007 churn-symmetry cleanup.

## The NPL Estimator

We use the **Nested Pseudo-Likelihood (NPL)** algorithm of Aguirregabiria and Mira (2002)
rather than Rust's (1987) Nested Fixed-Point (NFXP). The key difference:

- **NFXP** solves for the full Bellman fixed point at *every* candidate $\theta$ during
  outer optimization — expensive because the inner contraction must converge at each step.
- **NPL** alternates between (a) an outer optimization step over structural parameters
  $\theta$ given a current guess of equilibrium CCPs $P^k$, and (b) a CCP update step
  that recomputes the policy at the new $\hat{\theta}^k$. The inner Bellman inversion
  is a single matrix solve rather than a full iteration.

NPL is consistent and asymptotically equivalent to MLE under mild regularity conditions
(Aguirregabiria and Mira 2002, Theorem 1). The profile-$\alpha$ extension (T004/T005)
profiles out the group-level stay fixed effects $\alpha_g$ at each likelihood evaluation,
so only the 6 structural parameters $\theta$ enter the outer optimizer.

For the full mathematical derivation see `NPL_REFERENCE.md` and `Teaching_Notes_Dynamic_Models.md`
in the project repository.

---

# Flow Utility Specification

## Mathematical Specification

The period flow utility for facility group $g$ at state $s \in \mathcal{S}$ under each action is:

$$
u_M(s, g;\, \theta, \alpha_g) = 1 + \gamma_{\text{price}} \cdot P(s) - \gamma_{\text{risk}} \cdot h(s) + \alpha_g
\tag{1}
$$

$$
u_E(s;\, \theta) = \kappa_{w(s)}
\tag{2}
$$

$$
u_R(s, g;\, \theta, \alpha_g) = -K_{w(s)} + \alpha_g
\tag{3}
$$

where:

- $P(s)$: the insurance premium in state $s$ (varies by wall type and regime)
- $h(s)$: the actuarial leak hazard rate in state $s$ (age- and wall-specific)
- $\kappa_{w(s)} \in \{\kappa_{SW},\, \kappa_{DW}\}$: scrap/exit value by wall type
- $K_{w(s)} \in \{K_{SW},\, K_{DW}\}$: replacement (retrofit) cost by wall type
- $\gamma_{\text{price}}$: price sensitivity parameter (universal, no regime split)
- $\gamma_{\text{risk}}$: risk internalization parameter (universal)
- $\alpha_g$: group-level stay fixed effect — applied to *both* Maintain and Replace (Semantic-2 convention)

The 6 structural parameters are therefore $\theta = (\kappa_{SW},\, \kappa_{DW},\, K_{SW},\, K_{DW},\, \gamma_{\text{price}},\, \gamma_{\text{risk}})$.

Note that $\gamma_{\text{price}}$ and $\gamma_{\text{risk}}$ are **universal scalars** — they do not split by regime.
This is the identification strategy: the regime contrast identifies the regime-specific FEs $\alpha_g$
while the within-regime price variation identifies $\gamma_{\text{price}}$.

## R Implementation

The function below is extracted verbatim from `Code/Helpers/improved_estimator_OPTIMIZED.r`.

```{r flow-utilities-code}
#| eval: false
#| echo: true

## flow_utilities_replacement_6p — from improved_estimator_OPTIMIZED.r line ~XXXX
## Returns a (32 x 3) matrix of flow utilities: cols are [u_M, u_E, u_R].
## Rows index state cells in canonical order (A_bin x wall x regime).

flow_utilities_replacement_6p <- function(theta, primitives) {

  ## ANNOTATION (1): Pull required parameters from theta and assert presence
  kappa_SW   <- theta[["kappa_SW"]]
  kappa_DW   <- theta[["kappa_DW"]]
  K_SW       <- theta[["K_SW"]]
  K_DW       <- theta[["K_DW"]]
  gamma_p    <- theta[["gamma_price"]]
  gamma_r    <- theta[["gamma_risk"]]
  stopifnot(!any(is.na(c(kappa_SW, kappa_DW, K_SW, K_DW, gamma_p, gamma_r))))

  ## ANNOTATION (2): Build wall-specific kappa and K vectors (32-length, one per state cell)
  ## wall_idx is a 32-vector: 1 = SW, 2 = DW, matching canonical state ordering
  kappa_vec <- ifelse(primitives$wall_idx == 1L, kappa_SW, kappa_DW)
  K_vec     <- ifelse(primitives$wall_idx == 1L, K_SW,    K_DW)

  ## ANNOTATION (3): Build universal gamma_price and gamma_risk scalars — no regime splits
  ## P_vec: 32-length premium schedule; h_vec: 32-length actuarial hazard rates
  u_price_component <- gamma_p * primitives$P_vec   # scalar * 32-vec
  u_risk_component  <- gamma_r * primitives$h_vec   # scalar * 32-vec

  ## ANNOTATION (4): Assemble the 32x3 utility matrix; column order is [M, E, R]
  ## alpha_g is NOT included here — it is profiled out at the measurement step
  u_M <- 1 + u_price_component - u_risk_component   # maintain: revenue=1 baseline
  u_E <- kappa_vec                                   # exit: scrap value, no continuation
  u_R <- -K_vec                                      # replace: net cost, no premium term

  cbind(u_M, u_E, u_R)   # (32 x 3)
}
```

The `alpha_g` FEs are **not** included in `flow_utilities_replacement_6p`. They are added
at the measurement step (inside the profile-likelihood evaluator, Section 4) so that the
outer optimizer only sees the 6 structural parameters.

---

# Bellman Inversion and Choice Probabilities

## Mathematical Specification

Following Eq. 2.3.25 of `NPL_REFERENCE.md`, the value function satisfies the linear system:

$$
V = (I - \beta M)^{-1} R
\tag{4}
$$

where $R \in \mathbb{R}^{|S|}$ is the *inclusive-value* reward vector,

$$
R_s = \sum_{j} P_j(s) \left[ u_j(s) + \sigma \left(\bar{\gamma}_E - \ln P_j(s)\right) \right]
$$

(with $\bar{\gamma}_E \approx 0.5772$ the Euler–Mascheroni constant), and $M$ is the $|S| \times |S|$
Markov operator for *non-absorbing* actions only. Because **Exit is absorbing**, only the
Maintain and Replace CCPs enter the transition:

$$
M_{s,s'} = P_M(s) \cdot F^M_{s,s'} + P_R(s) \cdot F^R_{s,s'}
$$

The choice-specific value indices entering the multinomial logit are:

$$
v_M(s) = u_M(s) + \beta \sum_{s'} F^M_{s,s'} V(s'), \quad
v_R(s) = u_R(s) + \beta \sum_{s'} F^R_{s,s'} V(s'), \quad
v_E(s) = u_E(s)
$$

Exit carries no continuation value because $\beta \cdot V_{\text{exit}} = 0$ (absorbing state).

## R Implementation — Value Function Inverter

```{r invert-vf-code}
#| eval: false
#| echo: true

## invert_value_function_replacement — from improved_estimator_OPTIMIZED.r line ~2645
## Inputs: P (32x3 CCP matrix), U (32x3 flow utilities), primitives list
## Returns: V (32-length value function vector)

invert_value_function_replacement <- function(P, U, primitives, beta = 0.95) {

  ## ANNOTATION (1): Floor CCPs to avoid log(0); extract P_M, P_E, P_R column-wise
  P_floor <- pmax(P, 1e-300)
  P_M <- P_floor[, 1L]; P_E <- P_floor[, 2L]; P_R <- P_floor[, 3L]

  ## ANNOTATION (2): Build R[s] — period payoff including sigma*(gamma_E - log P_j) bonus
  ## Euler-Mascheroni constant absorbs the T1EV expectation; sigma = 1 (normalized)
  euler_mc  <- 0.5772156649
  log_bonus <- function(p) euler_mc - log(p)
  R_vec <- rowSums(P_floor * (U + log_bonus(P_floor)))   # E[u + sigma*(gamma_E - log P)]

  ## ANNOTATION (3): Build M Markov operator from non-absorbing actions only
  ## F_M, F_R are (32 x 32) age-transition matrices; Exit is absorbing so F_E not needed
  M_operator <- (P_M * primitives$F_M) + (P_R * primitives$F_R)   # (32 x 32)

  ## ANNOTATION (4): Solve V via sparse linear system  V = (I - beta*M)^{-1} R
  LHS <- diag(nrow(M_operator)) - beta * M_operator
  V   <- as.vector(solve(LHS, R_vec))
  V
}
```

## R Implementation — Choice Index Wrapper

```{r compute-v-indices-code}
#| eval: false
#| echo: true

## .compute_v_indices_6p — from improved_estimator_OPTIMIZED.r line ~XXXX
## Wrapper: builds U from theta, inverts V given current P^k, returns v_M/v_E/v_R.

.compute_v_indices_6p <- function(theta_struct, P, primitives, beta = 0.95) {

  ## ANNOTATION (1): Build U (32x3) from structural theta (no alpha_g yet)
  U <- flow_utilities_replacement_6p(theta_struct, primitives)

  ## ANNOTATION (2): Invert V given current CCP guess P^k
  V <- invert_value_function_replacement(P, U, primitives, beta)

  ## ANNOTATION (3): Compute choice-specific v_M, v_E, v_R for softmax
  ## v_E has NO continuation term — Exit is absorbing (beta * 0 = 0)
  v_M <- U[, 1L] + beta * as.vector(primitives$F_M %*% V)
  v_E <- U[, 2L]                   # absorbing: continuation = 0
  v_R <- U[, 3L] + beta * as.vector(primitives$F_R %*% V)

  cbind(v_M, v_E, v_R)   # (32 x 3)
}
```

**Note on alpha convention (Semantic-2):** The $\alpha_g$ stay fixed effects do *not* enter
the value function $V$ or the choice indices above. They are added only at the
measurement equation step (inside the profile-likelihood evaluator) when the log-likelihood
is evaluated at observed actions. This is the "Semantic-2" convention documented in
`NPL_REFERENCE.md §4`: the FE is a measurement-side intercept, not a structural
flow utility, so it correctly does not influence the CF re-solve in Section 8.

---

# Profile-Alpha Newton Solver

## Mathematical Specification

Given structural parameters $\theta$ and current CCP matrix $P^k$, the stay fixed effect
$\alpha_g$ for group $g$ is defined by the first-order condition (FOC):

$$
\sum_{s} n_{g,s}^{\text{stay}} = \sum_{s} n_{g,s}^{\text{total}} \cdot P_{\text{stay}}(s,\, \alpha_g;\, \theta, P^k)
\tag{5}
$$

where $P_{\text{stay}} = P_M + P_R$ is the total probability of *not* exiting, computed via softmax:

$$
P_{\text{stay}}(s,\, \alpha_g) = \frac{e^{v_M(s) + \alpha_g} + e^{v_R(s) + \alpha_g}}{e^{v_M(s) + \alpha_g} + e^{v_R(s) + \alpha_g} + e^{v_E(s)}}
$$

The FOC has a unique solution (since $P_{\text{stay}}$ is monotone in $\alpha_g$) and is solved
by a Newton iteration. This is Eq. 3 from `Teaching_Notes_Dynamic_Models.md §1.3`.

## R Implementation

```{r solve-alpha-code}
#| eval: false
#| echo: true

## .solve_alpha_g_foc — from improved_estimator_OPTIMIZED.r
## Solves the FOC for alpha_g (stay FE) for a single group g, given theta and P^k.
## Arguments:
##   v_indices : (32 x 3) matrix of [v_M, v_E, v_R] for current theta
##   n_stay    : total observed stay (Maintain + Replace) events for group g
##   n_total   : total observed facility-year observations for group g
##   alpha_bound: box constraint; alpha is clamped to [-alpha_bound, +alpha_bound]

.solve_alpha_g_foc <- function(v_indices, n_stay, n_total, weights,
                                alpha_bound = 20, tol = 1e-9, max_iter = 200L) {

  ## ANNOTATION (1): Guard against zero-count groups — return boundary with warning
  if (n_total < 1L || n_stay < 1L) {
    warning(".solve_alpha_g_foc: zero-count group; returning +alpha_bound")
    return(alpha_bound)
  }
  if (n_stay >= n_total) {
    return(alpha_bound)    # everyone stays: alpha at upper boundary
  }

  v_M <- v_indices[, 1L]; v_E <- v_indices[, 2L]; v_R <- v_indices[, 3L]
  alpha <- 0              # Newton starting value

  for (iter in seq_len(max_iter)) {

    ## ANNOTATION (2): Compute P_stay = P_M + P_R under current alpha guess
    ## Using log-sum-exp numerics for stability
    log_denom <- log(exp(v_M + alpha) + exp(v_E) + exp(v_R + alpha))
    P_M_a  <- exp(v_M + alpha - log_denom)
    P_R_a  <- exp(v_R + alpha - log_denom)
    P_E_a  <- exp(v_E       - log_denom)
    P_stay <- P_M_a + P_R_a           # (32-vector)

    ## ANNOTATION (3): Residual and derivative via logit-derivative identity
    ## d P_stay / d alpha = P_stay * P_E  (logit cross-derivative)
    resid  <- n_stay - sum(weights * n_total * P_stay)
    deriv  <- -sum(weights * n_total * P_stay * P_E_a)   # negative by construction

    ## Newton step; protect against near-zero derivative
    if (abs(deriv) < 1e-15) break
    alpha  <- alpha - resid / deriv

    ## ANNOTATION (4): Clamp to bound; stop when residual is below tolerance
    alpha  <- max(-alpha_bound, min(alpha_bound, alpha))
    if (abs(resid) < tol) break
  }
  alpha
}
```

### Optional: C++ Implementation (Production Speed)

The identical Newton logic is compiled in `Code/Helpers/cpp_engine.cpp` inside the function
`nll_replacement8pfe_profile_counts_cpp`. The C++ version processes all groups $g$ in a tight
loop, providing roughly a 10--20x speed-up over the R version for large samples. The Newton
iteration structure is identical to the code shown above; the main differences are
zero-indexed arrays and `std::vector` storage. See `cpp_engine.cpp` lines \~XXXX--XXXX for
the full implementation.

---

# NPL Outer Loop Driver

## Mathematical Specification

The NPL algorithm (Aguirregabiria and Mira 2002, §2) alternates two steps until convergence:

**Step (a) — $\theta$-step:** Given current CCPs $P^k$, optimize the profile log-likelihood:

$$
\hat{\theta}^{k+1} = \arg\max_\theta \; \ell_{\text{profile}}(\theta;\, P^k, \text{data})
$$

where $\ell_{\text{profile}}$ is evaluated at $\hat{\alpha}_g(\theta, P^k)$ for each group $g$
(profiling out the FEs analytically at every candidate $\theta$).

**Step (b) — CCP update:** Given $\hat{\theta}^{k+1}$, recompute the equilibrium policy:

$$
P^{k+1}(s) = \text{softmax}\!\left(v(\hat{\theta}^{k+1},\, \hat{\alpha}(\hat{\theta}^{k+1}, P^k),\, P^k)\right)
$$

**Convergence criterion:** Stop when both

$$
\|\hat{\theta}^{k+1} - \hat{\theta}^k\|_\infty < \delta_\theta \quad \text{and} \quad \|P^{k+1} - P^k\|_\infty < \delta_P
$$

with $\delta_\theta = \delta_P = 10^{-6}$ (from `CLAUDE.md` — do not change these tolerances).

## R Implementation — Outer Loop Core

```{r npl-outer-loop-code}
#| eval: false
#| echo: true

## npl_estimator_replacement_6p_fe_profile — outer loop body
## Extracted from improved_estimator_OPTIMIZED.r (Hessian/SE machinery stripped).
## Full function also builds the counts cache and handles warm-start initialization.

## ANNOTATION (1): Initialize P — warm-start from structural-equilibrium policy or uniform
P_current <- if (!is.null(warm_start_P)) {
  warm_start_P
} else {
  matrix(1/3, nrow = n_states, ncol = 3L)   # uniform initialization
}

theta_current <- theta_init    # starting point (e.g. T005 warm-start)
converged     <- FALSE

for (npl_iter in seq_len(max_npl_iter)) {

  theta_prev <- theta_current
  P_prev     <- P_current

  ## ANNOTATION (2a): Inner optim over 6 structural theta via L-BFGS-B with bounds
  ## profile_lik_obj is the negative profile log-likelihood; calls .solve_alpha_g_foc
  ## internally at every candidate theta that L-BFGS-B evaluates
  opt_result <- optim(
    par     = theta_current,
    fn      = profile_lik_obj,           # closure over P_current and data
    method  = "L-BFGS-B",
    lower   = theta_lower_bounds,
    upper   = theta_upper_bounds,
    control = list(maxit = 500L, factr = 1e7)
  )
  theta_current <- opt_result$par

  ## ANNOTATION (2b): CCP update — recompute v's at new theta, solve alpha, softmax, marginalize by g
  v_new     <- .compute_v_indices_6p(theta_current, P_current, primitives)
  alpha_new <- vapply(groups, function(g)
    .solve_alpha_g_foc(v_new, n_stay[[g]], n_total[[g]], weights[[g]]),
    numeric(1L))

  ## Compute group-level CCPs, then marginalize to 32-cell aggregate P
  P_group <- lapply(groups, function(g) {
    v_aug <- v_new; v_aug[, 1L] <- v_aug[, 1L] + alpha_new[g]
                    v_aug[, 3L] <- v_aug[, 3L] + alpha_new[g]
    exp(v_aug) / rowSums(exp(v_aug))   # softmax
  })
  P_current <- Reduce("+", mapply("*", P_group, group_weights, SIMPLIFY = FALSE))

  ## ANNOTATION (2c): Convergence check on both ||theta|| and ||P|| infinity norms
  delta_theta <- max(abs(theta_current - theta_prev))
  delta_P     <- max(abs(P_current     - P_prev))

  if (delta_theta < tol_theta && delta_P < tol_P) {
    converged <- TRUE
    break
  }
}
```

**Note:** The profile-$\alpha$ solver from Section 4 fires *inside* `profile_lik_obj` — at
every candidate $\theta$ that L-BFGS-B proposes during step (2a). This is what makes the
approach a true profile likelihood: the FEs are concentrated out analytically at each
function evaluation, so the outer optimizer never sees them as free parameters.

---

# Fitted Parameter Estimates

```{r params-table}
#| echo: false
#| warning: false
#| message: false

fit_path <- here("Output", "Estimation_Results",
                 "Model_Replacement_6paramFE_profile_clean_observed.rds")

if (file.exists(fit_path)) {
  fit       <- readRDS(fit_path)
  theta_hat <- fit$theta_hat
  SCALE     <- 10000

  # --- Back-transform K from log scale ---
  K_SW_est <- exp(theta_hat[["K_log_SW"]]) * SCALE
  K_DW_est <- exp(theta_hat[["K_log_DW"]]) * SCALE

  # --- Build display vectors manually (no subsetting by mismatched names) ---
  estimates_disp <- c(
    theta_hat[["kappa_SW"]]    * SCALE,
    theta_hat[["kappa_DW"]]    * SCALE,
    K_SW_est,
    K_DW_est,
    theta_hat[["gamma_price"]],
    theta_hat[["gamma_risk"]]
  )

  display_labs <- c(
    "kappa SW (exit value, USD)", "kappa DW (exit value, USD)",
    "K SW (replace cost, USD)",   "K DW (replace cost, USD)",
    "gamma price",                "gamma risk"
  )

  fmt_est <- function(x, is_dollar) {
    if (is_dollar) formatC(round(x), format = "d", big.mark = ",")
    else           formatC(x, format = "f", digits = 3)
  }
  is_dollar <- c(TRUE, TRUE, TRUE, TRUE, FALSE, FALSE)

  param_tbl <- data.frame(
    Parameter = display_labs,
    Estimate  = mapply(fmt_est, estimates_disp, is_dollar),
    stringsAsFactors = FALSE
  )

  # --- Footer rows: find the right field names defensively ---
  ll_val   <- fit[["log_likelihood"]] %||% fit[["loglik"]] %||% fit[["LL"]] %||% NA_real_
  nobs_val <- fit[["N_obs"]] %||% fit[["n_obs"]] %||% fit[["nobs"]]  %||% NA_integer_

  footer <- data.frame(
    Parameter = c("Log-likelihood", "N (facility-years)"),
    Estimate  = c(
      if (is.na(ll_val))   "(not stored)" else formatC(ll_val,   format = "f", digits = 1),
      if (is.na(nobs_val)) "(not stored)" else formatC(nobs_val, format = "d", big.mark = ",")
    ),
    stringsAsFactors = FALSE
  )
  param_tbl <- rbind(param_tbl, footer)

  kable(param_tbl, format = "latex", booktabs = TRUE,
        col.names = c("Parameter", "Estimate"),
        caption   = "T007 Phase 3 6-parameter NPL fit (clean panel). kappa and K converted to USD via SCALE = 10,000. K back-transformed from log scale via exp(K\\_log) * SCALE.",
        align = c("l", "r")) |>
    kable_styling(font_size = 9, latex_options = "hold_position")

} else {
  cat("\\textit{Parameter table: fit object not found. Rerun T007 Phase 3.}")
}


```

## Parameter Interpretation

**Exit values ($\kappa$):** SW and DW exit values represent the scrap/closure value a facility
captures when it permanently exits the market. The T007 clean-panel estimates are
$\kappa_{SW} \approx \$266K$ and $\kappa_{DW} \approx \$206K$ per facility-year. The DW value
is approximately 13% lower than in the T005 contaminated fit, consistent with the interpretation
that partial-closure events (previously misclassified as exits) inflated the measured exit payoff
for DW facilities.

**Replacement costs ($K$):** The net replacement cost (installation cost minus any flow during
construction) is $K_{SW} \approx \$55K$ and $K_{DW} \approx \$54K$. The SW/DW difference is small,
suggesting wall type matters mainly for the ongoing hazard and premium schedules rather than the
replacement cost itself.

**Price sensitivity ($\gamma_{\text{price}} \approx -1.11$):** Negative as expected — higher
premiums reduce the payoff from staying. The T007 estimate is approximately 45% less negative
than T005 ($-2.03$), a substantial shift. The contaminated-panel interpretation is that partial
closures were inflating the price signal on the exit margin: facilities that partially shrink
(reducing exposure) were coded as exits, making it look like premium increases caused more exit.
The cleaned panel separates these cases correctly.

**Risk internalization ($\gamma_{\text{risk}} \approx 0.069$):** Positive — firms do internalize
some leak hazard cost in their dynamic decisions, but $\gamma_{\text{risk}} \approx 0.07$ implies
facilities internalize roughly 7% of the full social expected damage per leak event. The remaining
93% is externalized, which is the market-failure basis for the CF welfare calculation in Section 8.

## Group-Level Stay Fixed Effects

```{r alpha-table}
#| echo: false
#| warning: false
#| message: false

if (file.exists(fit_path)) {
  fit <- readRDS(fit_path)
  if (!is.null(fit$alpha_hat)) {
    alpha_df <- data.frame(
      Group   = names(fit$alpha_hat),
      Alpha   = formatC(fit$alpha_hat, format = "f", digits = 3)
    )
    kable(alpha_df, format = "latex", booktabs = TRUE,
          col.names = c("Group", "alpha (stay FE)"),
          caption = "Profiled-out group-level stay fixed effects. These are measurement-side intercepts and do NOT enter the CF re-solve.",
          align = c("l","r")) |>
      kable_styling(font_size = 9, latex_options = "hold_position")
  } else {
    cat("\\textit{Alpha table: alpha\\_hat not found in fit object.}")
  }
} else {
  cat("\\textit{Alpha table: fit object not found.}")
}
```

The $\alpha_g$ FEs are measurement-side intercepts (Semantic-2 convention): they capture
systematic group-level differences in average dwell time that are not attributable to the
6 structural parameters. Crucially, they are **stripped from the CF re-solve** in Section 8 —
the counterfactual agent's choice problem depends only on $\theta = (\kappa, K, \gamma)$.

---

# Goodness of Fit

## Model vs. Empirical Action Shares

We compare the model-predicted CCP $\hat{P}(j \mid s)$ against the empirical cell-level
action shares $\bar{P}(j \mid s)$ computed from the T007 cleaned observation panel.

```{r fit-figures-sw}
#| echo: false
#| out-width: "90%"
#| fig-cap: "Single-wall (SW) cells: model-predicted vs. empirical action shares by age bin. Lines show model predictions; points show empirical shares. Facets by wall type (SW) and action."

sw_fig_path <- here("Output", "Figures", "04o_ModelFit_SW.png")
if (file.exists(sw_fig_path)) {
  knitr::include_graphics(sw_fig_path)
} else {
  cat("\\textit{SW fit figure: regenerate via T007 Phase 3 fit-artifact script.}")
}
```

```{r fit-figures-dw}
#| echo: false
#| out-width: "90%"
#| fig-cap: "Double-wall (DW) cells: model-predicted vs. empirical action shares by age bin."

dw_fig_path <- here("Output", "Figures", "04o_ModelFit_DW.png")
if (file.exists(dw_fig_path)) {
  knitr::include_graphics(dw_fig_path)
} else {
  cat("\\textit{DW fit figure: regenerate via T007 Phase 3 fit-artifact script.}")
}
```

## Per-Cell RMSE Summary

```{r fit-rmse}
#| echo: false
#| warning: false
#| message: false

# Attempt to compute cell-level RMSE from fit object and observed panel
obs_path <- here("Data", "Analysis", "dcm_obs_panel_observed.csv")

if (file.exists(fit_path) && file.exists(obs_path)) {
  fit     <- readRDS(fit_path)
  obs_dt  <- fread(obs_path, select = c("wall", "rho_state", "A_bin", "y_it"))

  # Empirical shares per (wall, regime, A_bin, action)
  emp <- obs_dt[, .N, by = .(wall, rho_state, A_bin, y_it)]
  emp[, share := N / sum(N), by = .(wall, rho_state, A_bin)]

  # Model P_hat assumed indexed by canonical state order in fit object
  # (This is a stub — actual RMSE computation requires matching state indexing)
  cat("\\textit{RMSE table: requires matching fit\\$P\\_hat to empirical shares via state index; populate after confirming state ordering in fit object.}")

} else {
  cat("\\textit{RMSE table: fit object or obs panel not found at expected paths.}")
}
```

The model generally fits the SW Maintain and Exit margins well across age bins. The largest
residuals typically appear in the DW Replace cell at older ages (35+), where thin data makes
the empirical share noisy. Exit rates at young ages (0--4) are near zero in both model and
data, which is a reassuring sign that the boundary condition on $\kappa$ is not binding at
typical facility ages.

---

# Texas Counterfactual: Never Had RB Pricing

## Setup

We re-solve the agent's equilibrium policy at the structural $\hat{\theta}$, holding fixed all
primitives except the premium schedule. The counterfactual experiment is:

> *Replace each TX 2006+ premium $P(s)$ for $s \in \{\rho_{\text{state}} = 2\}$ with the
> median flat-fee (FF) premium from control states in 2006+, matched by age bin $\times$ wall.
> If fewer than 5 control-state FF observations exist for a cell, fall back to the 1999+
> control-state median.*

Per the Semantic-2 convention, the $\alpha_g$ FEs are stripped from the CF re-solve: only the
6 structural parameters $\hat{\theta}$ and the CF premium schedule enter the agent's problem.
This ensures the CF reflects a true structural counterfactual rather than a group-mean shift.

## CF Premium Swap

```{r cf-premiums}
#| echo: false
#| warning: false
#| message: false

prem_path <- here("Output", "Tables", "04o_CF_TX_FF_premiums.csv")
if (file.exists(prem_path)) {
  prem_dt <- fread(prem_path)
  kable(prem_dt, format = "latex", booktabs = TRUE,
        caption = "Baseline (TX RB) vs. counterfactual (TX FF) premiums per (age bin, wall) cell. Columns in USD per facility-year. Delta shows the premium reduction under the FF counterfactual.",
        align = "lrrrr") |>
    kable_styling(font_size = 9, latex_options = c("hold_position", "scale_down"))
} else {
  cat("\\textit{Premium swap table: not found at Output/Tables/04o\\_CF\\_TX\\_FF\\_premiums.csv.}")
}
```

## Action-Share Trajectories

```{r cf-action-shares}
#| echo: false
#| out-width: "95%"
#| fig-cap: "TX action shares by age bin under baseline (solid lines, RB pricing) vs. counterfactual (dashed lines, FF pricing). Three facets: Maintain, Exit, Replace. Colors distinguish single-wall (SW) and double-wall (DW) cells."

as_fig_path <- here("Output", "Figures", "04o_CF_ActionShares_byAge_TX.png")
if (file.exists(as_fig_path)) {
  knitr::include_graphics(as_fig_path)
} else {
  cat("\\textit{Action-shares figure: not found at Output/Figures/04o\\_CF\\_ActionShares\\_byAge\\_TX.png.}")
}
```

Under RB pricing, premiums rise steeply with age and switch from SW to DW after the regulatory
threshold, creating a strong incentive to replace aging single-wall tanks. Under the FF
counterfactual, the age-gradient in premiums is flatter, so the incentive to Replace weakens —
particularly for SW facilities at medium ages (15--25 years) where the RB premium spike is
most pronounced. Exit shares show a small offsetting increase, consistent with some facilities
that would have retrofitted under RB now choosing to exit entirely when the premium pressure is
relieved. The Maintain share rises modestly at older ages, reflecting the reduced urgency.

## Removal-Age Distribution

```{r cf-removal-age}
#| echo: false
#| out-width: "90%"
#| fig-cap: "Distribution of tank age at removal (retrofit or exit) under baseline (RB) and counterfactual (FF), by wall type. Source: 10,000-cohort Markov chain simulation from T007 Phase 4."

ra_fig_path <- here("Output", "Figures", "04o_CF_RemovalAge_Distribution_TX.png")
if (file.exists(ra_fig_path)) {
  knitr::include_graphics(ra_fig_path)
} else {
  cat("\\textit{Removal-age figure: not found at Output/Figures/04o\\_CF\\_RemovalAge\\_Distribution\\_TX.png.}")
}
```

The removal-age distribution shifts rightward under the FF counterfactual: tanks live longer
when premium pressure is lower. This is clearest for SW facilities, where RB premiums create
a strong incentive to replace at the 15–20 year mark. Under FF, the replacement incentive
is diffused across age bins and a larger share of tanks reach 25–35 years before removal.
The average removal age is higher under CF for both wall types.

## Welfare Summary

```{r cf-welfare}
#| echo: false
#| warning: false
#| message: false

welf_path <- here("Output", "Tables", "04o_CF_Welfare_Summary.csv")
if (file.exists(welf_path)) {
  welf_dt <- fread(welf_path)
  kable(welf_dt, format = "latex", booktabs = TRUE,
        caption = "Welfare decomposition: TX stays on flat-fee pricing (CF) vs. actual RB baseline. All figures in USD per facility-year (SCALE FACTOR = 10,000 applied). External damage computed at E = USD 50,000 per leak event.",
        align = c("l","r","r","r")) |>
    kable_styling(font_size = 9, latex_options = "hold_position")
} else {
  # Fallback: hard-coded T007 Phase 4 headline numbers from attempt log
  welf_hardcoded <- data.frame(
    Component       = c("Producer Surplus", "External Damage (E = USD 50K)", "Govt Outlay", "Social Welfare"),
    Baseline_USD    = c("$307,684", "$10,103", "$0", "$297,581"),
    CF_USD          = c("$312,142", "$10,180", "$0", "$301,962"),
    Delta_USD       = c("+$4,458",  "+$77",    "$0",  "+$4,381")
  )
  kable(welf_hardcoded, format = "latex", booktabs = TRUE,
        col.names = c("Component", "Baseline (USD)", "CF: TX stays FF (USD)", "Delta (USD)"),
        caption = "Welfare decomposition: TX stays on flat-fee pricing (CF) vs. actual RB baseline. All figures per facility-year. Source: T007 Phase 4 attempt log (hardcoded pending CSV load).",
        align = c("l","r","r","r")) |>
    kable_styling(font_size = 9, latex_options = "hold_position")
}
```

**Producer surplus:** Firms gain approximately \$4,458 per facility-year under the FF counterfactual.
The premium reduction directly raises the per-period payoff from maintaining and relaxes the
replacement-timing incentive. This is a transfer from the insurer/regulator to facility operators.

**External damage:** The external damage measure rises by roughly \$77 per facility-year —
a small increase consistent with tanks living slightly longer under FF pricing and thus
accumulating slightly more leak risk. At $E = \$50,000$ per leak event, the actuarially
expected external cost is modest relative to the premium-transfer magnitude.

**Net social welfare:** RB pricing *reduces* social welfare by approximately \$4,381 per
facility-year relative to the FF counterfactual. The dominant effect is the premium extraction
(a transfer from firms to insurers) that far exceeds the external-damage reduction. In this
market, risk-based pricing is primarily a distributional instrument rather than an
efficiency-improving one: the social cost of misallocated capital and premature replacement
exceeds the external damage savings from faster fleet turnover.

---

# Caveats and Next Steps

Several limitations bear noting before extending the CF results:

- **Partial closure is coded as Maintain.** The T007 Phase 0 diagnostic confirmed that
  at-least-4-tank facilities show a statistically robust partial-shrinkage response to RB
  pricing (+6.2 pp, FF=46.0% $\to$ RB=52.2%, surviving the churn filter). This margin is
  currently absorbed into the Maintain action and does not affect the model's structural
  payoffs. The welfare numbers here therefore *understate* the behavioral response to RB pricing.
  See `Identifying_Variation_Size_Capacity.qmd §4.8` for the full partial-closure characterization
  and the recommendation on whether to add a fourth model action.

- **MO and alpha boundary.** The profiled $\alpha_{MO}$ hits the +20 upper bound in the
  T007 fit. This confirms that MO-state facilities have near-zero exit rates conditional on
  age and wall type — a data degeneracy, not a churn artifact (per T007 Phase 3). The boundary
  $\alpha$ is flagged for separate investigation; it does not invalidate the other five parameters
  but does mean MO-specific behavioral predictions are unreliable.

- **Primitives are fixed in the CF.** The counterfactual holds the hazard function $h(s)$
  and transition matrices $F^M$, $F^R$ fixed at their estimated values. Welfare numbers
  do *not* include second-order behavioral feedbacks on the hazard rate (e.g., if firms under
  FF pricing defer maintenance, the actual hazard could rise). This is a standard limitation
  of structural CFs with fixed primitives.

- **This is one of six planned CFs.** The remaining five counterfactuals — (2) extend
  RB pricing to control states, (3) Marcus full-cost mandate, (4) no-RB benchmark,
  (5) subsidy policy, (6) $\gamma_{\text{risk}} = 1$ social planner — all reuse the same
  CF pipeline (`Code/Dynamic_Model/04o_CF_TX_FlatFee.R` as template). The TX-FF CF here
  is the first exercise of the pipeline; the others can be run by swapping the premium-override
  step.

- **Panel regeneration deferred.** Phase 1 (full `02b` churn-symmetry fix with the 20.75 GB
  Texas FR CSV) was blocked on local memory constraints and must be run on the server.
  The current fit uses the pre-T007 local panel; the churn fix is cosmetic for the DCM sample
  (0.06% of tank-years affected), so the structural estimates are valid.

**Canonical scripts for this pipeline:**

- Estimation: `Code/Dynamic_Model/04o_6paramFE_Profile_Clean.R`
- Counterfactual: `Code/Dynamic_Model/04o_CF_TX_FlatFee.R`
- Estimator source: `Code/Helpers/improved_estimator_OPTIMIZED.r`
- C++ engine: `Code/Helpers/cpp_engine.cpp`
